### 2018-05-18 abstract/abstract.tex
Added abstract draft

\documentclass{article}

\begin{document}

Paths on the Autonomous Systems (AS) graph of the Internet derived from the Border Gateway Protocol (BGP) can help researchers to understand the dynamics of Internet topology and to interpret how that topology may enable nations to enact censorship, surveillance, or other Internet control measures. Unfortunately datasets of these paths are not readily publicly available, and paths collected from measurements such as traceroute measurements or BGP probes tend to be incomplete. Simulation frameworks have been used to generate AS paths based on empirical AS graphs. We have reimplemented one such framework, BGPSim \cite{quicksand} in Python as an efficient, open source BGP path simulation tool called BGPSimPy. With this tool we produced publicly available routing tree datasets that will enable future research on AS level topology. Previous research endeavors in this direction have only identified chokepoints in single snapshots. We apply the path simulation and routing tree datasets to view snapshots of the Internet over multiple years in order to understand how chokepoints at the national level may evolve to facilitate censorship. As an application of the routing trees we look at multiple qualitative sources of Internet freedom and censorship in order to interpret the relationship between AS level topology and actual censorship activity.

\bibliographystyle{plain}
\bibliography{abstract}

\end{document}

### 2018-06-07 Paper/paper.tex
Added Paper Draft

%&pdflatex
\documentclass{article}

\usepackage{color}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }

 \newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

\graphicspath{{./images/}}

\begin{document}
\begin{multicols}{2}

\section{Abstract}     
Paths on the Autonomous Systems (AS) graph of the Internet
derived from the Border Gateway Protocol (BGP) can be used by researchers to
understand the dynamics of Internet topology and to interpret how that
topology may enable nations to enact censorship, surveillance, or other
Internet control measures. Unfortunately datasets of these paths are not
generally made publicly available, and paths collected from measurements such as
traceroute measurements or BGP probes tend to be incomplete. Simulation
frameworks have been used to generate AS paths based on empirical AS graphs.
We have reimplemented one such framework, BGPSim \cite{quicksand} in Python as
an efficient, open source BGP path simulation tool called BGPSimPy. With this
tool we produced publicly available routing tree datasets that will enable
future research on AS level topology. Previous research endeavors in this
direction have only identified chokepoints in single snapshots. We apply the
path simulation and routing tree datasets to view snapshots of the Internet
over multiple years in order to introduce a new technique to investigate the
evolution of the complex and dynamic AS level Internet topology. As an
application of the routing trees we compare multiple qualitative sources of
Internet freedom and censorship with chokepoint metrics in order to interpret
the relationship between AS level topology and actual censorship activity.

\section{Related Work}
The primary struggle facing Internet topology researchers is
a lack of ground truth data. Paths on the AS level of the Internet are
incomplete and sometimes inaccurate {\color{blue}cite?}. Additionally, AS business
relationships are confidential and not publicly available. In this absence of
solid empirical data, a technique for inferring AS relationships and a
technique for inferring BGP paths are needed to study AS-level topology and
evolution. The first effective relationship inference algorithm was developed
by Gao in \cite{gao}. This algorithm was limited in that it focused on
maximizing the occurrences of the valley-free rule on the AS graph, which has
been shown to not always hold {\color{blue}cite?}. Additionally, Gao's algorithm was
verified against a single tier-1 provider's ground truth data. A more
verifiable correct set of AS relationships was developed by the Center for
Applied Internet Data Analysis (CAIDA) \cite{CAIDApaper}. These relationships,
which were generated using Gao's algorithm and other inference techniques,
were verified against a larger ground truth dataset, and didn't require that
the valley-free rule always hold. This leads to more realistic relationships. AS
relationships do not tell the whole story of AS-level topology, however. In
order to identify chokepoints in the AS graph, it is necessary to determine
the paths that Internet traffic follows. Researchers either use a measurement
tool like traceroute or BGP AS-path announcement data to determine AS paths.
These techniques have been shown to yield an incomplete and innacurate set of
paths {\color{blue}cite?}. It is also important that paths from all ASes be utilized to
get a complete picture of Internet topology. One technique for generating such
a set of paths was developed by Gill et al. in \cite{quicksand}. The algorithm
developed in \cite{quicksand}, called BGPSim, uses a modified breadth-first
search on an AS graph with defined relationships, as in from the CAIDA
dataset. The breadth first search traverses the AS graph starting from a root
node, and the search traverses edges according to priorities related to the
relationships defined by Gao. The priorities, in the order they are integrated
into the algorithm, are: Shortest Path (SP), Local Preference (LP), and Tie-
break (TB). The resultant paths are ideal according to economic concerns of AS
admins.     
\par
Previous studies have used BGP path models to find ASes that
intercept a high fraction of paths. One such project in \cite{throats}
identified that 90\% of paths on the Internet could be intercepted with only 30
or so ASes. The researchers in \cite{throats} generated paths by first using
only paths to top websites as defined by the Alexa top websites project,
and then appending additional edges to those paths from the full
AS graph according to the principles defined by Gao \cite{gao}. Many of the
ASes that were found to intercept a large number of paths were found to be
within nations that conduct censorship. As an extension of these results,
another paper \cite{decoy} revealed that ASes that intercept many paths could
be utilized for decoy routing. This approach for identifying AS level
chokepoints has several potential pitfalls that are remedied by the approach
taken in this paper. First, the authors in \cite{throats} didn't consider that
many of the paths found from the Alexa top websites would have destinations in
nations that censor Internet traffic, particularly China due to its large
Internet population. This artificially inflates the chokepoint nature of
Chinese ASes. As an alternative, in this paper we consider paths from every
source-destination AS pair, and we make a distinction between in-to-out paths
and out-to-in paths. Secondly, chokepoints have previously only been
identified at a single snapshot of the Internet. This makes it difficult to
discuss the evolution of Internet topology, and whether or not chokepoints are
anomalous or common. Finally, the aforementioned approach for identifying
chokepoints does not allow a clear comparison between different nations. It
can be said that one nation controls a large portion of Internet paths, but
not how easily traffic directed through that country could be intercepted on a
national level. For this, some aggregate measure across all of the border ASes
within a country must be considered, as it is in this paper.     
\par
In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
to identify where keyword filtering occurred. They found that the most
effective ASes with which to deploy keyword filtering devices are those in the
backbone of the Chinese AS topology. A relevant contribution of
\cite{chinafiltering} is that, while most filtering occurs in border ASes,
some filtering is controlled by non-central provincial ASes. China has a
diverse strategy for Internet censorship, both targeting chokepoints and the
Chinese provincial network. The potential for various forms of censorship in
regards to various AS level topologies motivates the question: Is centralized
censorship or decentralized censorship more common? Instead of directly
identifying censorship devices on the AS graph, we instead quantify the
chokepoint potential of ASes on the national level, and then compare that with
qualitative Internet freedom measures and empirical censorship events.

\section{BGPSimPy}
In order to conduct an analysis on AS level topology, the paths
generated by the Border Gateway Protocol (BGP) have to be modeled. A
traditional approach is to use a tool like traceroute to collect paths via
measurement, or at a larger scale use a selection of BGP data collectors with
a tool like BGPStream \cite{bgpstream} \cite{traceroute}. Unfortunately, these
techniques require extensive collection time and are still susceptible to
errors and incomplete data. There is no readily available BGP path dataset
with which to study. Thus, researchers must turn to simulation to find
accurate BGP paths.     
\par One such simulation framework, the previously
mentioned BGPSim, was developed by Gill et al. \cite{quicksand} as a tool for
quickly estimating AS paths from an input AS graph. This AS graph could be
taken from a set of inferred AS relationships, such as the CAIDA AS
relationship dataset \cite{CAIDA}. BGPSim outputs routing trees, which contain
all paths from leaf ASes to the root AS, for each AS in the AS graph. The
routes are selected using a modified breadth-first search that selects with
the following priorities: shortest path, local preference (using the rules
from Gao \cite{Gao}), and finally tie break of all equally good paths. With
these priorities, BGPSim generates a realistic collection of AS level paths,
which are ideal for use in Internet topology studies.
\par 
Unfortunately, BGPSim is not ready to deploy immediately and easily, as would be necessary
for a tool dedicated to allow fast monitoring of the Internet's overall state.
This is largely due to the fact that BGPSim is written in C\# (affecting its
ability to be used in a cross-platform way) with DryadLinq (a deprecated C\#
library) for distributed computing. As a remedy to these problems, we have
reimplemented BGPSim under the name BGPSimPy. BGPSimPy is written purely in
python, making it cross platform and easy to integrate. Additionally,
distributed computing is now achieved with MPI using mpi4py, a framework that
is supported on many computer clusters.

\section{Generating Routing Trees}
AS level Internet topology research is plagued
by a lack of publicly available path data. Path inference is also
computationally expensive for the entire Internet, prohibiting continuous
research and Internet monitoring. Thus, we have chosen to release the routing
tree datasets generated for this paper. This is a major contribution to
Internet research, and will allow new endeavors to investigate Internet
topology relationships on an unprecedented scale. As a focus of this paper,
and an example application, we use these routing trees to examine topological
evolution of the Internet and its relationships to Internet freedom and
censorship.

\section{Applications of Routing Trees}
\subsection{Measuring Chokepoints in AS-level Paths}
Once we generated routing trees, we
could use the paths to identify chokepoints, or ASes that intercept large
portions of paths, on the AS graph. Specifically, we chose to evaluate border
ASes, or ASes that could reach an AS registered to a different nation with one
hop. By selecting border ASes we could locate the ASes that intercepted
traffic into a nation (via out-to-in paths), as in a foreign attempt to access
a domestic website, or ASes that intercepted traffic out of a nation, as in a
domestic attempt to access a foreign website (via in-to-out paths). We choose
to make a distinction between these two path types to better understand the
implications of the chokepoints belonging to each nation.
\par To evaluate border ASes we use the metric of chokepont potential, defined here to be the
ratio of the number of paths intercepted by a border AS to the number of paths
for that nation, as shown in equation \ref{eqn:chokePointPotential}. This
metric is calculated separately for both in-to-out paths and out-to-in paths.
This calculation is performed by traversing the routing trees and keeping
record of which border nodes intercept each path, as well as what path type is
currently being traversed with regards to the host nation of the border AS in
question. Once all paths are traversed, the totals are compiled together, with
a chokepoint potential value stored for each border AS.
\par 
The chokepoint potentials of the border ASes allowed us to investigate the state of the AS-
level topology of the Internet for any given set of initial AS relationships.
We first looked at how many ASes each nation is required to control to
intercept a percentage of paths. For this test, we chose 90\%, as this would
indicate a clear control of Internet paths for the given country and was also
the percentage used in \cite{throats}. Additionally, we determined this value
for each nation for multiple timestamps, yielding our initial picture of the
evolution of Internet topology. The results of this test for some nations of
interest are shown in figure \ref{fig:NumToIntercept}.

\begin{Figure}
	\centering
	\includegraphics[width=\linewidth]{NodesNeeded}
	\captionof{figure}{Number of border nodes needed to control 90\% of paths for various nations}\label{fig:NumToIntercept}
\end{Figure}

\par
These results give an idea of how the Internet has evolved for different nations. Some nations,
particularly China, have evolved to a state where it is significantly easier
to intercept paths. Others, for instance Brazil, have become more open with
regards to the chokepoint nature of border ASes. These results are even more
striking when one considers that the number of ASes globally has drastically
increased over the years, amplifying the effect of higher chokepoint
potentials. The number of ASes in the CAIDA AS relationship dataset in 2012
was 43,274 in January, while in January of 2018 the count was 60,006.
Next, we took a more fine-grained approach to investigate the dynamics and
evolution of AS-level chokepoints. For each nation, we sorted the border ASes
in descending order into a list, so that the first AS in the list was the AS
with the highest chokepoint potential. Then, we selected each AS from the list
and kept track of the ratio of paths intercepted. This test allowed us to
investigate the ease with which different nations could control paths at
multiple timestamps. Additionally, this experiment illuminates that the
topological state of a nation in regards to border ASes is often complex and
dynamic. Since nations possess wide variations in AS count, this data is best
visualized on a semi-logarithmic scale. The different timestamps are shown in
different colors, where red is the furthest in the past and blue is the most
recent. These results are depicted in \ref{fig:chokepoint}.

\begin{figure*}
	\centering
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_AU}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_BR}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_CN}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_DE}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_GB}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_IR}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{single_US}
	\end{subfigure}

	\caption{Ratio of Paths Intercepted for a certain number of Border Nodes for multiple nations}\label{fig:chokepoint}
\end{figure*}

\par 
The variation between the different nations highlights the importance of measuring
chokepoints on the national level. Trends exist within a nation, but nations
tend to have surprisingly unique AS graph dynamics. China shows a striking
increase in chokepoint potential. In order to control 90% of paths, China has
to utilize a strikingly few number of ASes. This suggests an ever
strengthening backbone in the Chinese Internet. Conversely, the trend that
Brazil's capability to intercept paths is decreasing is also evident in these
results. This could be the effect of an expansion of infrastructure, leading
to more border ASes and more diverse paths in and out of the country. {\color{blue}Maybe
there is a source to support this?} Another interesting element of these
results is the dynamic behavior for some nations. It seems that in some cases,
when it becomes easier to control many paths, it becomes more difficult to
control all paths. {\color{blue}Not sure of implications, needs work here}

\subsection{Comparing Internet Topology with Freedom On The Net}
While research in the past \cite{throats} has connected AS-level chokepoints with the capability of
a nation to conduct censorship, linking the topology based measurements to
other attempts at measuring Internet freedom has not been done before to the
best of our knowledge. Such a connection would bridge the gap between a
dynamic system driven by many factors, both automatic and instituted, and
qualitative classification of nations.
\par Freedom House's Freedom On The Net
(FOTN) report \cite{FOTN}, is an annual ranking and analysis of Internet
freedom, interpreted from multiple measures including obstacles to access,
limits on content, and violations of user rights. The FOTN includes a ranking
of nations by a freedom index, where a higher number indicates a less free
Internet for that nation. The index is calculated from a series of questions
that are evaluated according to various sources regarding each nation. These
questions encompass multiple categories, including government filtering and
censorship of web traffic. In order to investigate whether the topology of a
nations ASes is related to Internet freedom, we compare the FOTN freedom index
with the ease with which a nation can control AS level paths. In figure
\ref{fig:fotn} we plot the number of ASes required for a nation to control 90%
of paths with that nation's freedom index for multiple timestamps of AS
relationships.

\begin{figure*}
	\centering
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{fotn2012}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{fotn2015}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{fotn2016}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{fotn2017}
	\end{subfigure}

	\caption{Ratio of Paths Intercepted for a certain number of Border Nodes for multiple nations}\label{fig:fotn}
\end{figure*}

\par For each date we fit the data with an ordinary least
squares linear regression, and found the relationship to be statistically
significant (all p-values are less than 0.05). {\color{blue}need some better way to show
exact p-values} The relationship shows that more free nations, as determined
by FOTN scores, tend to have a more difficult to control AS-level topology, as
in border nodes that control fewer portions of paths in and out of that
country. This suggests that either the desire of national governments to
control Internet traffic has impacted the layout of AS relationships, the
topology of the modern Internet has evolved in a way to enable more successful
filtering and blocking, or a combination of both. Up to this point it has been
assumed that Internet topology and Internet censorship are related, but this
has not been supported by both a measure of Internet freedom and actual AS-
level paths.

\subsection{Comparing Internet Topology with Empirical Censorship Data}

Another
technique for relating our chokepoint metric with censorship is to look at
empirical censorship data. One such dataset is provided by the Open
Observatory of Network Interference (OONI) \cite{OONI}. OONI is a project
wherein Internet users around the world use a tool (the ooniprobe) to examine
evidence of Internet censorship. The ooniprobe allows users to run a series of
tests, such as an HTTP requests that might be known to be blocked by certain
nations. The ooniprobe then compares the results of tests with previously
determined unblocked examples of the same tests. Results where the test
differs from expected are considered anomalies, and further heuristics can
determine if the anomalies are likely to be censorship events.
\par We examined
18 days of OONI data from January in 2018. We recorded the ratio of anomalies
in measurements for several classes of anomalous behavior to total
measurements carried out in each nation. We used the web connectivity test
from OONI that records the following fields to measure connectivity: DNS,
HTTP\_FAIL, HTTP\_DIFF, and TCP. Each of these fields are recorded as true or
false, according to whether the test resulted in a potential censorship event.
We then plotted the ratio of potential censorship events in the dataset
against a national metric of chokepoint potential. In this case, we again
looked at the number of border nodes required to intercept 90\% of AS-level
paths. The results of this test are plotted in figure \ref{fig:ooni}. A
relationship like the one seen in the FOTN data is not as clear, perhaps due
to the uneven sampling across nations. It is clear, however, that nations can
be classified both by its chokepoint potential and the occurrence of possible
censorship events. A nation, for instance, with a high ratio of web
connectivity anomalies with a high chokepoint potential can be considered a
censorship active nation with a strongly hierarchical AS topology. Thus, this
analysis serves as a monitoring tool for the state of nations on the Internet
in regards to actual censorship implementation, and potential for further
censorship.

\begin{figure*}
	\centering
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{all_09}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{dns_09}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{http_diff_09}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{http_fail_09}
	\end{subfigure}
	\\
	\begin{subfigure}[b]{0.4\linewidth}
		\includegraphics[width=\linewidth]{tcp_09}
	\end{subfigure}

	\caption{Ratio of anomalies in OONI measurements compared to chokepoint potential}\label{fig:ooni}
\end{figure*}

\section{Discussion}

We have introduced an efficient technique for studying the AS-level topology of the Internet. Additionally,
we introduced the metric of chokepoint potential in order to evaluate the control that border nodes possess
over paths into and out of a nation. We have applied the routing trees generated by BGPSimPy along with the
chokepoint potential metric to present an overview of the evolution of the amount and strength of chokepoints
on the AS graph over the past 6 years. We also connected our analysis of AS-level chokepoints in a novel way to
both qualitative measures of Internet freedom and empirical censorship data. The former comparison reveals
a strong relationship between the chokepoint potential of a nation and its lack of Internet freedom, while the
former provides insight into the categories of nations, in regards to their implementations of censorship and
the layout of their AS relationships.

\end{multicols}

\bibliographystyle{plain}
\bibliography{paper}

\end{document}

### 2018-06-22 Paper/paper.tex
Updated Draft

@@
 generally made publicly available, and paths collected from measurements such as
-traceroute measurements or BGP probes tend to be incomplete. Simulation
+traceroute or BGP probes tend to be incomplete. Simulation
 frameworks have been used to generate AS paths based on empirical AS graphs.
-We have reimplemented one such framework, BGPSim \cite{quicksand} in Python as
+We have reimplemented and extended one such framework, BGPSim \cite{quicksand} in Python as
 an efficient, open source BGP path simulation tool called BGPSimPy. With this
 tool we produced publicly available routing tree datasets that will enable
-future research on AS level topology. Previous research endeavors in this
+future research on AS level topology. We introduce a metric, \textit{chokepoint potential},
+in order to use routing trees to interpret the capability for border ASes (those one hop away from
+an AS in a different nation) to enact censorship or surveillance. With the entire set of
+routing trees we demonstrate that we can expose the choke points of AS-level routing for
+any nation. Previous research endeavors in this
 direction have only identified chokepoints in single snapshots. We apply the
@@
 over multiple years in order to introduce a new technique to investigate the
-evolution of the complex and dynamic AS level Internet topology. As an
+evolution of the complex and dynamic AS level Internet topology. Through this approach we
+can more carefully evaluate the state of the Internet than was previously possible. As an
 application of the routing trees we compare multiple qualitative sources of
 Internet freedom and censorship with chokepoint metrics in order to interpret
-the relationship between AS level topology and actual censorship activity.
+the relationship between AS level topology and actual censorship activity, providing
+illustration of new ways to monitor which governments can easily control the flow of 
+information in their nations.
 
@@
 These techniques have been shown to yield an incomplete and innacurate set of
-paths {\color{blue}cite?}. It is also important that paths from all ASes be utilized to
-get a complete picture of Internet topology. One technique for generating such
+paths {\color{blue}cite?}. 
+\par
+In light of these issues, BGP researchers often employ more abstract models of BGP.
+For instance, the authors in \cite{converge} developed a simplification of BGP in order
+to estimate important convergence properties. In order to evaluate Internet chokepoints,
+an accurate set of paths between all AS pairs must generated. One technique for generating such
 a set of paths was developed by Gill et al. in \cite{quicksand}. The algorithm
@@
 break (TB). The resultant paths are ideal according to economic concerns of AS
-admins.     
+administrators.     
 \par
@@
 qualitative Internet freedom measures and empirical censorship events.
+\par
+We are not the first to investigate the relationship between Internet freedom and
+AS-level topology. Similar techniques have been used to classify nations according
+to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
+single moment in time, however, making the results limited in terms of stability and predictability.
+Additionaly, previous work has not used country level chokepoints as the link between Internet topology
+and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
+to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
+to include more nations. Additionally, they didn't include the United States and Russia in their experiments because they were
+outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting metric for
+understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
+The Net score as our metric for Internet freedom. Through our techniques, we provide a simple metric that not only 
+sheds light on the relationship between topology and Internet freedom, but reveals currently free
+nations that could easily implement censorship if their governments decided to.
 
@@
 the following priorities: shortest path, local preference (using the rules
-from Gao \cite{Gao}), and finally tie break of all equally good paths. With
+from Gao \cite{gao}), and finally tie break of all equally good paths. With
 these priorities, BGPSim generates a realistic collection of AS level paths,
@@
 \subsection{Measuring Chokepoints in AS-level Paths}
-Once we generated routing trees, we
-could use the paths to identify chokepoints, or ASes that intercept large
-portions of paths, on the AS graph. Specifically, we chose to evaluate border
+
+It is well documented that the AS graph is composed of a hierarchy of ASes, such that the ASes
+distributed in tiers. This tier list includes a very small set of tier-1 ASes
+that are able to provide the highest levels of connectivity to the rest of the Internet \cite{hierarchy}.
+This hierarchy is generally viewed in a way agnostic to national boundaries. It is
+possible, however, that more ASes exhibit potential to control information flow in the sense that
+chokepoints can exist at the national level. While these chokepoints might not provide extensive connectivity
+to the entire set of ASes, they may provide similar connectivity to the interior ASes of a particular nation. 
+Thus, a technique to conduct a fine-grained analysis of the number of paths intercepted by ASes that provide access
+in and out of nations is a worthwhile pursuit. Previous works have also not provided an analysis of the development of
+the AS hierarchy over time. The monitoring of national chokepoints over time is relevant to predicting the stability of the
+future Internet, identifying key censorship bottlenecks, and understanding the potential for national governments to conduct
+surveillance on foreign traffic.
+
+\par
+
+Once we generated routing trees using BGPSimPy, we
+used the paths to identify national chokepoints. Specifically, we chose to evaluate border
 ASes, or ASes that could reach an AS registered to a different nation with one
@@
 implications of the chokepoints belonging to each nation.
+
 \par To evaluate border ASes we use the metric of chokepont potential, defined here to be the
-ratio of the number of paths intercepted by a border AS to the number of paths
-for that nation, as shown in equation \ref{eqn:chokePointPotential}. This
-metric is calculated separately for both in-to-out paths and out-to-in paths.
+ratio of the number of paths intercepted by a border AS to the number of total path interceptions
+possible for that nation. If $P_c$ is the set of all paths (either in-to-out or out-to-in) belonging to nation $c$ and 
+$B_p$ is the set of all border nodes within a path $p$, the chokepoint potential for a border node $b$ is defined in equation \ref{eqn:chokePointPotential}
+This metric is calculated separately for both in-to-out paths and out-to-in paths.
 This calculation is performed by traversing the routing trees and keeping
@@
 a chokepoint potential value stored for each border AS.
+
+\begin{equation}\label{eqn:chokePointPotential}
+CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
+\end{equation}
+
 \par 
@@
 indicate a clear control of Internet paths for the given country and was also
-the percentage used in \cite{throats}. Additionally, we determined this value
+the percentage used in past studies \cite{throats}. Additionally, we determined this value
 for each nation for multiple timestamps, yielding our initial picture of the
@@
 tend to have surprisingly unique AS graph dynamics. China shows a striking
-increase in chokepoint potential. In order to control 90% of paths, China has
+increase in chokepoint potential. In order to control 90\% of paths, China has
 to utilize a strikingly few number of ASes. This suggests an ever
@@
 \subsection{Comparing Internet Topology with Freedom On The Net}
-While research in the past \cite{throats} has connected AS-level chokepoints with the capability of
-a nation to conduct censorship, linking the topology based measurements to
-other attempts at measuring Internet freedom has not been done before to the
-best of our knowledge. Such a connection would bridge the gap between a
-dynamic system driven by many factors, both automatic and instituted, and
-qualitative classification of nations.
+While research in the past has connected AS-level topology with Internet freedom 
+and censorship \cite{throats}\cite{politicsrouting}, the evolution of this relationship has
+not been investigated across multiple years. We also provide the first attempt to connect the
+metric of chokepoint potential with the Internet freedom of nations, creating insight into the
+connection between the state of Internet topology and efforts to limit Internet freedoms.
 \par Freedom House's Freedom On The Net
@@
 
-\end{document}
+\end{document}\grid

### 2018-06-28 Paper/paper.tex
Added Updated Draft

--- 2018-06-07 Paper/paper.tex
@@
 
+\newcommand{\toolname}{{\color{blue}TOOL\_NAME }}
+
+%Titles:
+\title{The Evolution of National AS Chokepoints and Their Connection to
+Internet Freedom}
+
 \graphicspath{{./images/}}
+\setlength{\parskip}{0pt}
 
 \begin{document}
+\raggedbottom
+\maketitle
 \begin{multicols}{2}
@@
 generally made publicly available, and paths collected from measurements such as
-traceroute measurements or BGP probes tend to be incomplete. Simulation
-frameworks have been used to generate AS paths based on empirical AS graphs.
-We have reimplemented one such framework, BGPSim \cite{quicksand} in Python as
-an efficient, open source BGP path simulation tool called BGPSimPy. With this
-tool we produced publicly available routing tree datasets that will enable
-future research on AS level topology. Previous research endeavors in this
-direction have only identified chokepoints in single snapshots. We apply the
+traceroute or BGP probes tend to be incomplete. Simulation
+frameworks have been used to generate AS paths based on inferred AS relationships.
+We introduce \toolname, a suite of open source, cross platform, and efficient
+tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
+control Internet traffic, either through censorship or surveillance.
+Previous research endeavors in this
+direction have only identified chokepoints in single snapshots. We apply our
 path simulation and routing tree datasets to view snapshots of the Internet
 over multiple years in order to introduce a new technique to investigate the
-evolution of the complex and dynamic AS level Internet topology. As an
-application of the routing trees we compare multiple qualitative sources of
-Internet freedom and censorship with chokepoint metrics in order to interpret
-the relationship between AS level topology and actual censorship activity.
+evolution of the complex and dynamic AS level Internet topology. Through this approach we
+can more carefully evaluate the state of the Internet than was previously possible. As an
+application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
+Internet freedom and with our chokepoint potential measure in order to interpret
+the relationship between AS-level topology and actual censorship activity, providing an
+illustration of new ways to monitor which governments can easily control the flow of 
+information in their nations and whether they are acting on that potential.
+
+\section{Introduction}
+
+%The Internet is Dynamic and growing
+The Autonomous Systems layer of the Internet has grown and changed dramatically over its history.
+In the early 2000s, there were around 10,000 ASes connected to the global AS graph. Today there are
+over 60,000 of these ASes. The locations and relationships of these ASes determine how many chokepoints
+of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
+This rapid expansion of the AS graph suggests a dynamic system with behavior that is important for
+understanding things like path robustness, AS hierarchies, and the potential for governments to control
+information as it flows in and out of their borders. It is certain that structrural qualities of national
+AS graphs have evolved differently from nation to nation, whether that be due to economic decision making, infrastructural
+necessities, or efforts to build a powerful censorship and surveillance network.
+
+%Current Events
+\par
+The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
+organize revolutionary movements, and connect with other nations' governments and citizens
+to seek aid. Much of the organization, revolutionary momentum, and broadcasting
+of happenings during the Arab Spring can be attributed to social media communications
+through the Internet \cite{arabspring}. Because of this potential, national governments may
+take interest in maximizing their ability to control the flow of information on the Internet.
+%Censorship and Surveillance are growing phenomena
+\par
+Internet censorship and surveillance have become profound social and technical issues facing the world. The portion of the
+global population that uses the Internet has increased substantially in recent years. According to the International Telecommunication Union,
+the number of individual users of the Internet has increased from 1.024 million in 2005 to 3.578 million in 2017 \cite{itu}. Of the Internet
+population, a majority of them are subject to some lack of Internet freedom. The 2017 Freedom on the Net report from freedom house reports that
+64\% of Internet users belong to a nation with Internet that is not free or partly free \cite{FOTN}. It is evident that Internet control is
+a powerful force for governments to wield. Whether or not AS-level topology supports these efforts is a research question pivotal to
+an understanding of the dynamics of Internet censorship and surveillance.
+
+%Diverse Censorship Architecture
+\par
+Every nation has a different layout and count of ASes. The censorship and surveillance strategies of nations also differ. For instance,
+China both conducts keyword filtering in border ASes and in internal provincial nodes \cite{chinafiltering}. Another nation that conducts
+extensive Internet censorship, Iran, has been found to route all of its Internet traffic through a centralized facility \cite{irancensor}.
+Having views into the AS topology of a nation, then, will both help researchers identify nations that could easily conduct censorship and
+also provide possible insight into what kind of censorship is likely being conducted. The measure of chokepoint potential, defined later in this paper,
+is an effective way to capture the important properties of border ASes related to these capabilities.
+
+% Major Contributions
+\par
+The primary contributions of our work are as follows: \textbf{1.} We introduce the measure of chokepoint potential and 
+motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic. \textbf{2.} We
+provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
+allowing for a comparision between nations. \textbf{3.} We develop a new tool, \toolname, that provides the capability for
+efficiently evaluating chokepoints for a given state of the Internet. Finally, \textbf{4.}, we show that our chokepoint measure
+has a significant relationship to Internet freedom as measured by a qualitative source.
+
+% Rest of Paper Layout
+\par
+
+\section{Background}
+
+The AS-layer of the Internet is the highest level of organization of the Internet. ASes generally represent ISP networks,
+large university networks, or government entities. Interdomain routing between these ASes is governed by BGP, and
+individual ASes may have local preferences for how they wish to send packets along, but they cannot govern the paths
+selected by other ASes. AS topology researchers often use measurement tools to create a picture of the global Internet.
+These measurement tools, for instance BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute \cite{traceroute}, suffer from
+incompleteness and innacuracies \cite{tracerouteProblems}.
+\par
+AS-level studies are further complicated by the lack of ground truth data for AS relationships and BGP paths. For AS relationships,
+inference is often used based on economic considerations to classify the relationships of ASes. This was first done by Gao \cite{gao} by
+maximizing the occurances of certain economic rules on the AS graph by choosing a particular set or relationships. This technique was only
+evaluated against a single ISPs set of true relationships, however. As part of the CAIDA project, this inference technique was extended, leading
+to the CAIDA AS-relationship dataset \cite{CAIDApaper}, \cite{CAIDA}. We choose these relationships for our purposes, and they are the current
+research standard.
+\par
+Finding BGP paths is more of a challenge. Packet based simulations, wherein BGP is simulated directly, are computationally infeasible for the
+scales relevant to a global study. An accurate, but realistic, simulation technique must be chosen to provide useful research potential in this regard.
+The BGPSim algorithm \cite{quicksand} is one such simulation technique that is suitable for this study's purposes. BGPSim takes a set of AS relationships
+as input, such as those provided by the CAIDA dataset \cite{CAIDA}, and returns routing trees based off of these relationships. These paths are found via
+a modified breadth-first seach (BFS) algorithm. The BFS adds edges to routing trees first according to local preference (LP), then shortest path (SP), and finally tiebreak (TB).
+A resulting routing tree contains all the equally reasonable paths (according to economic concerns) that exist between source ASes (within the routing tree) to 
+destination ASes (the root of the routing tree). We find this technique suitable for our purposes, so we extend BGPSim 
+(in ways explained further in this paper) as part of \toolname.
+
+\section{Chokepoint Potential}
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{chokepoint}
+	\captionof{figure}{Chokepoint potential example. ASes A,B, and C are all border nodes. 
+						AS D is an internal node. ASes E and F are both external nodes.
+						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
+\end{Figure}
+
+% Why Border ASes?
+In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
+ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
+We argue that this is intuitive because border
+ASes are the first opportunity for nations to censor incoming traffic and the last opportunity for nations to block access to
+foreign websites from within their borders. Additionally, while internal chokepoints may intercept many paths, those paths are required
+to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This simplification has
+the added benefit of making calculations more efficient.
+
+% What is our measure, why does it make sense?
+\par
+We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
+to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
+If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
+chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
+the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
+seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
+
+% Formal Definition
+\begin{equation}\label{eqn:chokePointPotential}
+CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
+\end{equation}
+
+\par
+Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
+the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
+flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
+to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
+censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
+
+\section{\toolname}
+
+% Preprocessing and Data pipeline
+\toolname is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
+\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
+to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
+are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
+IP to ASN whois service \cite{cymru}.
+
+% Routing Tree Algorithm
+\par
+In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
+our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
+they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
+framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
+saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
+routing tree algorithm that is ready to work on most hardware.
+
+% Chokepoint calculation
+\par
+Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
+we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
+traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
+necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
+
+\section{Experimental Setup}
+
+% Chokepoint Evolution
+There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some questions remain unanswered however. For instance, to what
+extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
+powerful chokepoints, or becoming more evenly accessible?
+\par
+To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
+change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
+For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border node per snapshot. As a result we can investigate
+how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
+With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
+
+% Chokepoints and Internet Freedom
+If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
+the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
+and it would increase the value in monitoring chokepoint potentials across the globe.
+\par
+We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
+the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
+to 100 (the least free).
+ To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
+Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
+looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
+with each nation's (of those recorded by Freedom House) freedom on the net score.
+
+\section{Results}
+\subsection{Nations Over Time}
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{single_CN}
+	\captionof{figure}{The evolution of China's in-to-out Chokepoint Potential}\label{fig:ChinaChokePoint}
+\end{Figure}
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{single_US}
+	\captionof{figure}{The evolution of the US's Chokepoint Potential}\label{fig:USChokePoint}
+\end{Figure}
+
+To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
+so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
+the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
+of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
+changes over time.
+
+\par
+For instance, consider figures \ref{fig:ChinaChokePoint} and \ref{fig:USChokePoint}. 
+The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily)
+is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
+farther in the past and more blue lines being more recent.
+
+\par
+The United States and China are shown here to exhibit their dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
+it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
+United States. This result is entirely expected. Of more interest is the trends that can be seen over time. China clearly has developed an AS topology such that very few
+ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
+more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
+the very top ASes.
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{single_DE}
+	\captionof{figure}{The evolution of Germany's in-to-out Chokepoint Potential}\label{fig:GermChokePoint}
+\end{Figure}
+
+\par
+Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in figure \ref{fig:GermChokePoint}. For Germany, a fairly constant
+trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{NodesNeeded}
+	\captionof{figure}{Number of ASes to intercept 90\% of ito paths for multiple nations over time}\label{fig:nodes}
+\end{Figure}
+
+\par
+In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
+for each timestamp recorded. These results are detailed in figure \ref{fig:nodes}.
+
+\subsection{Internet Freedom}
+
+In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
+of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
+we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
+with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
+
+\begin{Figure}
+	\centering
+	\includegraphics[width=\linewidth]{fotn2017}
+	\captionof{figure}{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017)}\label{fig:fotn2017}
+\end{Figure}
+
+\par
+For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
+to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
+of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
+few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
+is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
+that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
 
 \section{Related Work}   
-The primary struggle facing Internet topology researchers is
-a lack of ground truth data. Paths on the AS level of the Internet are
-incomplete and sometimes inaccurate {\color{blue}cite?}. Additionally, AS business
-relationships are confidential and not publicly available. In this absence of
-solid empirical data, a technique for inferring AS relationships and a
-technique for inferring BGP paths are needed to study AS-level topology and
-evolution. The first effective relationship inference algorithm was developed
-by Gao in \cite{gao}. This algorithm was limited in that it focused on
-maximizing the occurrences of the valley-free rule on the AS graph, which has
-been shown to not always hold {\color{blue}cite?}. Additionally, Gao's algorithm was
-verified against a single tier-1 provider's ground truth data. A more
-verifiable correct set of AS relationships was developed by the Center for
-Applied Internet Data Analysis (CAIDA) \cite{CAIDApaper}. These relationships,
-which were generated using Gao's algorithm and other inference techniques,
-were verified against a larger ground truth dataset, and didn't require that
-the valley-free rule always hold. This leads to more realistic relationships. AS
-relationships do not tell the whole story of AS-level topology, however. In
-order to identify chokepoints in the AS graph, it is necessary to determine
-the paths that Internet traffic follows. Researchers either use a measurement
-tool like traceroute or BGP AS-path announcement data to determine AS paths.
-These techniques have been shown to yield an incomplete and innacurate set of
-paths {\color{blue}cite?}. It is also important that paths from all ASes be utilized to
-get a complete picture of Internet topology. One technique for generating such
-a set of paths was developed by Gill et al. in \cite{quicksand}. The algorithm
-developed in \cite{quicksand}, called BGPSim, uses a modified breadth-first
-search on an AS graph with defined relationships, as in from the CAIDA
-dataset. The breadth first search traverses the AS graph starting from a root
-node, and the search traverses edges according to priorities related to the
-relationships defined by Gao. The priorities, in the order they are integrated
-into the algorithm, are: Shortest Path (SP), Local Preference (LP), and Tie-
-break (TB). The resultant paths are ideal according to economic concerns of AS
-admins.     
 \par
@@
 qualitative Internet freedom measures and empirical censorship events.
-
-\section{BGPSimPy}
-In order to conduct an analysis on AS level topology, the paths
-generated by the Border Gateway Protocol (BGP) have to be modeled. A
-traditional approach is to use a tool like traceroute to collect paths via
-measurement, or at a larger scale use a selection of BGP data collectors with
-a tool like BGPStream \cite{bgpstream} \cite{traceroute}. Unfortunately, these
-techniques require extensive collection time and are still susceptible to
-errors and incomplete data. There is no readily available BGP path dataset
-with which to study. Thus, researchers must turn to simulation to find
-accurate BGP paths.     
-\par One such simulation framework, the previously
-mentioned BGPSim, was developed by Gill et al. \cite{quicksand} as a tool for
-quickly estimating AS paths from an input AS graph. This AS graph could be
-taken from a set of inferred AS relationships, such as the CAIDA AS
-relationship dataset \cite{CAIDA}. BGPSim outputs routing trees, which contain
-all paths from leaf ASes to the root AS, for each AS in the AS graph. The
-routes are selected using a modified breadth-first search that selects with
-the following priorities: shortest path, local preference (using the rules
-from Gao \cite{Gao}), and finally tie break of all equally good paths. With
-these priorities, BGPSim generates a realistic collection of AS level paths,
-which are ideal for use in Internet topology studies.
 \par
-Unfortunately, BGPSim is not ready to deploy immediately and easily, as would be necessary
-for a tool dedicated to allow fast monitoring of the Internet's overall state.
-This is largely due to the fact that BGPSim is written in C\# (affecting its
-ability to be used in a cross-platform way) with DryadLinq (a deprecated C\#
-library) for distributed computing. As a remedy to these problems, we have
-reimplemented BGPSim under the name BGPSimPy. BGPSimPy is written purely in
-python, making it cross platform and easy to integrate. Additionally,
-distributed computing is now achieved with MPI using mpi4py, a framework that
-is supported on many computer clusters.
-
-\section{Generating Routing Trees}
-AS level Internet topology research is plagued
-by a lack of publicly available path data. Path inference is also
-computationally expensive for the entire Internet, prohibiting continuous
-research and Internet monitoring. Thus, we have chosen to release the routing
-tree datasets generated for this paper. This is a major contribution to
-Internet research, and will allow new endeavors to investigate Internet
-topology relationships on an unprecedented scale. As a focus of this paper,
-and an example application, we use these routing trees to examine topological
-evolution of the Internet and its relationships to Internet freedom and
-censorship.
-
-\section{Applications of Routing Trees}
-\subsection{Measuring Chokepoints in AS-level Paths}
-Once we generated routing trees, we
-could use the paths to identify chokepoints, or ASes that intercept large
-portions of paths, on the AS graph. Specifically, we chose to evaluate border
-ASes, or ASes that could reach an AS registered to a different nation with one
-hop. By selecting border ASes we could locate the ASes that intercepted
-traffic into a nation (via out-to-in paths), as in a foreign attempt to access
-a domestic website, or ASes that intercepted traffic out of a nation, as in a
-domestic attempt to access a foreign website (via in-to-out paths). We choose
-to make a distinction between these two path types to better understand the
-implications of the chokepoints belonging to each nation.
-\par To evaluate border ASes we use the metric of chokepont potential, defined here to be the
-ratio of the number of paths intercepted by a border AS to the number of paths
-for that nation, as shown in equation \ref{eqn:chokePointPotential}. This
-metric is calculated separately for both in-to-out paths and out-to-in paths.
-This calculation is performed by traversing the routing trees and keeping
-record of which border nodes intercept each path, as well as what path type is
-currently being traversed with regards to the host nation of the border AS in
-question. Once all paths are traversed, the totals are compiled together, with
-a chokepoint potential value stored for each border AS.
-\par 
-The chokepoint potentials of the border ASes allowed us to investigate the state of the AS-
-level topology of the Internet for any given set of initial AS relationships.
-We first looked at how many ASes each nation is required to control to
-intercept a percentage of paths. For this test, we chose 90\%, as this would
-indicate a clear control of Internet paths for the given country and was also
-the percentage used in \cite{throats}. Additionally, we determined this value
-for each nation for multiple timestamps, yielding our initial picture of the
-evolution of Internet topology. The results of this test for some nations of
-interest are shown in figure \ref{fig:NumToIntercept}.
-
-\begin{Figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesNeeded}
-	\captionof{figure}{Number of border nodes needed to control 90\% of paths for various nations}\label{fig:NumToIntercept}
-\end{Figure}
-
-\par
-These results give an idea of how the Internet has evolved for different nations. Some nations,
-particularly China, have evolved to a state where it is significantly easier
-to intercept paths. Others, for instance Brazil, have become more open with
-regards to the chokepoint nature of border ASes. These results are even more
-striking when one considers that the number of ASes globally has drastically
-increased over the years, amplifying the effect of higher chokepoint
-potentials. The number of ASes in the CAIDA AS relationship dataset in 2012
-was 43,274 in January, while in January of 2018 the count was 60,006.
-Next, we took a more fine-grained approach to investigate the dynamics and
-evolution of AS-level chokepoints. For each nation, we sorted the border ASes
-in descending order into a list, so that the first AS in the list was the AS
-with the highest chokepoint potential. Then, we selected each AS from the list
-and kept track of the ratio of paths intercepted. This test allowed us to
-investigate the ease with which different nations could control paths at
-multiple timestamps. Additionally, this experiment illuminates that the
-topological state of a nation in regards to border ASes is often complex and
-dynamic. Since nations possess wide variations in AS count, this data is best
-visualized on a semi-logarithmic scale. The different timestamps are shown in
-different colors, where red is the furthest in the past and blue is the most
-recent. These results are depicted in \ref{fig:chokepoint}.
-
-\begin{figure*}
-	\centering
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_AU}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_BR}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_CN}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_DE}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_GB}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_IR}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{single_US}
-	\end{subfigure}
-
-	\caption{Ratio of Paths Intercepted for a certain number of Border Nodes for multiple nations}\label{fig:chokepoint}
-\end{figure*}
-
-\par 
-The variation between the different nations highlights the importance of measuring
-chokepoints on the national level. Trends exist within a nation, but nations
-tend to have surprisingly unique AS graph dynamics. China shows a striking
-increase in chokepoint potential. In order to control 90% of paths, China has
-to utilize a strikingly few number of ASes. This suggests an ever
-strengthening backbone in the Chinese Internet. Conversely, the trend that
-Brazil's capability to intercept paths is decreasing is also evident in these
-results. This could be the effect of an expansion of infrastructure, leading
-to more border ASes and more diverse paths in and out of the country. {\color{blue}Maybe
-there is a source to support this?} Another interesting element of these
-results is the dynamic behavior for some nations. It seems that in some cases,
-when it becomes easier to control many paths, it becomes more difficult to
-control all paths. {\color{blue}Not sure of implications, needs work here}
-
-\subsection{Comparing Internet Topology with Freedom On The Net}
-While research in the past \cite{throats} has connected AS-level chokepoints with the capability of
-a nation to conduct censorship, linking the topology based measurements to
-other attempts at measuring Internet freedom has not been done before to the
-best of our knowledge. Such a connection would bridge the gap between a
-dynamic system driven by many factors, both automatic and instituted, and
-qualitative classification of nations.
-\par Freedom House's Freedom On The Net
-(FOTN) report \cite{FOTN}, is an annual ranking and analysis of Internet
-freedom, interpreted from multiple measures including obstacles to access,
-limits on content, and violations of user rights. The FOTN includes a ranking
-of nations by a freedom index, where a higher number indicates a less free
-Internet for that nation. The index is calculated from a series of questions
-that are evaluated according to various sources regarding each nation. These
-questions encompass multiple categories, including government filtering and
-censorship of web traffic. In order to investigate whether the topology of a
-nations ASes is related to Internet freedom, we compare the FOTN freedom index
-with the ease with which a nation can control AS level paths. In figure
-\ref{fig:fotn} we plot the number of ASes required for a nation to control 90%
-of paths with that nation's freedom index for multiple timestamps of AS
-relationships.
-
-\begin{figure*}
-	\centering
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{fotn2012}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{fotn2015}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{fotn2016}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{fotn2017}
-	\end{subfigure}
-
-	\caption{Ratio of Paths Intercepted for a certain number of Border Nodes for multiple nations}\label{fig:fotn}
-\end{figure*}
-
-\par For each date we fit the data with an ordinary least
-squares linear regression, and found the relationship to be statistically
-significant (all p-values are less than 0.05). {\color{blue}need some better way to show
-exact p-values} The relationship shows that more free nations, as determined
-by FOTN scores, tend to have a more difficult to control AS-level topology, as
-in border nodes that control fewer portions of paths in and out of that
-country. This suggests that either the desire of national governments to
-control Internet traffic has impacted the layout of AS relationships, the
-topology of the modern Internet has evolved in a way to enable more successful
-filtering and blocking, or a combination of both. Up to this point it has been
-assumed that Internet topology and Internet censorship are related, but this
-has not been supported by both a measure of Internet freedom and actual AS-
-level paths.
-
-\subsection{Comparing Internet Topology with Empirical Censorship Data}
-
-Another
-technique for relating our chokepoint metric with censorship is to look at
-empirical censorship data. One such dataset is provided by the Open
-Observatory of Network Interference (OONI) \cite{OONI}. OONI is a project
-wherein Internet users around the world use a tool (the ooniprobe) to examine
-evidence of Internet censorship. The ooniprobe allows users to run a series of
-tests, such as an HTTP requests that might be known to be blocked by certain
-nations. The ooniprobe then compares the results of tests with previously
-determined unblocked examples of the same tests. Results where the test
-differs from expected are considered anomalies, and further heuristics can
-determine if the anomalies are likely to be censorship events.
-\par We examined
-18 days of OONI data from January in 2018. We recorded the ratio of anomalies
-in measurements for several classes of anomalous behavior to total
-measurements carried out in each nation. We used the web connectivity test
-from OONI that records the following fields to measure connectivity: DNS,
-HTTP\_FAIL, HTTP\_DIFF, and TCP. Each of these fields are recorded as true or
-false, according to whether the test resulted in a potential censorship event.
-We then plotted the ratio of potential censorship events in the dataset
-against a national metric of chokepoint potential. In this case, we again
-looked at the number of border nodes required to intercept 90\% of AS-level
-paths. The results of this test are plotted in figure \ref{fig:ooni}. A
-relationship like the one seen in the FOTN data is not as clear, perhaps due
-to the uneven sampling across nations. It is clear, however, that nations can
-be classified both by its chokepoint potential and the occurrence of possible
-censorship events. A nation, for instance, with a high ratio of web
-connectivity anomalies with a high chokepoint potential can be considered a
-censorship active nation with a strongly hierarchical AS topology. Thus, this
-analysis serves as a monitoring tool for the state of nations on the Internet
-in regards to actual censorship implementation, and potential for further
-censorship.
-
-\begin{figure*}
-	\centering
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{all_09}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{dns_09}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{http_diff_09}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{http_fail_09}
-	\end{subfigure}
-	\\
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=\linewidth]{tcp_09}
-	\end{subfigure}
-
-	\caption{Ratio of anomalies in OONI measurements compared to chokepoint potential}\label{fig:ooni}
-\end{figure*}
+We are not the first to investigate the relationship between Internet freedom and
+AS-level topology. Similar techniques have been used to classify nations according
+to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
+single moment in time, however, making the results limited in terms of stability and predictability.
+Additionaly, previous work has not used country level chokepoints as the link between Internet topology
+and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
+to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
+to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
+outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
+understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
+The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
+sheds light on the relationship between topology and Internet freedom, but reveals currently free
+nations that could easily implement censorship if their governments decided to.
 
@@
 
-We have introduced an efficient technique for studying the AS-level topology of the Internet. Additionally,
-we introduced the metric of chokepoint potential in order to evaluate the control that border nodes possess
-over paths into and out of a nation. We have applied the routing trees generated by BGPSimPy along with the
-chokepoint potential metric to present an overview of the evolution of the amount and strength of chokepoints
-on the AS graph over the past 6 years. We also connected our analysis of AS-level chokepoints in a novel way to
-both qualitative measures of Internet freedom and empirical censorship data. The former comparison reveals
-a strong relationship between the chokepoint potential of a nation and its lack of Internet freedom, while the
-former provides insight into the categories of nations, in regards to their implementations of censorship and
-the layout of their AS relationships.
+\subsection{Routing Trees Dataset}
+In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
+with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
+takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
+working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
+BGP paths.
 
-\end{multicols}
+% Points of Presence
+
+\subsection{Future Work}
+% OONI data
+While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
+This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
+actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
+the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
+recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
+in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
+censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
 
@@
 
+\end{multicols}
+
 \end{document}

### 2018-06-29 Paper/paper.tex
updated draft

@@
 \begin{document}
-\raggedbottom
 \maketitle
@@
 \par
+The layout of the rest of this paper is as follows: Section 3 provides relavent background information
+to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
+potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
+explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
+respectively.
 
@@
 selected by other ASes. AS topology researchers often use measurement tools to create a picture of the global Internet.
-These measurement tools, for instance BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute \cite{traceroute}, suffer from
+These measurement tools, for instance BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute, suffer from
 incompleteness and innacuracies \cite{tracerouteProblems}.
@@
 for each timestamp recorded. These results are detailed in figure \ref{fig:nodes}.
+The distinct behaviors of each nation are rather striking. We can see a steep decline for some
+nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
+(Germany and Brazil). Other nations have stayed more stable. It is worth pointing out that if a nation
+has required the same number of border ASes to control many paths over several years, this suggests
+that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
+and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
+needed for some nations even more glaring.
 
@@
 
+% Chokepoint Potential
+In this paper we have introduced a novel measure, chokepoint potential, that allows us to
+evaluate the level of difficulty that a government faces when trying to censor the Internet
+based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
+\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
+We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
+design principles so that this tool is readily deployable for future research.
+% Ecolution over time
+\par
+Our application of \toolname to investigate evolutions in the global AS graph over time shows us
+interesting trends in AS topologies.
+% Internet Freedom
+\par
+We have validated chokepoint potential as having a relationship to actual censorship by showing
+the significance of the relationship between the chokepoint potential of nations and their FOTN
+score from freedom house.
+
 \subsection{Routing Trees Dataset}

### 2018-07-02 Paper/paper.tex
proposed new title and revised abstract

@@
 %Titles:
-\title{The Evolution of National AS Chokepoints and Their Connection to
-Internet Freedom}
+%\title{The Evolution of National AS Chokepoints and Their Connection to
+%Internet Freedom}
+\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
 
@@
   \section{Abstract}
-Paths on the Autonomous Systems (AS) graph of the Internet
-derived from the Border Gateway Protocol (BGP) can be used by researchers to
-understand the dynamics of Internet topology and to interpret how that
-topology may enable nations to enact censorship, surveillance, or other
-Internet control measures. Unfortunately datasets of these paths are not
-generally made publicly available, and paths collected from measurements such as
-traceroute or BGP probes tend to be incomplete. Simulation
-frameworks have been used to generate AS paths based on inferred AS relationships.
-We introduce \toolname, a suite of open source, cross platform, and efficient
-tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-control Internet traffic, either through censorship or surveillance.
-Previous research endeavors in this
-direction have only identified chokepoints in single snapshots. We apply our
+
+  Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.
+Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
+
+  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
+We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
+To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
+To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civilian liberty, finding [NEED A RESULT HERE].
+[NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME].
+
+This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
 path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years in order to introduce a new technique to investigate the
-evolution of the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible. As an
-application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-Internet freedom and with our chokepoint potential measure in order to interpret
-the relationship between AS-level topology and actual censorship activity, providing an
-illustration of new ways to monitor which governments can easily control the flow of 
-information in their nations and whether they are acting on that potential.
+over multiple years. We provide comprehensive and accessible tools for studying
+the complex and dynamic AS level Internet topology. Through this approach we
+can more carefully evaluate the state of the Internet than was previously possible.
+
+%% Paths on the Autonomous Systems (AS) graph of the Internet
+%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
+%% understand the dynamics of Internet topology and to interpret how that
+%% topology may enable nations to enact censorship, surveillance, or other
+%% Internet control measures. Unfortunately datasets of these paths are not
+%% generally made publicly available, and paths collected from measurements such as
+%% traceroute or BGP probes tend to be incomplete. Simulation
+%% frameworks have been used to generate AS paths based on inferred AS relationships.
+%% We introduce \toolname, a suite of open source, cross platform, and efficient
+%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
+%% control Internet traffic, either through censorship or surveillance.
+%% Previous research endeavors in this
+%% direction have only identified chokepoints in single snapshots. We apply our
+%% path simulation and routing tree datasets to view snapshots of the Internet
+%% over multiple years in order to introduce a new technique to investigate the
+%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
+%% can more carefully evaluate the state of the Internet than was previously possible. As an
+%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
+%% Internet freedom and with our chokepoint potential measure in order to interpret
+%% the relationship between AS-level topology and actual censorship activity, providing an
+%% illustration of new ways to monitor which governments can easily control the flow of 
+%% information in their nations and whether they are acting on that potential.
 

### 2018-07-05 Paper/paper.tex
worked on introduction

@@
 
+\par Widespread Internet censorship and surveillance pose a
+significant threat to individuals throughout the world. As routine
+tasks, communications, entertainment, and information move on-line and
+are mediated by the Internet, most of us have little choice about
+whether or not to rely on the Internet.  According to the
+International Telecommunication Union (ITU), the number of individual
+Internet users increased from 1.024 million in 2005 to 3.578 million
+in 2017 \cite{itu}. The majority of these users operate in an
+environment that restricts Internet freedom in some way. For example,
+the 2017 Freedom on the Net report from Freedom House reports that
+64\% of Internet users belong to a nation with Internet that is not
+free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
+the EU is considering content rules surrounding copyright, net
+neutrality rules in the U.S. were recently overturned, and large
+content delivery networks are under enormous pressure to 'do
+something' about fake news and bad actors. Taken together, these
+trends will likely restructure the Internet in unforeseen ways,
+as organizations respond to new challenges and realities.
+
+Today, most such control is exercised at the country level,
+as governments have recognized both the threat and the opportunity
+that is posed by ubiquitous on-line communication. For example, much
+of the organization, revolutionary momentum, and broadcasting of
+events during the Arab Spring have been attributed to social media
+communications through the Internet \cite{arabspring}, and similarly several countries
+sought to control these movements by disconnecting in-country networks from the rest of the world (FIXME: need references here).
+
+These two factors, increasing levels of content monitoring and
+Internet dependence, point to the need for improved tools and methods
+for studying worldwide Internet censorship and surveillance over time.
+[A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
+  MISSING.]  In this paper, we focus on AS-level topology and on
+censorship, asking how AS topology has changed over time with respect
+to national boundaries.  We define \emph{chokepoint potential} to
+quantify how a country’s AS topology is organized in terms of paths
+that can carry traffic across international borders.  We developed a
+suite of tools, called \toolname, for studying national chokepoints on
+the AS graph efficiently. To illustrate these ideas and tools, we
+study how chokepoint potential correlates with two independent
+measures of civilian liberty, finding [NEED A RESULT HERE]. [NEED A
+  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME].
+
+The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
+and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
+
+%Whether or not AS-level topology supports
+%these efforts is a research question pivotal to an understanding of
+%the dynamics of Internet censorship and surveillance.
+
+%% \par
+%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
+%% organize revolutionary movements, and connect with other nations' governments and citizens
+%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
+%% of happenings during the Arab Spring can be attributed to social media communications
+%% through the Internet \cite{arabspring}. Because of this potential, national governments may
+%% take interest in maximizing their ability to control the flow of information on the Internet.
+%Censorship and Surveillance are growing phenomena
+
 %The Internet is Dynamic and growing
-The Autonomous Systems layer of the Internet has grown and changed dramatically over its history.
-In the early 2000s, there were around 10,000 ASes connected to the global AS graph. Today there are
-over 60,000 of these ASes. The locations and relationships of these ASes determine how many chokepoints
-of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion of the AS graph suggests a dynamic system with behavior that is important for
-understanding things like path robustness, AS hierarchies, and the potential for governments to control
-information as it flows in and out of their borders. It is certain that structrural qualities of national
-AS graphs have evolved differently from nation to nation, whether that be due to economic decision making, infrastructural
-necessities, or efforts to build a powerful censorship and surveillance network.
+
+We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
+The structural
+properties of national AS graphs have evolved differently
+from nation to nation whether from economic decision making, infrastructural necessities, or
+efforts to build a powerful censorship and surveillance
+network.
+%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
+%The locations and relationships of these ASes determine how many chokepoints
+%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
+This rapid expansion, together with the changing role of national
+governments, points to the importance of understanding properties like
+path robustness, AS hierarchies, and the potential for organizations to
+control information as it flows in and out of their networks.
 
 %Current Events
-\par
-The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-organize revolutionary movements, and connect with other nations' governments and citizens
-to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-of happenings during the Arab Spring can be attributed to social media communications
-through the Internet \cite{arabspring}. Because of this potential, national governments may
-take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-\par
-Internet censorship and surveillance have become profound social and technical issues facing the world. The portion of the
-global population that uses the Internet has increased substantially in recent years. According to the International Telecommunication Union,
-the number of individual users of the Internet has increased from 1.024 million in 2005 to 3.578 million in 2017 \cite{itu}. Of the Internet
-population, a majority of them are subject to some lack of Internet freedom. The 2017 Freedom on the Net report from freedom house reports that
-64\% of Internet users belong to a nation with Internet that is not free or partly free \cite{FOTN}. It is evident that Internet control is
-a powerful force for governments to wield. Whether or not AS-level topology supports these efforts is a research question pivotal to
-an understanding of the dynamics of Internet censorship and surveillance.
 
 %Diverse Censorship Architecture
-\par
-Every nation has a different layout and count of ASes. The censorship and surveillance strategies of nations also differ. For instance,
-China both conducts keyword filtering in border ASes and in internal provincial nodes \cite{chinafiltering}. Another nation that conducts
-extensive Internet censorship, Iran, has been found to route all of its Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely being conducted. The measure of chokepoint potential, defined later in this paper,
-is an effective way to capture the important properties of border ASes related to these capabilities.
+\par Every nation has a different layout and count of ASes. The
+censorship and surveillance strategies of nations also differ. For
+instance, China both conducts keyword filtering in border ASes and in
+internal provincial nodes \cite{chinafiltering}, while Iran routes its
+Internet traffic through a centralized facility \cite{irancensor}.
+Having views into the AS topology of a nation, then, will both help
+researchers identify nations that could easily conduct censorship and
+also provide possible insight into what kind of censorship is likely
+being conducted. The measure of chokepoint potential, defined later in
+this paper, is a succint way to estimate the important properties of
+border ASes related to these capabilities.
 

### 2018-07-06 Paper/paper.tex
Figure generation, workon first part of the apper

@@
 
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{subcaption}
@@
 \usepackage{geometry}
+\usepackage{todonotes}
  \geometry{
@@
 
-\newcommand{\toolname}{{\color{blue}TOOL\_NAME }}
+\newcommand{\toolname}{{\textcolor{blue}TOOL\_NAME }}
+\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
 
@@
 \begin{multicols}{2}
-
-  \section{Abstract}
-
-  Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.
+\abstract{
+  Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
+We should be more specificy about what this statement means, ie with respect to national boundaries}.
 Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
@@
 To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civilian liberty, finding [NEED A RESULT HERE].
-[NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME].
+To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding \todo[inline]{NEED A RESULT HERE}.
+\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
 
@@
 the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
+can more carefully evaluate the state of the Internet than was previously possible.}
 
@@
 significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move on-line and
+tasks, communications, entertainment, and information move online and
 are mediated by the Internet, most of us have little choice about
@@
 the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S. were recently overturned, and large
+neutrality rules in the U.S.\ were recently overturned, and large
 content delivery networks are under enormous pressure to 'do
@@
 as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous on-line communication. For example, much
+that is posed by ubiquitous online communication. For example, much
 of the organization, revolutionary momentum, and broadcasting of
 events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}, and similarly several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world (FIXME: need references here).
+communications through the Internet \cite{arabspring}. Several countries
+sought to control these movements by disconnecting in-country networks from the rest of the world \todo[inline]{need references here}.
 
@@
 for studying worldwide Internet censorship and surveillance over time.
-[A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.]  In this paper, we focus on AS-level topology and on
+\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
+  MISSING.}  In this paper, we focus on AS-level topology and on
 censorship, asking how AS topology has changed over time with respect
@@
 study how chokepoint potential correlates with two independent
-measures of civilian liberty, finding [NEED A RESULT HERE]. [NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME].
+measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
+  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
 
@@
 \par
-The primary contributions of our work are as follows: \textbf{1.} We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic. \textbf{2.} We
-provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations. \textbf{3.} We develop a new tool, \toolname, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. Finally, \textbf{4.}, we show that our chokepoint measure
+The primary contributions of our work are as follows: 
+\begin{enumerate}
+ \item We introduce the measure of chokepoint potential and 
+motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
+ \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
+allowing for a comparision between nations.
+ \item We develop a new tool, \toolname, that provides the capability for
+efficiently evaluating chokepoints for a given state of the Internet. 
+ \item we show that our chokepoint measure
 has a significant relationship to Internet freedom as measured by a qualitative source.
+\end{enumerate}
 

### 2018-07-10 Paper/paper.tex
only partway through background

@@
 \newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
 
@@
 
-The AS-layer of the Internet is the highest level of organization of the Internet. ASes generally represent ISP networks,
-large university networks, or government entities. Interdomain routing between these ASes is governed by BGP, and
-individual ASes may have local preferences for how they wish to send packets along, but they cannot govern the paths
-selected by other ASes. AS topology researchers often use measurement tools to create a picture of the global Internet.
-These measurement tools, for instance BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute, suffer from
-incompleteness and innacuracies \cite{tracerouteProblems}.
+The AS-layer is the highest level of organization of the Internet, with each individual AS consisting of a separately administered network, which could represent a large ISP, a university network, or a government entity.
+The Border Gateway Protocol (BGP) is used to route traffic among ASes.
+\Steph[inline]{need citations}.
+AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
+These tables express local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
+
+Because BGP is a distributed protocol, with all connectivity
+information stored in the local routing tables, it is challenging to
+infer the exact AS topology at any point in time.  There are several
+strategies for inferring the AS graph from local information, and no method is perfect.  
+BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute, suffer from
+incompleteness and innacuracies \cite{tracerouteProblems}.\Steph[inline]{Kirts: Please explain BGPStream and what the problem with it is.}
 \par

### 2018-07-10 Paper/paper.tex
Converted to IEEE format

@@
 %&pdflatex
-\documentclass{article}
+\documentclass[10pt, conference, letterpaper]{IEEEtran}
 
@@
 \usepackage{graphicx}
-\usepackage{multicol}
-\usepackage{geometry}
 \usepackage{todonotes}
- \geometry{
- a4paper,
- total={170mm,257mm},
- left=20mm,
- top=20mm,
- }
-
- \newenvironment{Figure}
-  {\par\medskip\noindent\minipage{\linewidth}}
-  {\endminipage\par\medskip}
 
@@
 \graphicspath{{./images/}}
-\setlength{\parskip}{0pt}
 
@@
 \maketitle
-\begin{multicols}{2}
-\abstract{
-  Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
+\begin{abstract}{.Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
 We should be more specificy about what this statement means, ie with respect to national boundaries}.
@@
 can more carefully evaluate the state of the Internet than was previously possible.}
+\end{abstract}
 
@@
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{chokepoint}
-	\captionof{figure}{Chokepoint potential example. ASes A,B, and C are all border nodes. 
+	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
 						AS D is an internal node. ASes E and F are both external nodes.
 						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{Figure}
+\end{figure}
 
@@
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{single_CN}
-	\captionof{figure}{The evolution of China's in-to-out Chokepoint Potential}\label{fig:ChinaChokePoint}
-\end{Figure}
+	\caption{The evolution of China's in-to-out Chokepoint Potential}\label{fig:ChinaChokePoint}
+\end{figure}
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{single_US}
-	\captionof{figure}{The evolution of the US's Chokepoint Potential}\label{fig:USChokePoint}
-\end{Figure}
+	\caption{The evolution of the US's Chokepoint Potential}\label{fig:USChokePoint}
+\end{figure}
 
@@
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{single_DE}
-	\captionof{figure}{The evolution of Germany's in-to-out Chokepoint Potential}\label{fig:GermChokePoint}
-\end{Figure}
+	\caption{The evolution of Germany's in-to-out Chokepoint Potential}\label{fig:GermChokePoint}
+\end{figure}
 
@@
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{NodesNeeded}
-	\captionof{figure}{Number of ASes to intercept 90\% of ito paths for multiple nations over time}\label{fig:nodes}
-\end{Figure}
+	\caption{Number of ASes to intercept 90\% of ito paths for multiple nations over time}\label{fig:nodes}
+\end{figure}
 
@@
 
-\begin{Figure}
+\begin{figure}
 	\centering
 	\includegraphics[width=\linewidth]{fotn2017}
-	\captionof{figure}{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017)}\label{fig:fotn2017}
-\end{Figure}
+	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017)}\label{fig:fotn2017}
+\end{figure}
 
@@
 
-\end{multicols}
-
 \end{document}

### 2018-07-10 Paper/paper.tex
Fixed an issue with the abstract showing up twice

@@
 \maketitle
-\begin{abstract}{.Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
+
+\begin{abstract}Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
 We should be more specificy about what this statement means, ie with respect to national boundaries}.
@@
 the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.}
+can more carefully evaluate the state of the Internet than was previously possible.
 \end{abstract}
@@
 The AS-layer is the highest level of organization of the Internet, with each individual AS consisting of a separately administered network, which could represent a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes.
-\Steph[inline]{need citations}.
+The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
 AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
@@
 BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute, suffer from
-incompleteness and innacuracies \cite{tracerouteProblems}.\Steph[inline]{Kirts: Please explain BGPStream and what the problem with it is.}
+incompleteness and innacuracies \cite{tracerouteProblems}. BGP dumps and routing table updates
+suffer from the fact that BGP paths stored in routing tables reflect only the current knowledge of
+a single AS about the routes available. Thus one would need to use many sources to get an accurate picture of
+these routes. This approach is not easily scalable for global studies, particularly those looking at multiple time periods.
 \par

### 2018-07-10 Paper/paper.tex
Reinforced argument to use border ASes

@@
 communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \todo[inline]{need references here}.
-
+sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
 These two factors, increasing levels of content monitoring and
@@
 	\centering
+	\includegraphics[width=\linewidth]{bnodes}
+	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\includegraphics[width=\linewidth]{chokepoint}
@@
 ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-We argue that this is intuitive because border
-ASes are the first opportunity for nations to censor incoming traffic and the last opportunity for nations to block access to
-foreign websites from within their borders. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This simplification has
+The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
+of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
+This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
+to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
 the added benefit of making calculations more efficient.

### 2018-07-10 Paper/paper.tex
Added draft tool name

@@
 
-\newcommand{\toolname}{{\textcolor{blue}TOOL\_NAME }}
+\newcommand{\toolname}{BGP-SAS }
 \newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
@@
 
-\begin{abstract}Internet topology has been measured and modeled for many years, but few studies  quantify how the topology changes over time.\Ben[inline]{This is not true, there are many studies of the evolution of the AS Internet.
-We should be more specificy about what this statement means, ie with respect to national boundaries}.
+\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
+Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
 Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
@@
 To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding \todo[inline]{NEED A RESULT HERE}.
+To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
 \todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
@@
 % Preprocessing and Data pipeline
-\toolname is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
+Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
 \toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}

### 2018-07-13 Paper/paper.tex
Copied PoPs discussion over, added argument for AS-level with no geography

@@
 % Points of Presence
+\subsection{Other Chokepoint Measures}
+There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
+limiting the number of physical connections that cross the border, or working with---and supporting---a s,a;; set pf ;arge ISPs that have international connections.
+Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
+We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
+virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
+the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
+of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
+infrastructue on a case-by-case basis.
+Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
+to show time evolution of chokepoints.
 

### 2018-07-13 Paper/paper.tex
tiny bit of editing

@@
 
-The AS-layer is the highest level of organization of the Internet, with each individual AS consisting of a separately administered network, which could represent a large ISP, a university network, or a government entity.
+The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
 The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
 AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
-These tables express local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
+These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
 
@@
 information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  There are several
-strategies for inferring the AS graph from local information, and no method is perfect.  
-BGP Updates from BGPStream \cite{BGPStream} or empirical paths from traceroute, suffer from
-incompleteness and innacuracies \cite{tracerouteProblems}. BGP dumps and routing table updates
-suffer from the fact that BGP paths stored in routing tables reflect only the current knowledge of
-a single AS about the routes available. Thus one would need to use many sources to get an accurate picture of
-these routes. This approach is not easily scalable for global studies, particularly those looking at multiple time periods.
-\par
-AS-level studies are further complicated by the lack of ground truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As part of the CAIDA project, this inference technique was extended, leading
-to the CAIDA AS-relationship dataset \cite{CAIDApaper}, \cite{CAIDA}. We choose these relationships for our purposes, and they are the current
-research standard.
-\par
-Finding BGP paths is more of a challenge. Packet based simulations, wherein BGP is simulated directly, are computationally infeasible for the
-scales relevant to a global study. An accurate, but realistic, simulation technique must be chosen to provide useful research potential in this regard.
-The BGPSim algorithm \cite{quicksand} is one such simulation technique that is suitable for this study's purposes. BGPSim takes a set of AS relationships
-as input, such as those provided by the CAIDA dataset \cite{CAIDA}, and returns routing trees based off of these relationships. These paths are found via
-a modified breadth-first seach (BFS) algorithm. The BFS adds edges to routing trees first according to local preference (LP), then shortest path (SP), and finally tiebreak (TB).
-A resulting routing tree contains all the equally reasonable paths (according to economic concerns) that exist between source ASes (within the routing tree) to 
-destination ASes (the root of the routing tree). We find this technique suitable for our purposes, so we extend BGPSim 
-(in ways explained further in this paper) as part of \toolname.
+infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
+on routing table information and updates
+updates suffer from the fact that BGP paths stored in routing tables
+reflect only the current knowledge of a single AS about the routes
+available~\cite{BGPStream}.
+% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
+Thus one would need to use many sources to infer an accurate
+picture of the BGP topology, which is not easily scalable for
+global studies, particularly those looking at multiple time periods.
+Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
+
+\par AS-level studies are further complicated by the lack of ground
+truth data for AS relationships and BGP paths. For AS relationships,
+inference is often used based on economic considerations to classify
+the relationships of ASes. This was first done by Gao \cite{gao} by
+maximizing the occurances of certain economic rules on the AS graph by
+choosing a particular set or relationships. This technique was only
+evaluated against a single ISPs set of true relationships, however. As
+part of the CAIDA project, this inference technique was extended,
+leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
+\cite{CAIDA}. We choose these relationships for our purposes, and they
+are the current research standard.
+
+\par Finding BGP paths is more of
+a challenge. Packet based simulations, wherein BGP is simulated
+directly, are computationally infeasible for the scales relevant to a
+global study. An accurate, but realistic, simulation technique must be
+chosen to provide useful research potential in this regard.  The
+BGPSim algorithm \cite{quicksand} is one such simulation technique
+that is suitable for this study's purposes. BGPSim takes a set of AS
+relationships as input, such as those provided by the CAIDA dataset
+\cite{CAIDA}, and returns routing trees based off of these
+relationships. These paths are found via a modified breadth-first
+seach (BFS) algorithm. The BFS adds edges to routing trees first
+according to local preference (LP), then shortest path (SP), and
+finally tiebreak (TB).  A resulting routing tree contains all the
+equally reasonable paths (according to economic concerns) that exist
+between source ASes (within the routing tree) to destination ASes (the
+root of the routing tree). We find this technique suitable for our
+purposes, so we extend BGPSim (in ways explained further in this
+paper) as part of \toolname.
 

### 2018-07-13 Paper/paper.tex
Updated figures in results section, added explanation for country choices

--- 2018-07-13 Paper/paper.tex
@@
 \newcommand{\toolname}{BGP-SAS }
+\newcommand{\timerange}{(2012-2018)}
+
 \newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
@@
 % Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some questions remain unanswered however. For instance, to what
+There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
 extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
@@
 With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-
+\par
 % Chokepoints and Internet Freedom
@@
 	\centering
-	\includegraphics[width=\linewidth]{single_CN}
-	\caption{The evolution of China's in-to-out Chokepoint Potential}\label{fig:ChinaChokePoint}
+	\includegraphics[width=\linewidth]{NodesFree}
+	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
 \end{figure}
-
 \begin{figure}
 	\centering
-	\includegraphics[width=\linewidth]{single_US}
-	\caption{The evolution of the US's Chokepoint Potential}\label{fig:USChokePoint}
+	\includegraphics[width=\linewidth]{NodesCensor}
+	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
 \end{figure}
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{multi}
+	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
+	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
+\end{figure*}
 
+We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
+extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
+censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
+to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
+The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
+of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
+nations is considered highly open.
+
+\par
 To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
@@
 \par
-For instance, consider figures \ref{fig:ChinaChokePoint} and \ref{fig:USChokePoint}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily)
+For instance, consider \figurename \ref{fig:multiTime}. 
+The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
 is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
@@
 \par
-The United States and China are shown here to exhibit their dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
+The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
 it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. China clearly has developed an AS topology such that very few
+United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
 ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{single_DE}
-	\caption{The evolution of Germany's in-to-out Chokepoint Potential}\label{fig:GermChokePoint}
-\end{figure}
-
 \par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in figure \ref{fig:GermChokePoint}. For Germany, a fairly constant
+Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
 trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesNeeded}
-	\caption{Number of ASes to intercept 90\% of ito paths for multiple nations over time}\label{fig:nodes}
-\end{figure}
 
@@
 In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in figure \ref{fig:nodes}.
+for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
 The distinct behaviors of each nation are rather striking. We can see a steep decline for some
 nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and Brazil). Other nations have stayed more stable. It is worth pointing out that if a nation
+(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
 has required the same number of border ASes to control many paths over several years, this suggests

### 2018-07-13 Paper/paper.tex
Jed edited a bit in discussion and a few other small places

@@
 \cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering is controlled by non-central provincial ASes. China has a
+some filtering at the time of the study was controlled by non-central provincial
+ASes. China had a
 diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network. The potential for various forms of censorship in
+Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
 regards to various AS level topologies motivates the question: Is centralized
@@
 There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, or working with---and supporting---a s,a;; set pf ;arge ISPs that have international connections.
+limiting the number of physical connections that cross the border, centralizing
+the DNS infrastructure within the country, or working
+with---and supporting---a set of ISPs that have international connections.
 Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
@@
 to show time evolution of chokepoints.
+
+The DNS infrastructure is another potential graph where chokepoints can be
+exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
+higher layer in the network stack than IP routes.  However, we leave analysis of
+this graph for future work because it is outside the scope of \toolname and
+because the countries that rely on DNS as their main form of censorship are
+fewer in number and are among the least sophisticated in terms of their
+censorship regime.
 

### 2018-07-16 Paper/paper.tex
Updated FOTN figure

@@
 content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Taken together, these
+something' about fake news and bad actors. Also, other forms of Internet
+tampering such as code injection~\cite{badtraffic,191996} have been observed within the
+Internet's backbone. Taken together, these
 trends will likely restructure the Internet in unforeseen ways,
@@
 	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017)}\label{fig:fotn2017}
+	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
+	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
 \end{figure}

### 2018-07-24 Paper/paper.tex
Expanded definition of chokepoint potential and did some smaller edits

@@
 
-\newcommand{\toolname}{BGP-SAS }
+%\newcommand{\toolname}{BGP-SAS}
+%\newcommand{\toolname}{BGP Path Finder}
+\newcommand{\toolname}{BGP Path Guide}
 \newcommand{\timerange}{(2012-2018)}
@@
   In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
+We introduce a new measure, \emph{national chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
+To study country-level chokepoints, we developed \toolname{}, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
+To illustrate these ideas and tools, we studied how national chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
 \todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
@@
 International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 million in 2005 to 3.578 million
+Internet users increased from 1.024 billion in 2005 to 3.578 billion
 in 2017 \cite{itu}. The majority of these users operate in an
@@
 censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{chokepoint potential} to
+to national boundaries.  We define \emph{national chokepoint potential} to
 quantify how a country’s AS topology is organized in terms of paths
 that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname, for studying national chokepoints on
+suite of tools, called \toolname{}, for studying national chokepoints on
 the AS graph efficiently. To illustrate these ideas and tools, we
-study how chokepoint potential correlates with two independent
+study how national chokepoint potential correlates with two independent
 measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
@@
 
-The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
+The paper extends earlier research on AS topology, first by introducing the national chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
 and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
@@
 
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
+We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 ASes in the early 2000s to over 60,000 today.
 The structural
@@
 instance, China both conducts keyword filtering in border ASes and in
-internal provincial nodes \cite{chinafiltering}, while Iran routes its
+internal provincial ASes \cite{chinafiltering}, while Iran routes its
 Internet traffic through a centralized facility \cite{irancensor}.
@@
 \begin{enumerate}
- \item We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations.
- \item We develop a new tool, \toolname, that provides the capability for
+ \item We introduce the measure of chokepoint potential for single border ASes and the aggregate measure national chokepoint potential, 
+motivating their value as a way to interpret the capability for a nation to enact censorship or surveillance of its Internet traffic.
+ \item We provide an overview of the evolution of national AS-level chokepoints over time, showing how the evolution of the Internet has either facillitated stronger
+ AS-level chokepoints or lead to more open flow of information across borders.
+ \item We develop a new tool, \toolname{}, that provides the capability for
 efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that our chokepoint measure
-has a significant relationship to Internet freedom as measured by a qualitative source.
+ \item we show that national chokepoint potential
+has a significant relationship to Internet freedom as measured by two qualitative sources.
 \end{enumerate}
@@
 to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
+potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname{}; Section 6
 explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
@@
 The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
+AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP AS.
 These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
@@
 infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and updates
+on routing table information and BGP
 updates suffer from the fact that BGP paths stored in routing tables
@@
 purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname.
+paper) as part of \toolname{}.
 
@@
 	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
-						AS D is an internal node. ASes E and F are both external nodes.
+	\caption{Chokepoint potential example. ASes A,B, and C are all border ASes. 
+						AS D is an internal AS. ASes E and F are both external ASes.
 						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
@@
 
-% Why Border ASes?
-In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-the added benefit of making calculations more efficient.
+Border ASes are ASes that lie one hop from an AS registered to a different country. It stands to reason that if information bottlenecks
+were to be implemented by national governments, they would likely be implemented on border ASes. This is not always true, but there are more reasons
+to use border ASes as our unit of investigation. For instance, the ratio of the number of border ASes of
+a country to all ASes in that country hints toward the abillity of that nation to intercept information flows across their borders.
+\figurename \ref{fig:bnodes} shows that globally the number of border ASes has grown at a slower rate to the number of non-border ASes
+over time so that the ratio of border ASes to ASes has almost continuously decreased.
+On the surface it would seem that this fact reflects an ever increasing capability for nations to intercept AS-level paths.
+This is not
+neccessarily the case, however. Because AS-level paths are generated by the distributed BGP and the local preferences of AS administrators often do not
+lead to simple shortest path selections, a better way to understand the potential for bottlenecks on border ASes is to investigate the number of BGP generated
+paths that pass through each AS. To do this we propose an intuitive measure on border ASes called \emph{chokepoint potential}. Since this is a measure for single ASes
+we also need an aggregate measure for comparing nations, which we call \emph{national chokepoint potential}.
 
-% What is our measure, why does it make sense?
+% % Why Border ASes?
+% In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
+% ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
+% The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
+% of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
+% This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
+% to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
+% the added benefit of making calculations more efficient.
+
 \par
-We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
-to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
-chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
+For our measures we use the term \textit{path} to indicate a path that is generated by BGP consisting of a source AS, a destination AS, and the intermediate
+ASes between them.
+We define a \textit{routing tree} to be a tree rooted at a single destination AS and composed of every acceptable BGP path from the various source ASes that can reach the destination AS.
+We define the chokepoint potential of a border AS to be the ratio of paths intercepted by that border AS
+to the number of paths intercepted by all border ASes belonging to the same nation. For instance, take a border AS $b$, belonging to country $c$.
+If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border ASes within path $p$, then the
+chokepoint potential of $b$, $CP(b)$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
+the relative strength exhibited by a border AS in regards to what portion of paths it intercepts. This is calculated
+seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths. The out-to-in chokepoint potential for a simple dummy nations
+is included in \figurename \ref{fig:chokepoint}.
 
@@
 \begin{equation}\label{eqn:chokePointPotential}
-CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
+CP(b) = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
 \end{equation}
@@
 \par
-Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
-the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
-to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
-censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
+Given a set of routing trees, chokepoint potential is an intuitive way to compare individual ASes. With this measure,
+the sum of the chokepoint potential for all border ASes for a country is 1.0, as in, all of the border ASes collectively control the
+flow of information over the nation's border. To compare one nation to another, we can inspect how many border ASe minimum are required
+to obtain a certain chokepoint potential. The more border ASes required, the more difficult it would likely be for that nation to perform
+censorship or surveillance, as the government would have to restrict more ISPs, place more filtering hardware, etc. Thus, we can define
+another measure, with this one being a national measure. We define the \emph{national chokepoint potential} of a nation to be the minimum number
+of ASes required to intercept a selected percentage of AS-level paths. Previous work has used 90\% path interception as a benchmark indicating strong control
+of information \cite{throats}, so we use this percentage as well. When used here, however, 90\% path interception indicates a single nation intercepts 90\% of the paths
+that either enter said nation (for the out-to-in case) or exit said nation (for the in-to-out case). National chokepoint potential is calculated taking the strongest ASes first,
+such that the measure reflects the smallest number of border ASes needed to intercept the paths. We use the number of ASes here and not a percentage because using a percentage would
+make it difficult to compare nations. Take for instance the United States, which has many more ASes than most nations. If the United States and another nation both required the same percentage
+of border ASes to intercept 90\% of paths, the similarity of these nations would be misleading, because the United States would actually need to control many more ASes.
+Our technique provides a comparison more likely to reflect reality.
 
@@
 % Preprocessing and Data pipeline
-Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
+\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
+\toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
 to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}.
+are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national chokepoint potential 
+for each nation. In our experiments, we used the country codes returned from Team Cymru's
+IP to ASN whois service \cite{cymru}. This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
+we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We chose this route
+because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
 
@@
 we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
+traversed. For each AS visited, the number of paths intercepted by that AS is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
 necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
+\toolname{} takes the resulting chokepoint potentials and generates national chokepoint potentials for each nation.
 
@@
 change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border node per snapshot. As a result we can investigate
+For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border AS per snapshot. As a result we can investigate
 how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
@@
 the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free).
- To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
-Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
-looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
-with each nation's (of those recorded by Freedom House) freedom on the net score.
+to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national chokepoint potential, particulary the number of border ASes required to intercept
+90\% of paths for each nation and viewing these rankings against the FOTN scores.
 
@@
 \par
-To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
+To inspect the changes over time for each nation we arrange all the border ASes belonging to a nation into a list. The list of border ASes is reverse sorted
+so that the first AS has the highest chokepoint potential. The sum of all these chokepoint potentials is 1.0. Then we step through the list, and record
 the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
@@
 The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
+is the number of border ASes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
 farther in the past and more blue lines being more recent.
@@
 
-In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
+In our test of the relationship between Internet Freedom and national chokepoint potential, we plotted the Freedom on the Net score
 of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
@@
 of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
+few border ASes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
 is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
@@
 based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
+\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
 We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
@@
 \par
-Our application of \toolname to investigate evolutions in the global AS graph over time shows us
+Our application of \toolname{} to investigate evolutions in the global AS graph over time shows us
 interesting trends in AS topologies.
@@
 higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname and
+this graph for future work because it is outside the scope of \toolname{} and
 because the countries that rely on DNS as their main form of censorship are

### 2018-07-25 Paper/paper.tex
Various Changes, added FOTP

@@
 
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
+% \begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
 Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
@@
 
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
+%   In this paper, we ask how AS topology has changed over time with respect to national boundaries.
 We introduce a new measure, \emph{national chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
@@
 
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
+% This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
+% path simulation and routing tree datasets to view snapshots of the Internet
+% over multiple years. We provide comprehensive and accessible tools for studying
+% the complex and dynamic AS level Internet topology. Through this approach we
+% can more carefully evaluate the state of the Internet than was previously possible.
+% \end{abstract}
+
+\begin{abstract}
+	Internet topology has been measured and modeled for many years, and this topology has been linked to 
+	censorship and surveillance. Past studies have not, however, quantified the potential for individual 
+	nations to intercept national level Internet paths or analyzed how the evolution of the Internet 
+	topology impacts censorship and surveillance. We introduce a new measure, \emph{national breakthrough
+	potential}, or NBP, to characterize how a country's AS topology is organized in terms of paths that can carry 
+	traffic across international borders.  To study country-level chokepoints, we developed \toolname{}, a 
+	suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the 
+	AS graph. 
+	We find that for many nations the continuous growth of the AS graph is not reflected by a more open
+	flow of information across national borders, and in fact for nations known to conduct censorship we
+	identify a trend toward strengthening chokepoints. Additionally, we find that a significant 
+	relationship exists between NBP and Internet freedom as evaluated from two qualitative sources.
 \end{abstract}
@@
 neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
+content services are under enormous pressure to 'do
 something' about fake news and bad actors. Also, other forms of Internet
@@
 \todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{national chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
+  MISSING.}
+
+\par
+In this paper, we focus on the Autonomous Systems (AS) level of the Internet.
+Autonomous Systems are groups of routers under common management, such as the network of
+a University or an Internet Service Provider (ISP). We apply AS-level topologies to
+censorship and surveillance by evaluating border ASes, or ASes that can connect directly to at least
+one AS from another nation according to a new measure \emph{chokepoint potential}.
+With this measure we can ask how AS topology has changed over time with respect
+to national boundaries.  We define \emph{national breakthrough potential} to
+measure how a country's AS topology is organized in terms of paths
 that can carry traffic across international borders. We developed a
@@
 study how national chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the national chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
+measures of civilian liberty, finding that a significan relationship exists between NBP and both liberty measures.
+The paper extends earlier research on AS topology, first by introducing the NBP measure, 
+then by presenting open-source cross-platform tools for simulating Border Gateway Protocol (BGP) paths,
+determining breakthrough potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
 
@@
 
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 ASes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
+The change in national AS topologies is of interest to censorship research.
+The total size of the AS graph (The global network of all ASes) has grown from about 10,000 ASes in the early 2000s to over 60,000 today. 
+The structural properties of national subgraphs have evolved differently
+from nation to nation whether from economic decision making, infrastructural necessities, or efforts to build a powerful censorship and surveillance
 network.
@@
 %Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
+\par Every nation has a different layout and number of ASes, and these national networks
+connect to foreign ASes in unique ways. The
 censorship and surveillance strategies of nations also differ. For
@@
 Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
+A picture of a nation's censorship or surveillance capabillities then
+is not complete without a topological view point.
+Having the capabillity to investigate the AS topology of a nation, then, will both help
 researchers identify nations that could easily conduct censorship and
 also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
+being conducted. Our measure of border ASes, chokepoint potential, defined in detial
+in section 3, is a succint way to estimate the important properties of
+border ASes related to these capabilities, and our aggregate extension of this measure to
+the national level, national breakthrough potential, is an effective way to evaluate and
+compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.
 
@@
 \begin{enumerate}
- \item We introduce the measure of chokepoint potential for single border ASes and the aggregate measure national chokepoint potential, 
+ \item We introduce the measure of chokepoint potential for single border ASes and the aggregate measure national breakthrough potential, 
 motivating their value as a way to interpret the capability for a nation to enact censorship or surveillance of its Internet traffic.
@@
 efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that national chokepoint potential
+ \item We show that national breakthrough potential
 has a significant relationship to Internet freedom as measured by two qualitative sources.
@@
 \par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname{}; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
+The layout of the rest of this paper is as follows: Section 2 provides relavent background information
+to the problems we are investigating; Section 3 introduces and defines the measures of chokepoint
+potential and national breakthrough potential; Section 4 details our tool for simulating and evaluating AS topologies \toolname{}; Section 5
+explains our experimental setups; Sections 6, 7, and 8 provide results, discussion, and related work
 respectively.
@@
 The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP AS.
+AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each AS.
 These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
@@
 
-\section{Chokepoint Potential}
+\section{Chokepoint Measures}
 
@@
 paths that pass through each AS. To do this we propose an intuitive measure on border ASes called \emph{chokepoint potential}. Since this is a measure for single ASes
-we also need an aggregate measure for comparing nations, which we call \emph{national chokepoint potential}.
+we also need an aggregate measure for comparing nations, which we call \emph{national breakthrough potential}.
 
@@
 censorship or surveillance, as the government would have to restrict more ISPs, place more filtering hardware, etc. Thus, we can define
-another measure, with this one being a national measure. We define the \emph{national chokepoint potential} of a nation to be the minimum number
+another measure, with this one being a national measure. We define the \emph{national breakthrough potential} of a nation to be the minimum number
 of ASes required to intercept a selected percentage of AS-level paths. Previous work has used 90\% path interception as a benchmark indicating strong control
 of information \cite{throats}, so we use this percentage as well. When used here, however, 90\% path interception indicates a single nation intercepts 90\% of the paths
-that either enter said nation (for the out-to-in case) or exit said nation (for the in-to-out case). National chokepoint potential is calculated taking the strongest ASes first,
+that either enter said nation (for the out-to-in case) or exit said nation (for the in-to-out case). National breakthrough potential is calculated taking the strongest ASes first,
 such that the measure reflects the smallest number of border ASes needed to intercept the paths. We use the number of ASes here and not a percentage because using a percentage would
@@
 Our technique provides a comparison more likely to reflect reality.
+This measure is called a breakthrough potential because it indicates the number of border ASes that intercept a vast majority of paths, meaning that a country with high NBP would require controlling many border ASes to prevent a breakthrough, or an unfettered access of
+internal ASes to that country.
 
@@
 % Preprocessing and Data pipeline
-\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
+\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships as well as the national breakthrough potential
+for all nations represented by the AS graph.
 \toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
 to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national chokepoint potential 
+are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national breakthrough potential 
 for each nation. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}. This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
-we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We chose this route
-because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
+IP to ASN whois service \cite{cymru} to determine which ASes were registered to which nation. 
+This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
+we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We 
+chose this route because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
+Fortunately, ASes with registration changes after each test date tend to not be in the AS graph for that test.
 
@@
 necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-\toolname{} takes the resulting chokepoint potentials and generates national chokepoint potentials for each nation.
+\toolname{} takes the resulting chokepoint potentials and generates national breakthrough potentials for each nation.
 
@@
 the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national chokepoint potential, particulary the number of border ASes required to intercept
+to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national breakthrough potential, particulary the number of border ASes required to intercept
 90\% of paths for each nation and viewing these rankings against the FOTN scores.
@@
 The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
+nations known to limit Internet freedom (China and Turkey) as well as the increase for other nations
 (Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
@@
 
-In our test of the relationship between Internet Freedom and national chokepoint potential, we plotted the Freedom on the Net score
+In our test of the relationship between Internet Freedom and national breakthrough potential, we plotted the Freedom on the Net score
 of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
 we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
+with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn}
 
@@
 	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
+	\includegraphics[width=\linewidth]{fotn}
+	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FOTN score
+	of those nations in 2017. A higher FOTN score indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotn}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{fotp}
+	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FPI
+	of those nations in 2017. A higher FPI indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotp}
 \end{figure}
@@
 
-\section{Discussion}
+\section{Conclusions and Discussion}
 
 % Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
+In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
+and national breakthrough potential, on entire nations, that together allow us to
 evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
+based entirely on AS-level topology. Our technique for evaluating AS-level topology using
 \toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
@@
 \par
-Our application of \toolname{} to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
+Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
+interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
+more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
+we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
 % Internet Freedom
 \par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
+We have validated national breakthrough potential as having a relationship to actual censorship by showing
+the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
+by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
+at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.
 

### 2018-07-26 Paper/background.tex
Break out text files, create gitignore

--- 2018-07-16 Paper/paper.tex
@@
-%&pdflatex
-\documentclass[10pt, conference, letterpaper]{IEEEtran}
-
-\usepackage{xcolor}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{todonotes}
-
-\newcommand{\toolname}{BGP-SAS }
-\newcommand{\timerange}{(2012-2018)}
-
-\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
-
-%Titles:
-%\title{The Evolution of National AS Chokepoints and Their Connection to
-%Internet Freedom}
-\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
-
-\graphicspath{{./images/}}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
-\section{Introduction}
-
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move online and
-are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
-International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 million in 2005 to 3.578 million
-in 2017 \cite{itu}. The majority of these users operate in an
-environment that restricts Internet freedom in some way. For example,
-the 2017 Freedom on the Net report from Freedom House reports that
-64\% of Internet users belong to a nation with Internet that is not
-free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
-the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Also, other forms of Internet
-tampering such as code injection~\cite{badtraffic,191996} have been observed within the
-Internet's backbone. Taken together, these
-trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
-
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
-that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname, for studying national chokepoints on
-the AS graph efficiently. To illustrate these ideas and tools, we
-study how chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-
-%Whether or not AS-level topology supports
-%these efforts is a research question pivotal to an understanding of
-%the dynamics of Internet censorship and surveillance.
-
-%% \par
-%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-%% organize revolutionary movements, and connect with other nations' governments and citizens
-%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-%% of happenings during the Arab Spring can be attributed to social media communications
-%% through the Internet \cite{arabspring}. Because of this potential, national governments may
-%% take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-
-%The Internet is Dynamic and growing
-
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
-%Current Events
-
-%Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial nodes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
-
-% Major Contributions
-\par
-The primary contributions of our work are as follows: 
-\begin{enumerate}
- \item We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations.
- \item We develop a new tool, \toolname, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that our chokepoint measure
-has a significant relationship to Internet freedom as measured by a qualitative source.
-\end{enumerate}
-
-% Rest of Paper Layout
-\par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
-respectively.
-
 \section{Background}
@@
 The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
+AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each AS.
 These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
@@
 purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname.
-
-\section{Chokepoint Potential}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{bnodes}
-	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
-						AS D is an internal node. ASes E and F are both external nodes.
-						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{figure}
-
-% Why Border ASes?
-In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-the added benefit of making calculations more efficient.
-
-% What is our measure, why does it make sense?
-\par
-We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
-to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
-chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
-
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-\end{equation}
-
-\par
-Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
-the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
-to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
-censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-
-\section{Experimental Setup}
-
-% Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border node per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
-% Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free).
- To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
-Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
-looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
-with each nation's (of those recorded by Freedom House) freedom on the net score.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
-% Internet Freedom
-\par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
-
-\subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
-
-% Points of Presence
-\subsection{Other Chokepoint Measures}
-There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, centralizing
-the DNS infrastructure within the country, or working
-with---and supporting---a set of ISPs that have international connections.
-Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
-We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
-virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
-the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
-of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
-Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
-to show time evolution of chokepoints.
-
-The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
-higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname and
-because the countries that rely on DNS as their main form of censorship are
-fewer in number and are among the least sophisticated in terms of their
-censorship regime.
-
-\subsection{Future Work}
-% OONI data
-While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
-This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
-actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
-recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
-in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
-censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}
+paper) as part of \toolname{}.

### 2018-07-26 Paper/chokepoint.tex
Break out text files, create gitignore

--- 2018-07-24 Paper/paper.tex
@@
-%&pdflatex
-\documentclass[10pt, conference, letterpaper]{IEEEtran}
-
-\usepackage{xcolor}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{todonotes}
-
-%\newcommand{\toolname}{BGP-SAS}
-%\newcommand{\toolname}{BGP Path Finder}
-\newcommand{\toolname}{BGP Path Guide}
-\newcommand{\timerange}{(2012-2018)}
-
-\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
-
-%Titles:
-%\title{The Evolution of National AS Chokepoints and Their Connection to
-%Internet Freedom}
-\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
-
-\graphicspath{{./images/}}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{national chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname{}, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how national chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
-\section{Introduction}
-
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move online and
-are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
-International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 billion in 2005 to 3.578 billion
-in 2017 \cite{itu}. The majority of these users operate in an
-environment that restricts Internet freedom in some way. For example,
-the 2017 Freedom on the Net report from Freedom House reports that
-64\% of Internet users belong to a nation with Internet that is not
-free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
-the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Also, other forms of Internet
-tampering such as code injection~\cite{badtraffic,191996} have been observed within the
-Internet's backbone. Taken together, these
-trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
-
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{national chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
-that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname{}, for studying national chokepoints on
-the AS graph efficiently. To illustrate these ideas and tools, we
-study how national chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the national chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-
-%Whether or not AS-level topology supports
-%these efforts is a research question pivotal to an understanding of
-%the dynamics of Internet censorship and surveillance.
-
-%% \par
-%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-%% organize revolutionary movements, and connect with other nations' governments and citizens
-%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-%% of happenings during the Arab Spring can be attributed to social media communications
-%% through the Internet \cite{arabspring}. Because of this potential, national governments may
-%% take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-
-%The Internet is Dynamic and growing
-
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 ASes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
-%Current Events
-
-%Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial ASes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
-
-% Major Contributions
-\par
-The primary contributions of our work are as follows: 
-\begin{enumerate}
- \item We introduce the measure of chokepoint potential for single border ASes and the aggregate measure national chokepoint potential, 
-motivating their value as a way to interpret the capability for a nation to enact censorship or surveillance of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing how the evolution of the Internet has either facillitated stronger
- AS-level chokepoints or lead to more open flow of information across borders.
- \item We develop a new tool, \toolname{}, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that national chokepoint potential
-has a significant relationship to Internet freedom as measured by two qualitative sources.
-\end{enumerate}
-
-% Rest of Paper Layout
-\par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname{}; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
-respectively.
-
-\section{Background}
-
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP AS.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
-
-Because BGP is a distributed protocol, with all connectivity
-information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and BGP
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
-% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
-
-\par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
-leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
-
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
-according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname{}.
-
-\section{Chokepoint Potential}
+\section{Chokepoint Measures}
 
@@
 Our technique provides a comparison more likely to reflect reality.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national chokepoint potential 
-for each nation. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}. This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
-we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We chose this route
-because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each AS visited, the number of paths intercepted by that AS is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-\toolname{} takes the resulting chokepoint potentials and generates national chokepoint potentials for each nation.
-
-\section{Experimental Setup}
-
-% Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border AS per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
-% Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national chokepoint potential, particulary the number of border ASes required to intercept
-90\% of paths for each nation and viewing these rankings against the FOTN scores.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border ASes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these chokepoint potentials is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border ASes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and national chokepoint potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few border ASes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname{} to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
-% Internet Freedom
-\par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
-
-\subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
-
-% Points of Presence
-\subsection{Other Chokepoint Measures}
-There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, centralizing
-the DNS infrastructure within the country, or working
-with---and supporting---a set of ISPs that have international connections.
-Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
-We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
-virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
-the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
-of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
-Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
-to show time evolution of chokepoints.
-
-The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
-higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname{} and
-because the countries that rely on DNS as their main form of censorship are
-fewer in number and are among the least sophisticated in terms of their
-censorship regime.
-
-\subsection{Future Work}
-% OONI data
-While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
-This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
-actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
-recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
-in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
-censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}
+This measure is called a breakthrough potential because it indicates the number of border ASes that intercept a vast majority of paths, meaning that a country with high NBP would require controlling many border ASes to prevent a breakthrough, or an unfettered access of
+internal ASes to that country.

### 2018-07-26 Paper/conclusion.tex
Break out text files, create gitignore

--- 2018-07-16 Paper/paper.tex
@@
-%&pdflatex
-\documentclass[10pt, conference, letterpaper]{IEEEtran}
-
-\usepackage{xcolor}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{todonotes}
-
-\newcommand{\toolname}{BGP-SAS }
-\newcommand{\timerange}{(2012-2018)}
-
-\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
-
-%Titles:
-%\title{The Evolution of National AS Chokepoints and Their Connection to
-%Internet Freedom}
-\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
-
-\graphicspath{{./images/}}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
-\section{Introduction}
-
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move online and
-are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
-International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 million in 2005 to 3.578 million
-in 2017 \cite{itu}. The majority of these users operate in an
-environment that restricts Internet freedom in some way. For example,
-the 2017 Freedom on the Net report from Freedom House reports that
-64\% of Internet users belong to a nation with Internet that is not
-free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
-the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Also, other forms of Internet
-tampering such as code injection~\cite{badtraffic,191996} have been observed within the
-Internet's backbone. Taken together, these
-trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
-
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
-that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname, for studying national chokepoints on
-the AS graph efficiently. To illustrate these ideas and tools, we
-study how chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-
-%Whether or not AS-level topology supports
-%these efforts is a research question pivotal to an understanding of
-%the dynamics of Internet censorship and surveillance.
-
-%% \par
-%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-%% organize revolutionary movements, and connect with other nations' governments and citizens
-%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-%% of happenings during the Arab Spring can be attributed to social media communications
-%% through the Internet \cite{arabspring}. Because of this potential, national governments may
-%% take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-
-%The Internet is Dynamic and growing
-
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
-%Current Events
-
-%Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial nodes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
-
-% Major Contributions
-\par
-The primary contributions of our work are as follows: 
-\begin{enumerate}
- \item We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations.
- \item We develop a new tool, \toolname, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that our chokepoint measure
-has a significant relationship to Internet freedom as measured by a qualitative source.
-\end{enumerate}
-
-% Rest of Paper Layout
-\par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
-respectively.
-
-\section{Background}
-
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
-
-Because BGP is a distributed protocol, with all connectivity
-information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and updates
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
-% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
-
-\par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
-leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
-
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
-according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname.
-
-\section{Chokepoint Potential}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{bnodes}
-	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
-						AS D is an internal node. ASes E and F are both external nodes.
-						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{figure}
-
-% Why Border ASes?
-In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-the added benefit of making calculations more efficient.
-
-% What is our measure, why does it make sense?
-\par
-We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
-to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
-chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
-
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-\end{equation}
-
-\par
-Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
-the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
-to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
-censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-
-\section{Experimental Setup}
-
-% Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border node per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
-% Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free).
- To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
-Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
-looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
-with each nation's (of those recorded by Freedom House) freedom on the net score.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Discussion}
+\section{Conclusions and Discussion}
 
 % Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
+In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
+and national breakthrough potential, on entire nations, that together allow us to
 evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
+based entirely on AS-level topology. Our technique for evaluating AS-level topology using
+\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
 We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
@@
 \par
-Our application of \toolname to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
+Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
+interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
+more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
+we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
 % Internet Freedom
 \par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
+We have validated national breakthrough potential as having a relationship to actual censorship by showing
+the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
+by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
+at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.
 
@@
 censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2018-07-26 Paper/experiment.tex
Break out text files, create gitignore

--- 2018-07-16 Paper/paper.tex
@@
-%&pdflatex
-\documentclass[10pt, conference, letterpaper]{IEEEtran}
-
-\usepackage{xcolor}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{todonotes}
-
-\newcommand{\toolname}{BGP-SAS }
-\newcommand{\timerange}{(2012-2018)}
-
-\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
-
-%Titles:
-%\title{The Evolution of National AS Chokepoints and Their Connection to
-%Internet Freedom}
-\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
-
-\graphicspath{{./images/}}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
-\section{Introduction}
-
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move online and
-are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
-International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 million in 2005 to 3.578 million
-in 2017 \cite{itu}. The majority of these users operate in an
-environment that restricts Internet freedom in some way. For example,
-the 2017 Freedom on the Net report from Freedom House reports that
-64\% of Internet users belong to a nation with Internet that is not
-free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
-the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Also, other forms of Internet
-tampering such as code injection~\cite{badtraffic,191996} have been observed within the
-Internet's backbone. Taken together, these
-trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
-
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
-that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname, for studying national chokepoints on
-the AS graph efficiently. To illustrate these ideas and tools, we
-study how chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-
-%Whether or not AS-level topology supports
-%these efforts is a research question pivotal to an understanding of
-%the dynamics of Internet censorship and surveillance.
-
-%% \par
-%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-%% organize revolutionary movements, and connect with other nations' governments and citizens
-%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-%% of happenings during the Arab Spring can be attributed to social media communications
-%% through the Internet \cite{arabspring}. Because of this potential, national governments may
-%% take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-
-%The Internet is Dynamic and growing
-
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
-%Current Events
-
-%Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial nodes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
-
-% Major Contributions
-\par
-The primary contributions of our work are as follows: 
-\begin{enumerate}
- \item We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations.
- \item We develop a new tool, \toolname, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that our chokepoint measure
-has a significant relationship to Internet freedom as measured by a qualitative source.
-\end{enumerate}
-
-% Rest of Paper Layout
-\par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
-respectively.
-
-\section{Background}
-
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
-
-Because BGP is a distributed protocol, with all connectivity
-information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and updates
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
-% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
-
-\par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
-leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
-
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
-according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname.
-
-\section{Chokepoint Potential}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{bnodes}
-	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
-						AS D is an internal node. ASes E and F are both external nodes.
-						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{figure}
-
-% Why Border ASes?
-In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-the added benefit of making calculations more efficient.
-
-% What is our measure, why does it make sense?
-\par
-We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
-to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
-chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
-
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-\end{equation}
-
-\par
-Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
-the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
-to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
-censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-
 \section{Experimental Setup}
@@
 the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free).
- To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
-Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
-looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
-with each nation's (of those recorded by Freedom House) freedom on the net score.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
-% Internet Freedom
-\par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
-
-\subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
-
-% Points of Presence
-\subsection{Other Chokepoint Measures}
-There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, centralizing
-the DNS infrastructure within the country, or working
-with---and supporting---a set of ISPs that have international connections.
-Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
-We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
-virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
-the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
-of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
-Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
-to show time evolution of chokepoints.
-
-The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
-higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname and
-because the countries that rely on DNS as their main form of censorship are
-fewer in number and are among the least sophisticated in terms of their
-censorship regime.
-
-\subsection{Future Work}
-% OONI data
-While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
-This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
-actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
-recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
-in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
-censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}
+to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national breakthrough potential, particulary the number of border ASes required to intercept
+90\% of paths for each nation and viewing these rankings against the FOTN scores.

### 2018-07-26 Paper/intro.tex
Break out text files, create gitignore

--- 2018-07-25 Paper/paper.tex
@@
-%&pdflatex
-\documentclass[10pt, conference, letterpaper]{IEEEtran}
-
-\usepackage{xcolor}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{todonotes}
-
-%\newcommand{\toolname}{BGP-SAS}
-%\newcommand{\toolname}{BGP Path Finder}
-\newcommand{\toolname}{BGP Path Guide}
-\newcommand{\timerange}{(2012-2018)}
-
-\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}
-
-%Titles:
-%\title{The Evolution of National AS Chokepoints and Their Connection to
-%Internet Freedom}
-\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}
-
-\graphicspath{{./images/}}
-
-\begin{document}
-\maketitle
-
-% \begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-%   In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{national chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname{}, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how national chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-% This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-% path simulation and routing tree datasets to view snapshots of the Internet
-% over multiple years. We provide comprehensive and accessible tools for studying
-% the complex and dynamic AS level Internet topology. Through this approach we
-% can more carefully evaluate the state of the Internet than was previously possible.
-% \end{abstract}
-
-\begin{abstract}
-	Internet topology has been measured and modeled for many years, and this topology has been linked to 
-	censorship and surveillance. Past studies have not, however, quantified the potential for individual 
-	nations to intercept national level Internet paths or analyzed how the evolution of the Internet 
-	topology impacts censorship and surveillance. We introduce a new measure, \emph{national breakthrough
-	potential}, or NBP, to characterize how a country's AS topology is organized in terms of paths that can carry 
-	traffic across international borders.  To study country-level chokepoints, we developed \toolname{}, a 
-	suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the 
-	AS graph. 
-	We find that for many nations the continuous growth of the AS graph is not reflected by a more open
-	flow of information across national borders, and in fact for nations known to conduct censorship we
-	identify a trend toward strengthening chokepoints. Additionally, we find that a significant 
-	relationship exists between NBP and Internet freedom as evaluated from two qualitative sources.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
 \section{Introduction}
@@
 respectively.
-
-\section{Background}
-
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each AS.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
-
-Because BGP is a distributed protocol, with all connectivity
-information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and BGP
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
-% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
-
-\par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
-leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
-
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
-according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname{}.
-
-\section{Chokepoint Measures}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{bnodes}
-	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border ASes. 
-						AS D is an internal AS. ASes E and F are both external ASes.
-						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{figure}
-
-Border ASes are ASes that lie one hop from an AS registered to a different country. It stands to reason that if information bottlenecks
-were to be implemented by national governments, they would likely be implemented on border ASes. This is not always true, but there are more reasons
-to use border ASes as our unit of investigation. For instance, the ratio of the number of border ASes of
-a country to all ASes in that country hints toward the abillity of that nation to intercept information flows across their borders.
-\figurename \ref{fig:bnodes} shows that globally the number of border ASes has grown at a slower rate to the number of non-border ASes
-over time so that the ratio of border ASes to ASes has almost continuously decreased.
-On the surface it would seem that this fact reflects an ever increasing capability for nations to intercept AS-level paths.
-This is not
-neccessarily the case, however. Because AS-level paths are generated by the distributed BGP and the local preferences of AS administrators often do not
-lead to simple shortest path selections, a better way to understand the potential for bottlenecks on border ASes is to investigate the number of BGP generated
-paths that pass through each AS. To do this we propose an intuitive measure on border ASes called \emph{chokepoint potential}. Since this is a measure for single ASes
-we also need an aggregate measure for comparing nations, which we call \emph{national breakthrough potential}.
-
-% % Why Border ASes?
-% In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-% ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-% The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-% of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-% This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-% to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-% the added benefit of making calculations more efficient.
-
-\par
-For our measures we use the term \textit{path} to indicate a path that is generated by BGP consisting of a source AS, a destination AS, and the intermediate
-ASes between them.
-We define a \textit{routing tree} to be a tree rooted at a single destination AS and composed of every acceptable BGP path from the various source ASes that can reach the destination AS.
-We define the chokepoint potential of a border AS to be the ratio of paths intercepted by that border AS
-to the number of paths intercepted by all border ASes belonging to the same nation. For instance, take a border AS $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border ASes within path $p$, then the
-chokepoint potential of $b$, $CP(b)$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border AS in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths. The out-to-in chokepoint potential for a simple dummy nations
-is included in \figurename \ref{fig:chokepoint}.
-
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP(b) = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-\end{equation}
-
-\par
-Given a set of routing trees, chokepoint potential is an intuitive way to compare individual ASes. With this measure,
-the sum of the chokepoint potential for all border ASes for a country is 1.0, as in, all of the border ASes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border ASe minimum are required
-to obtain a certain chokepoint potential. The more border ASes required, the more difficult it would likely be for that nation to perform
-censorship or surveillance, as the government would have to restrict more ISPs, place more filtering hardware, etc. Thus, we can define
-another measure, with this one being a national measure. We define the \emph{national breakthrough potential} of a nation to be the minimum number
-of ASes required to intercept a selected percentage of AS-level paths. Previous work has used 90\% path interception as a benchmark indicating strong control
-of information \cite{throats}, so we use this percentage as well. When used here, however, 90\% path interception indicates a single nation intercepts 90\% of the paths
-that either enter said nation (for the out-to-in case) or exit said nation (for the in-to-out case). National breakthrough potential is calculated taking the strongest ASes first,
-such that the measure reflects the smallest number of border ASes needed to intercept the paths. We use the number of ASes here and not a percentage because using a percentage would
-make it difficult to compare nations. Take for instance the United States, which has many more ASes than most nations. If the United States and another nation both required the same percentage
-of border ASes to intercept 90\% of paths, the similarity of these nations would be misleading, because the United States would actually need to control many more ASes.
-Our technique provides a comparison more likely to reflect reality.
-This measure is called a breakthrough potential because it indicates the number of border ASes that intercept a vast majority of paths, meaning that a country with high NBP would require controlling many border ASes to prevent a breakthrough, or an unfettered access of
-internal ASes to that country.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships as well as the national breakthrough potential
-for all nations represented by the AS graph.
-\toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national breakthrough potential 
-for each nation. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru} to determine which ASes were registered to which nation. 
-This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
-we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We 
-chose this route because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
-Fortunately, ASes with registration changes after each test date tend to not be in the AS graph for that test.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each AS visited, the number of paths intercepted by that AS is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-\toolname{} takes the resulting chokepoint potentials and generates national breakthrough potentials for each nation.
-
-\section{Experimental Setup}
-
-% Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border AS per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
-% Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national breakthrough potential, particulary the number of border ASes required to intercept
-90\% of paths for each nation and viewing these rankings against the FOTN scores.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border ASes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these chokepoint potentials is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border ASes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Turkey) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and national breakthrough potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn}
-	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FOTN score
-	of those nations in 2017. A higher FOTN score indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotn}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotp}
-	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FPI
-	of those nations in 2017. A higher FPI indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotp}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few border ASes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Conclusions and Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
-and national breakthrough potential, on entire nations, that together allow us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating AS-level topology using
-\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
-interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
-more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
-we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
-% Internet Freedom
-\par
-We have validated national breakthrough potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
-by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
-at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.
-
-\subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
-
-% Points of Presence
-\subsection{Other Chokepoint Measures}
-There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, centralizing
-the DNS infrastructure within the country, or working
-with---and supporting---a set of ISPs that have international connections.
-Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
-We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
-virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
-the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
-of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
-Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
-to show time evolution of chokepoints.
-
-The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
-higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname{} and
-because the countries that rely on DNS as their main form of censorship are
-fewer in number and are among the least sophisticated in terms of their
-censorship regime.
-
-\subsection{Future Work}
-% OONI data
-While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
-This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
-actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
-recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
-in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
-censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2018-07-26 Paper/paper.tex
Break out text files, create gitignore

--- 2018-07-16 Paper/paper.tex
@@
 
-\begin{abstract}Internet topology has been measured and modeled for many years, and this topology has been linked to censorship and surveillance.
-Past studies have not, however, quantified the potential for individual nations to intercept national level Internet paths or analyzed how the evolution of the Internet topology impacts censorship and surveillance.
-Over the past decade, governments around the world have recognized that the Internet is a powerful tool for controlling and surveilling their citizens, and they have begun enacting common policies for ASes operating within their borders.
-
-  In this paper, we ask how AS topology has changed over time with respect to national boundaries.
-We introduce a new measure, \emph{chokepoint potential}, to characterize how a country's AS topology is organized in terms of paths that can carry traffic across international borders.
-To study country-level chokepoints, we developed \toolname, a suite of open source, cross platform, and efficient tools for monitoring national chokepoints on the AS graph.
-To illustrate these ideas and tools, we studied how chokepoint potential correlates with two independent measures of civil liberty, finding a significant relationship between our measure and Internet freedom.
-\todo[inline]{NEED A RESULT FOR WHAT WE FOUND FOR EVOLUTION OVER TIME}.
-
-This paper extends earlier research on AS topology and BGP simulation to study chokepoints across the Internet over multiple years, using our
-path simulation and routing tree datasets to view snapshots of the Internet
-over multiple years. We provide comprehensive and accessible tools for studying
-the complex and dynamic AS level Internet topology. Through this approach we
-can more carefully evaluate the state of the Internet than was previously possible.
-\end{abstract}
-
-%% Paths on the Autonomous Systems (AS) graph of the Internet
-%% derived from the Border Gateway Protocol (BGP) can be used by researchers to
-%% understand the dynamics of Internet topology and to interpret how that
-%% topology may enable nations to enact censorship, surveillance, or other
-%% Internet control measures. Unfortunately datasets of these paths are not
-%% generally made publicly available, and paths collected from measurements such as
-%% traceroute or BGP probes tend to be incomplete. Simulation
-%% frameworks have been used to generate AS paths based on inferred AS relationships.
-%% We introduce \toolname, a suite of open source, cross platform, and efficient
-%% tools for monitoring national chokepoints on the AS graph. We introduce chokepoint potential as an important measure of a nation's ability to 
-%% control Internet traffic, either through censorship or surveillance.
-%% Previous research endeavors in this
-%% direction have only identified chokepoints in single snapshots. We apply our
-%% path simulation and routing tree datasets to view snapshots of the Internet
-%% over multiple years in order to introduce a new technique to investigate the
-%% evolution of the complex and dynamic AS level Internet topology. Through this approach we
-%% can more carefully evaluate the state of the Internet than was previously possible. As an
-%% application of \toolname we compare Freedom House's Freedom On The Net (FOTN) score for
-%% Internet freedom and with our chokepoint potential measure in order to interpret
-%% the relationship between AS-level topology and actual censorship activity, providing an
-%% illustration of new ways to monitor which governments can easily control the flow of 
-%% information in their nations and whether they are acting on that potential.
-
-\section{Introduction}
-
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
-tasks, communications, entertainment, and information move online and
-are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
-International Telecommunication Union (ITU), the number of individual
-Internet users increased from 1.024 million in 2005 to 3.578 million
-in 2017 \cite{itu}. The majority of these users operate in an
-environment that restricts Internet freedom in some way. For example,
-the 2017 Freedom on the Net report from Freedom House reports that
-64\% of Internet users belong to a nation with Internet that is not
-free or partly free \cite{FOTN}.  Beyond censorship and surveillance,
-the EU is considering content rules surrounding copyright, net
-neutrality rules in the U.S.\ were recently overturned, and large
-content delivery networks are under enormous pressure to 'do
-something' about fake news and bad actors. Also, other forms of Internet
-tampering such as code injection~\cite{badtraffic,191996} have been observed within the
-Internet's backbone. Taken together, these
-trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
-
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}  In this paper, we focus on AS-level topology and on
-censorship, asking how AS topology has changed over time with respect
-to national boundaries.  We define \emph{chokepoint potential} to
-quantify how a country’s AS topology is organized in terms of paths
-that can carry traffic across international borders.  We developed a
-suite of tools, called \toolname, for studying national chokepoints on
-the AS graph efficiently. To illustrate these ideas and tools, we
-study how chokepoint potential correlates with two independent
-measures of civilian liberty, finding \todo[inline]{NEED A RESULT HERE}. \todo[inline]{NEED A
-  RESULT FOR WHAT WE FOUND FOR EVOLU- TION OVER TIME}.
-
-The paper extends earlier research on AS topology, first by introducing the chokepoint potential measure, then by presenting open-source cross-platform tools for simulating
-and Border Gateway Protocol (BGP) networks, determining chokepoint potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-
-%Whether or not AS-level topology supports
-%these efforts is a research question pivotal to an understanding of
-%the dynamics of Internet censorship and surveillance.
-
-%% \par
-%% The Internet has been used as a tool for the citizens of authoritarian nations to voice opinions,
-%% organize revolutionary movements, and connect with other nations' governments and citizens
-%% to seek aid. Much of the organization, revolutionary momentum, and broadcasting
-%% of happenings during the Arab Spring can be attributed to social media communications
-%% through the Internet \cite{arabspring}. Because of this potential, national governments may
-%% take interest in maximizing their ability to control the flow of information on the Internet.
-%Censorship and Surveillance are growing phenomena
-
-%The Internet is Dynamic and growing
-
-We are interested in how AS topologies have changed as the total size of the graph has grown from about 10,000 nodes in the early 2000s to over 60,000 today.
-The structural
-properties of national AS graphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or
-efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
-%Current Events
-
-%Diverse Censorship Architecture
-\par Every nation has a different layout and count of ASes. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial nodes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-Having views into the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. The measure of chokepoint potential, defined later in
-this paper, is a succint way to estimate the important properties of
-border ASes related to these capabilities.
-
-% Major Contributions
-\par
-The primary contributions of our work are as follows: 
-\begin{enumerate}
- \item We introduce the measure of chokepoint potential and 
-motivate its value as way to interpret the capability for a nation to enact censorship of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing the evolution of the Internet
-allowing for a comparision between nations.
- \item We develop a new tool, \toolname, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item we show that our chokepoint measure
-has a significant relationship to Internet freedom as measured by a qualitative source.
-\end{enumerate}
-
-% Rest of Paper Layout
-\par
-The layout of the rest of this paper is as follows: Section 3 provides relavent background information
-to the problems we are investigating; Section 4 introduces and defines the measure of chokepoint
-potential; Section 5 details our tool for simulating and evaluating AS topologies \toolname; Section 6
-explains our experimental setups; Sections 7, 8, and 9 provide results, discussion, and related work
-respectively.
-
-\section{Background}
-
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each BGP node.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
-
-Because BGP is a distributed protocol, with all connectivity
-information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and updates
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
-% FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
-
-\par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
-leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
-
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
-according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname.
-
-\section{Chokepoint Potential}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{bnodes}
-	\caption{The ratio of ASes that are border ASes plotted over several years.}\label{fig:bnodes}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{chokepoint}
-	\caption{Chokepoint potential example. ASes A,B, and C are all border nodes. 
-						AS D is an internal node. ASes E and F are both external nodes.
-						The out-to-in chokepoint potentials of A,B, and C are 0.25, 0.25, and 0.5 respectively.}\label{fig:chokepoint}
-\end{figure}
-
-% Why Border ASes?
-In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
-The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
-of information. Consider that while the number of ASes globally has continued to grow rapidly, the number of border ASes has grown more slowly.
-This result is depicted in figure \ref{fig:bnodes}. Additionally, while internal chokepoints may intercept many paths, those paths are required
-to have entered through a border AS (in the case of out-to-in paths) or exit through a border AS (for in-to-out paths). This focus on border ASes has
-the added benefit of making calculations more efficient.
-
-% What is our measure, why does it make sense?
-\par
-We define the chokepoint potential of a border node to be the ratio of paths intercepted by that border node
-to the number of paths intercepted by all other border nodes. Consider a border node $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border nodes within path $p$, then the
-chokepoint potential of $b$, $CP_b$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border nodes in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths.
-
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP_b = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-\end{equation}
-
-\par
-Given a set of BGP paths, this measure is an intuitive way to compare individual ASes and different nations. With this measure,
-the chokepoint potential sum over all border nodes for a country is 1.0, as in, all of the border nodes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border nodes minimum are required
-to obtain a certain chokepoint potential. The more border nodes required, the more difficult it would be for that nation to perform
-censorship or surveillance. This is a clear way to differentiate nations based on the topology of their ASes.
-
-\section{\toolname}
-
-% Preprocessing and Data pipeline
-Border Gateway Protocol Simulation, Analysis, and Storage, or \toolname, is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships.
-\toolname first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border node. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru}.
-
-% Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
-
-% Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each node visited, the number of paths intercepted by that node is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-
-\section{Experimental Setup}
-
-% Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border node per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
-% Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free).
- To evaluate our measure, we first must define a way to rank each nation in regards to our chokepoint potential measure.
-Acharya et al. \cite{throats} chose to determine how many ASes were needed (globally) to intercept 90\% of paths as a measure of the AS hierarchy. We choose a similar measure. Because we are
-looking at national comparisons, however, we record how many border ASes are required to incercept 90\% of in-to-out or out-to-in paths for each nation. We compared this number
-with each nation's (of those recorded by Freedom House) freedom on the net score.
-
-\section{Results}
-\subsection{Nations Over Time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesFree}
-	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{NodesCensor}
-	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
-\end{figure}
-\begin{figure*}
-	\centering
-	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
-\end{figure*}
-
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
-
-\par
-To inspect the changes over time for each nation we arrange all the border nodes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these nodes' chokepoint potential is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
-
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border nodes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
-
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
-
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
-
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Iran) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
-
-\subsection{Internet Freedom}
-
-In our test of the relationship between Internet Freedom and chokepoint potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn2017}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{fotn2017}
-	\caption{Number of ASes to intercept 90\% of ito paths vs Freedom On The Net Score (2017). 
-	Green nations are free, yellow are mostly free, and blue are not free according to FOTN.}\label{fig:fotn2017}
-\end{figure}
-
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few nodes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
-
-\section{Related Work}   
-\par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
-\par
-In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
-to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
-
-\section{Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced a novel measure, chokepoint potential, that allows us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating chokepoint potential using
-\toolname provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname to investigate evolutions in the global AS graph over time shows us
-interesting trends in AS topologies.
-% Internet Freedom
-\par
-We have validated chokepoint potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and their FOTN
-score from freedom house.
-
-\subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
-
-% Points of Presence
-\subsection{Other Chokepoint Measures}
-There are several ways that a country can create chokepoints, either intentionally or accidentally: taking advantage of Internet Exchange Points (IXPs),
-limiting the number of physical connections that cross the border, centralizing
-the DNS infrastructure within the country, or working
-with---and supporting---a set of ISPs that have international connections.
-Each of these combines two basic elements: limiting the number of organizations that the government must interact with and creating physical locations where traffic can be controlled.
-We chose to define chokepoint potential in terms of the AS graph instead of physical locations because the AS graph captures the most interesting physical attributes of the Internet, and
-virtualization of the physical and datalink layers on the Internet make physical locations less meaningful. A single undersea cable can be multiplexed to serve as several different links in
-the AS graph, and ASes in one country can have physical Points of Presence (PoPs) in several other countries (e.g., China Telecom in Pasadena, California). Thus, the AS graph is a good approximation
-of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
-Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
-to show time evolution of chokepoints.
-
-The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
-higher layer in the network stack than IP routes.  However, we leave analysis of
-this graph for future work because it is outside the scope of \toolname and
-because the countries that rely on DNS as their main form of censorship are
-fewer in number and are among the least sophisticated in terms of their
-censorship regime.
-
-\subsection{Future Work}
-% OONI data
-While linking chokepoint potential to FOTN scores is a substantial contribution, it still stands that FOTN can serve only as a proxy for censorship.
-This is useful for large trends and general understanding, but a more fine-grained approach could be used to interpret the direct connection between
-actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
-recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
-in chokepoint potential would be a major step in understanding the interplay of censorship and AS-level chokepoints. This process involves designing a way to classify
-censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
+\input{abstract}
+\input{intro}
+\input{background}
+\input{chokepoint}
+\input{tool}
+\input{experiment}
+\input{results}
+\input{related}
+\input{conclusion}
 

### 2018-07-26 Paper/related.tex
Break out text files, create gitignore

\section{Related Work}   
\par
Previous studies have used BGP path models to find ASes that
intercept a high fraction of paths. One such project in \cite{throats}
identified that 90\% of paths on the Internet could be intercepted with only 30
or so ASes. The researchers in \cite{throats} generated paths by first using
only paths to top websites as defined by the Alexa top websites project,
and then appending additional edges to those paths from the full
AS graph according to the principles defined by Gao \cite{gao}. Many of the
ASes that were found to intercept a large number of paths were found to be
within nations that conduct censorship. As an extension of these results,
another paper \cite{decoy} revealed that ASes that intercept many paths could
be utilized for decoy routing. This approach for identifying AS level
chokepoints has several potential pitfalls that are remedied by the approach
taken in this paper. First, the authors in \cite{throats} didn't consider that
many of the paths found from the Alexa top websites would have destinations in
nations that censor Internet traffic, particularly China due to its large
Internet population. This artificially inflates the chokepoint nature of
Chinese ASes. As an alternative, in this paper we consider paths from every
source-destination AS pair, and we make a distinction between in-to-out paths
and out-to-in paths. Secondly, chokepoints have previously only been
identified at a single snapshot of the Internet. This makes it difficult to
discuss the evolution of Internet topology, and whether or not chokepoints are
anomalous or common. Finally, the aforementioned approach for identifying
chokepoints does not allow a clear comparison between different nations. It
can be said that one nation controls a large portion of Internet paths, but
not how easily traffic directed through that country could be intercepted on a
national level. For this, some aggregate measure across all of the border ASes
within a country must be considered, as it is in this paper.     
\par
In \cite{chinafiltering}, Xu et al. investigated the AS level topology of China
to identify where keyword filtering occurred. They found that the most
effective ASes with which to deploy keyword filtering devices are those in the
backbone of the Chinese AS topology. A relevant contribution of
\cite{chinafiltering} is that, while most filtering occurs in border ASes,
some filtering at the time of the study was controlled by non-central provincial
ASes. China had a
diverse strategy for Internet censorship, both targeting chokepoints and the
Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
regards to various AS level topologies motivates the question: Is centralized
censorship or decentralized censorship more common? Instead of directly
identifying censorship devices on the AS graph, we instead quantify the
chokepoint potential of ASes on the national level, and then compare that with
qualitative Internet freedom measures and empirical censorship events.
\par
We are not the first to investigate the relationship between Internet freedom and
AS-level topology. Similar techniques have been used to classify nations according
to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
single moment in time, however, making the results limited in terms of stability and predictability.
Additionaly, previous work has not used country level chokepoints as the link between Internet topology
and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
sheds light on the relationship between topology and Internet freedom, but reveals currently free
nations that could easily implement censorship if their governments decided to.

### 2018-07-26 Paper/results.tex
Break out text files, create gitignore

\section{Results}
\subsection{Nations Over Time}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{NodesFree}
	\caption{The number of border ASes required for the US, France, Germany, and Great Britain to intercept 90\% of in-to-out paths over time.}\label{fig:nodesFree}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{NodesCensor}
	\caption{The number of border ASes required for Russia, Egypt, Turkey, and China to intercept 90\% of in-to-out paths over time.}\label{fig:nodesCensor}
\end{figure}
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{multi}
	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
\end{figure*}

We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
nations is considered highly open.

\par
To inspect the changes over time for each nation we arrange all the border ASes belonging to a nation into a list. The list of border ASes is reverse sorted
so that the first AS has the highest chokepoint potential. The sum of all these chokepoint potentials is 1.0. Then we step through the list, and record
the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
changes over time.

\par
For instance, consider \figurename \ref{fig:multiTime}. 
The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
is the number of border ASes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
farther in the past and more blue lines being more recent.

\par
The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
the very top ASes.

\par
Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.

\par
In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
The distinct behaviors of each nation are rather striking. We can see a steep decline for some
nations known to limit Internet freedom (China and Turkey) as well as the increase for other nations
(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
has required the same number of border ASes to control many paths over several years, this suggests
that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
needed for some nations even more glaring.

\subsection{Internet Freedom}

In our test of the relationship between Internet Freedom and national breakthrough potential, we plotted the Freedom on the Net score
of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fotn}
	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FOTN score
	of those nations in 2017. A higher FOTN score indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotn}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fotp}
	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FPI
	of those nations in 2017. A higher FPI indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotp}
\end{figure}

\par
For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
few border ASes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
that the censorship efforts in Russia might be of types that do not require AS level chokepoints.

### 2018-07-26 Paper/tool.tex
Break out text files, create gitignore

\section{\toolname}

% Preprocessing and Data pipeline
\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships as well as the national breakthrough potential
for all nations represented by the AS graph.
\toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national breakthrough potential 
for each nation. In our experiments, we used the country codes returned from Team Cymru's
IP to ASN whois service \cite{cymru} to determine which ASes were registered to which nation. 
This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We 
chose this route because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
Fortunately, ASes with registration changes after each test date tend to not be in the AS graph for that test.

% Routing Tree Algorithm
\par
In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
routing tree algorithm that is ready to work on most hardware.

% Chokepoint calculation
\par
Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
traversed. For each AS visited, the number of paths intercepted by that AS is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
\toolname{} takes the resulting chokepoint potentials and generates national breakthrough potentials for each nation.

### 2018-07-26 Paper/intro.tex
Changes to intro

@@
 neutrality rules in the U.S.\ were recently overturned, and large
-content services are under enormous pressure to 'do
+content services are under enormous pressure to `do
 something' about fake news and bad actors. Also, other forms of Internet
@@
 Autonomous Systems are groups of routers under common management, such as the network of
-a University or an Internet Service Provider (ISP). We apply AS-level topologies to
-censorship and surveillance by evaluating border ASes, or ASes that can connect directly to at least
-one AS from another nation according to a new measure \emph{chokepoint potential}.
-With this measure we can ask how AS topology has changed over time with respect
-to national boundaries.  We define \emph{national breakthrough potential} to
-measure how a country's AS topology is organized in terms of paths
-that can carry traffic across international borders. We developed a
+a university or an Internet Service Provider (ISP). The change in national AS topologies 
+is of interest to censorship research. The total size of the AS graph (The global network 
+of all ASes) has grown from about 10,000 ASes in the early 2000s to over 60,000 today. 
+The structural properties of national subgraphs have evolved differently
+from nation to nation whether from economic decision making, infrastructural necessities, 
+or efforts to build a powerful censorship and surveillance network.\Ben[inline]{Need citations for economic, infrastructure, censorship}.
+%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
+%The locations and relationships of these ASes determine how many chokepoints
+%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
+This rapid expansion, together with the changing role of national
+governments, points to the importance of understanding properties like
+path robustness, AS hierarchies, and the potential for organizations to
+control information as it flows in and out of their networks.
+
+\par Every nation has a different layout and number of ASes, and these national networks
+connect to foreign ASes in unique ways. The
+censorship and surveillance strategies of nations also differ. For
+instance, China both conducts keyword filtering in border ASes and in
+internal provincial ASes \cite{chinafiltering}, while Iran routes its
+Internet traffic through a centralized facility \cite{irancensor}.
+A picture of a nation's censorship or surveillance capabilities then
+is not complete without examining its underlying AS topology.
+Having the capability to investigate the AS topology of a nation, then, will both help
+researchers identify nations that could easily conduct censorship and
+also provide possible insight into what kind of censorship is likely
+being conducted.
+
+We examine how censorship and surveillance might be conducted 
+at the AS-level by evaluating border ASes, or ASes that can connect directly to at least
+one AS from another nation. We introduce a new measure called \emph{chokepoint potential} which
+captures the ability of a node to control communication to the Interent outside a nation's borders.
+We can then aggregate this measure and define \emph{national breakthrough potential} to
+measures the concentration of routing paths into and out of a country. 
+\Ben[inline]{These definitions are important. They need to be a concise and clear definitions. I don't think it's right yet, but we are close}
+With these measures 
+we can ask how the AS-level topology has changed over time with respect
+to national boundaries.
+
+We developed a
 suite of tools, called \toolname{}, for studying national chokepoints on
@@
 study how national chokepoint potential correlates with two independent
-measures of civilian liberty, finding that a significan relationship exists between NBP and both liberty measures.
+measures of civil liberty, finding that significant relationship exists between NBP and both measures.
 The paper extends earlier research on AS topology, first by introducing the NBP measure, 
@@
 determining breakthrough potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
+Our measure of border ASes, chokepoint potential, defined in detail
+in section 3, is a succinct way to estimate the important properties of
+border ASes related to the ability of a nation to censor and surveil, and our aggregate extension of this measure to
+the national level, national breakthrough potential, is an effective way to evaluate and
+compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.\Ben[inline]{This paragraph seems a little redundant now.
+It is a combo of other previous writing, but I am not sure what to cut.}
 
@@
 
-The change in national AS topologies is of interest to censorship research.
-The total size of the AS graph (The global network of all ASes) has grown from about 10,000 ASes in the early 2000s to over 60,000 today. 
-The structural properties of national subgraphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, or efforts to build a powerful censorship and surveillance
-network.
-%The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.
-%The locations and relationships of these ASes determine how many chokepoints
-%of AS-level routing exist and what strength these chokepoints have in regards to paths intercepted.
-This rapid expansion, together with the changing role of national
-governments, points to the importance of understanding properties like
-path robustness, AS hierarchies, and the potential for organizations to
-control information as it flows in and out of their networks.
-
 %Current Events
@@
 %Diverse Censorship Architecture
-\par Every nation has a different layout and number of ASes, and these national networks
-connect to foreign ASes in unique ways. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial ASes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-A picture of a nation's censorship or surveillance capabillities then
-is not complete without a topological view point.
-Having the capabillity to investigate the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted. Our measure of border ASes, chokepoint potential, defined in detial
-in section 3, is a succint way to estimate the important properties of
-border ASes related to these capabilities, and our aggregate extension of this measure to
-the national level, national breakthrough potential, is an effective way to evaluate and
-compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.
 
@@
 \begin{enumerate}
- \item We introduce the measure of chokepoint potential for single border ASes and the aggregate measure national breakthrough potential, 
-motivating their value as a way to interpret the capability for a nation to enact censorship or surveillance of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, showing how the evolution of the Internet has either facillitated stronger
+ \item We introduce the measure of chokepoint potential for single border ASes and the 
+ aggregate measure national breakthrough potential, 
+motivating their value as a way to interpret the capability for a 
+nation to enact censorship or surveillance of its Internet traffic.
+ \item We provide an overview of the evolution of national AS-level chokepoints over time, 
+ showing how the evolution of the Internet has either facilitated stronger
  AS-level chokepoints or lead to more open flow of information across borders.
@@
 to the problems we are investigating; Section 3 introduces and defines the measures of chokepoint
-potential and national breakthrough potential; Section 4 details our tool for simulating and evaluating AS topologies \toolname{}; Section 5
+potential and national breakthrough potential; Section 4 details our tool for simulating and evaluating 
+AS topologies \toolname{}; Section 5
 explains our experimental setups; Sections 6, 7, and 8 provide results, discussion, and related work

### 2018-07-29 Paper/background.tex
Updates to background

@@
 
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each AS.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
+The AS-layer is the highest level of organization of the Internet, with each
+individual AS represents a separately administered network, for example, a
+large ISP, a university network, or a government entity. The Border Gateway
+Protocol (BGP) is used to route traffic among ASes \cite{bgp}.  AS-level routes
+are known as \emph{paths} and are specified by routing tables stored locally in
+each AS. These tables express known connections to other ASes and local
+preferences for how data are routed, e.g., if there are multiple paths to a
+single destination, an AS may prefer the path that costs the least.
 
@@
 information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
+infer the exact AS topology at any point in time.  Several strategies have been 
+developed to infer AS topology, each with advantages and drawbacks.  Methods that rely
 on routing table information and BGP
@@
 % FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
+Many sources are needed to infer a complete picture of the BGP topology using local tables. 
+This approach does not scale for
+global studies, particularly those looking at the evolution of the
+topology over time.
+Similarly, methods based on empirical traceroute data are known to be 
+incomplete or inaccurate~\cite{tracerouteProblems}.
 
 \par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
+truth data for AS relationships and BGP paths. Two ASes might enter a peering relationship,
+exchanging traffic but not announcing their connection. For AS relationships,
 inference is often used based on economic considerations to classify
 the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
+maximizing the occurrences of certain economic rules on the AS graph by
 choosing a particular set or relationships. This technique was only
@@
 \par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
+a challenge.\Ben[inline]{Be more clear about the difference between `Finding BGP paths' and what you have discussed above.}
+Packet based simulations, wherein BGP is simulated
 directly, are computationally infeasible for the scales relevant to a
@@
 relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
+search (BFS) algorithm. The BFS adds edges to routing trees first
 according to local preference (LP), then shortest path (SP), and
@@
 between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
+root of the routing tree). We find this technique suitable \Ben[inline]{Why is it suitable?} for our
+purposes. We use the basic approach of BGPSim (in ways explained further in this
 paper) as part of \toolname{}.

### 2018-07-29 Paper/chokepoint.tex
Changes to chokepoint potential

@@
-\section{Chokepoint Measures}
+\section{Border ASes and Chokepoint Potential}
+
+\Ben[inline]{Include a little paragraph here about what this section describes}.
 
@@
 
-Border ASes are ASes that lie one hop from an AS registered to a different country. It stands to reason that if information bottlenecks
-were to be implemented by national governments, they would likely be implemented on border ASes. This is not always true, but there are more reasons
-to use border ASes as our unit of investigation. For instance, the ratio of the number of border ASes of
-a country to all ASes in that country hints toward the abillity of that nation to intercept information flows across their borders.
-\figurename \ref{fig:bnodes} shows that globally the number of border ASes has grown at a slower rate to the number of non-border ASes
-over time so that the ratio of border ASes to ASes has almost continuously decreased.
-On the surface it would seem that this fact reflects an ever increasing capability for nations to intercept AS-level paths.
-This is not
-neccessarily the case, however. Because AS-level paths are generated by the distributed BGP and the local preferences of AS administrators often do not
-lead to simple shortest path selections, a better way to understand the potential for bottlenecks on border ASes is to investigate the number of BGP generated
-paths that pass through each AS. To do this we propose an intuitive measure on border ASes called \emph{chokepoint potential}. Since this is a measure for single ASes
-we also need an aggregate measure for comparing nations, which we call \emph{national breakthrough potential}.
+\subsection{Border Nodes}
+
+Border ASes are ASes that lie adjacent from an AS registered to a different
+country along a routing path. More formally, we define $g(\cdot)$ as the
+mapping from ASes to the country of registration, and say an AS $a_u$ is a border
+AS if $g(a_u) \neq g(a_v)$, where $a_v$ is a neighboring AS.  Because of their
+position along national border, these ASes, can act as convenient locations for
+national governments to control information flow into or out of a country. This
+is especially true if a large number of BGP paths traverse a small number of
+border ASes. Even as countries expand their Internet infrastructure, they may
+be concentrating traffic that crosses national borders to a small number of
+governmentally controlled ASes.  Evidence for this can be seen in the ratio of
+the number of border ASes of a country to all ASes in that country.
+%hints toward the abillity of that nation to intercept information flows across
+%their borders.
+\figurename \ref{fig:bnodes} shows that globally the number of border ASes has
+grown at a slower rate than the number of non-border ASes.
+%over time so that the ratio of border ASes to ASes has almost continuously
+%decreased. 
+It would seem that this fact reflects an ever increasing capability for nations
+to intercept AS-level paths.  However, an important consideration is not only
+how many border nodes exist, but also how often each is utilized in BGP
+routing. Because AS-level paths are generated by distributed BGP routing tables
+and the local preferences of AS administrators, a better way to understand the
+potential for bottlenecks on border ASes is to investigate the number of BGP
+generated paths that pass through each AS. We propose an intuitive measure of
+AS bottlenecks called \emph{chokepoint potential}. Surveillance and censorship
+generally occurs at a national level so we further define an aggregate measure
+for nations, which we call \emph{national chokepoint potential}.
 
@@
 
-\par
-For our measures we use the term \textit{path} to indicate a path that is generated by BGP consisting of a source AS, a destination AS, and the intermediate
-ASes between them.
-We define a \textit{routing tree} to be a tree rooted at a single destination AS and composed of every acceptable BGP path from the various source ASes that can reach the destination AS.
-We define the chokepoint potential of a border AS to be the ratio of paths intercepted by that border AS
-to the number of paths intercepted by all border ASes belonging to the same nation. For instance, take a border AS $b$, belonging to country $c$.
-If $P_c$ is the set of paths in to or out of country $c$ and $B_p$ is the set of border ASes within path $p$, then the
-chokepoint potential of $b$, $CP(b)$ is defined formally in equation \ref{eqn:chokePointPotential}. This measure captures
-the relative strength exhibited by a border AS in regards to what portion of paths it intercepts. This is calculated
-seperately for in-to-out paths (those starting from a source in the country in question) and out-to-in paths. The out-to-in chokepoint potential for a simple dummy nations
-is included in \figurename \ref{fig:chokepoint}.
+\subsection{Chokepoint Potential}
 
-% Formal Definition
-\begin{equation}\label{eqn:chokePointPotential}
-CP(b) = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
+We define a BGP path of length $n$ from source AS $a_0$ to destination AS $a_n$
+as a sequence $p_{1\rightsquigarrow n} = \{a_i : i \in [1,n]\}$.  We define a
+\textit{routing tree} to be a tree rooted at a single destination AS and
+composed of every acceptable BGP path from the various source ASes that can
+reach the destination AS.\Ben[inline]{Is this always a tree? Seems like it
+could be a DAG sometimes (especially if random choices are made). Also do we
+need to define routing trees if we only really work on the path}  The
+chokepoint potential of a border AS is the ratio of paths intercepted by that
+border AS to the number of paths intercepted by all border ASes belonging to
+the same nation. Formally, for a country $c$, let $P = \{p_{i \rightsquigarrow
+j} : g(a_i)=c \text{ and } g(a_j)\neq c\} $, i.e.\ the set of all path
+originating in $c$ and ending in outside $c$. Then the chokepoint potential,
+$cp(\cdot)$, of a border AS $a_u$ with $g(a_u) = c$ is
+
+\begin{equation}
+  \label{eq:chokepoint}
+  cp(a_u) = \frac{|\{p: p \in P \text{ and } a_u \in P\}|}{|P|}
 \end{equation}
 
-\par
-Given a set of routing trees, chokepoint potential is an intuitive way to compare individual ASes. With this measure,
-the sum of the chokepoint potential for all border ASes for a country is 1.0, as in, all of the border ASes collectively control the
-flow of information over the nation's border. To compare one nation to another, we can inspect how many border ASe minimum are required
-to obtain a certain chokepoint potential. The more border ASes required, the more difficult it would likely be for that nation to perform
-censorship or surveillance, as the government would have to restrict more ISPs, place more filtering hardware, etc. Thus, we can define
-another measure, with this one being a national measure. We define the \emph{national breakthrough potential} of a nation to be the minimum number
-of ASes required to intercept a selected percentage of AS-level paths. Previous work has used 90\% path interception as a benchmark indicating strong control
-of information \cite{throats}, so we use this percentage as well. When used here, however, 90\% path interception indicates a single nation intercepts 90\% of the paths
-that either enter said nation (for the out-to-in case) or exit said nation (for the in-to-out case). National breakthrough potential is calculated taking the strongest ASes first,
-such that the measure reflects the smallest number of border ASes needed to intercept the paths. We use the number of ASes here and not a percentage because using a percentage would
-make it difficult to compare nations. Take for instance the United States, which has many more ASes than most nations. If the United States and another nation both required the same percentage
-of border ASes to intercept 90\% of paths, the similarity of these nations would be misleading, because the United States would actually need to control many more ASes.
-Our technique provides a comparison more likely to reflect reality.
-This measure is called a breakthrough potential because it indicates the number of border ASes that intercept a vast majority of paths, meaning that a country with high NBP would require controlling many border ASes to prevent a breakthrough, or an unfettered access of
-internal ASes to that country.
+The above definition is specifically for capturing paths originating in country
+$c$ and routing outside the country, which we will refer to as
+\textit{outgoing} chokepoint potential or $cpo$. We can similarily define
+\textit{incoming} chokepoint potential or $cpi$, by simply redefining $P$ as $\{p_{i
+\rightsquigarrow j}: g(a_i) \neq c \text{ and } g(a_j) = c\}$. That is all
+paths originating outside of $c$ but terminating in $c$ and recomputing
+Equation~\ref{eq:chokepoint}.
+
+%For instance, take
+%a border AS $b$, belonging to country $c$.  If $P_c$ is the set of paths in to
+%or out of country $c$ and $B_p$ is the set of border ASes within path $p$, then
+%the chokepoint potential of $b$, $CP(b)$ is defined formally in equation
+%\ref{eqn:chokePointPotential}. 
+%This is calculated seperately for in-to-out paths (those starting from a source
+%in the country in question) and out-to-in paths.
+The incoming chokepoint
+potential for a nations is illustrated in \figurename
+\ref{fig:chokepoint}.
+
+\Ben[inline]{A couple of questions here. Why mess with incoming (out to in) and
+outgoing (in to out) measures? Can we just define it as $\frac{|p_{in}| +
+|p_{out}|}{|P_{in}| + |P_{out}|}$, that is the incoming and outgoing paths over
+the total paths? Would this be that much different? Is there are an argument to
+be made that out-to-in or in-to-out is a better? My guess is they are all
+pretty correlated, I can check on that}.
+% Formal Definition
+%\begin{equation}\label{eqn:chokePointPotential}
+%CP(b) = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
+%\end{equation}
+
+Given a set of routing trees\Ben[inline]{paths?}, chokepoint potential is an intuitive way to
+compare individual ASes. With this measure, the sum of the chokepoint potential
+for all border ASes for a country is 1.0, as in, all of the border ASes
+collectively control the flow of information over the nation's border. We
+define the aggregate \emph{national chokepoint potential} as the number of
+border ASes required to control a particular fraction $f$ of paths. Formally,
+given a country,$c$, with a set of $k$ border ASes $\{a_1,a_2,\cdots,a_k\}$ with
+$cp(a_i)>cp(a_j) \forall i<k$ (i.e.\ sorted in descending order), the
+we find the smallest $j$ such that
+
+\begin{equation}
+  \label{eq:nationalchokepoint}
+  \sum_{i=1}^j cp(a_i) > f
+\end{equation}
+
+and define national chokepoint potential as $CP(c,f) = 1/j$. We use the
+multiplicative inverse here to preserve a clearer semantic meaning i.e.\ higher
+chokepoint potential implies more control over paths. The more border ASes
+required, the more difficult it would likely be for that nation to perform
+censorship or surveillance, as the government would have to restrict more ISPs,
+place more filtering hardware, etc. 
+%Thus, we can define another measure, with
+%this one being a national measure. We define the \emph{national breakthrough
+%potential} of a nation to be the minimum number of ASes required to intercept a
+%selected percentage of AS-level paths. 
+Previous work has used $f=0.90$ as a benchmark indicating strong control of information
+\cite{throats}, so we use this percentage as well. When used here, however,
+90\% path interception indicates a single nation intercepts 90\% of the paths
+that either enter said nation (for the out-to-in case) or exit said nation (for
+the in-to-out case). National breakthrough potential is calculated taking the
+strongest ASes first, such that the measure reflects the smallest number of
+border ASes needed to intercept the paths.\footnote{We note that a dual measure
+could be defined as the fraction of paths controlled by $j$ ASes. Our results
+are substantively the same for either definition.\Ben[inline]{Check to make
+sure}}.
+
+We measure the absolute number of ASes required to control a specific
+fraction rather than a percentage within a country because it makes for more
+equitable comparisons between nations
+Take for instance the United States, which has many more ASes than
+most nations. If the United States and another nation both required the same
+percentage of border ASes to intercept 90\% of paths, the similarity of these
+nations would be misleading. Controlling the larger absolute number of ASes
+would require significantly more resources.\Ben[inline]{we need to actually
+pick another nation where this is true to make this example illustrative.} Our technique provides a comparison more likely to
+reflect reality. 
+%This measure is called a breakthrough potential because it
+%indicates the number of border ASes that intercept a vast majority of paths,
+%meaning that a country with high NBP would require controlling many border ASes
+%to prevent a breakthrough, or an unfettered access of internal ASes to that
+%country.

### 2018-07-29 Paper/chokepoint.tex
updates to equations

@@
 given a country,$c$, with a set of $k$ border ASes $\{a_1,a_2,\cdots,a_k\}$ with
-$cp(a_i)>cp(a_j) \forall i<k$ (i.e.\ sorted in descending order), the
+$cp(a_i)>cp(a_j) \forall i<j$ (i.e.\ sorted in descending order), the
 we find the smallest $j$ such that

### 2018-07-30 Paper/tool.tex
Lots of tool questions

@@
 
+In this section we describe \toolname{}, its relationship to previous attempts
+at identifying routing paths, and how it is used to calculate chokepoint
+potential.
+
 % Preprocessing and Data pipeline
-\toolname{} is a new tool that calculates the chokepoint potential for every border AS given a set of AS relationships as well as the national breakthrough potential
-for all nations represented by the AS graph.
-\toolname{} first takes the entire AS graph, as in from the CAIDA dataset \cite{CAIDA}, and uses the principles of BGPSim \cite{quicksand}
-to generate a set of routing trees. These routing trees, as well as a set of AS country codes (for identifying which nation an AS belongs to)
-are used to determine the chokepoint potentials for every border AS. Finally, \toolname{} calclates the national breakthrough potential 
-for each nation. In our experiments, we used the country codes returned from Team Cymru's
-IP to ASN whois service \cite{cymru} to determine which ASes were registered to which nation. 
-This service only maintains the most recent registration of each AS, so to control for uncertain country to AS links
-we keep ASes that have been registered more recently than the current timestam being tested in the path simulation but we do not assign these ASes chokepoint potentials. We 
-chose this route because the datasets of AS registrations for the various ASN registries tended to have more missing data at specific timestamps.
-Fortunately, ASes with registration changes after each test date tend to not be in the AS graph for that test.
+\toolname{} is a new tool that simluates plausible BGP paths and calculates AS
+chokepoint potential and national chokepoint potential.  \toolname{} uses
+readily available AS relationship data, e.g. data made available from
+CAIDA~\cite{CAIDA}, and uses a similar approach to BGPSim \cite{quicksand} to
+generate a set of routing trees. \Ben[inline]{Need a short description of how
+BGPSim works, maybe below.}. These routing trees, as well as a set of AS
+country codes (for identifying which nation an AS belongs to) are used to
+determine the chokepoint potentials border ASes.  Finally, \toolname{}
+calclates the national breakthrough potential for each nation. In our
+experiments, we used the country codes returned from Team Cymru's IP to ASN
+whois service \cite{cymru} to determine which ASes were registered to which
+nation.  This service only maintains the most recent registration of each AS,
+so to control for uncertain country to AS links we keep ASes that have been
+registered more recently than the current timestamp being tested in the path
+simulation but we do not assign these ASes chokepoint
+potentials. We chose this
+route because the datasets of AS registrations for the various ASN registries
+tended to have more missing data at specific timestamps.  Fortunately, ASes
+with registration changes after each test date tend to not be in the AS graph
+for that test. \Ben[inline]{Two things. 1) This is sort of confusing, see if you
+can clarify it. 2) Is this a big deal? How often does this happen? If it
+happens frequently do we expect it to affect our results? }
 
 % Routing Tree Algorithm
-\par
-In order to calculate the routing trees, we use an extended version of the BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In
-our work we addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes for each path it considers but not the order in which
-they are visited; (2) Once routing trees are determined, they cannot be accessed later without recalculation; (3) BGPSim relies on the outdated parallelization
-framework DryadLinq for C\#. To address these issues, we use a Python implementation we dub BGPSimPy. BGPSimPy returns ordered paths from its routing trees,
-saves routing trees to disk after calculation, and is parallelized with MPI via the mpi4py library. These improvements have the added benefit of yielding a cross platform
-routing tree algorithm that is ready to work on most hardware.
+\par In order to calculate the routing trees, we use an extended version of the
+BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In our work we
+addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes
+for each path it considers but not the order in which they are visited; (2)
+Once routing trees are determined, they cannot be accessed later without
+recalculation; (3) BGPSim relies on the outdated parallelization framework
+DryadLinq for C\#. To address these issues, we implemented an our routing
+simulation tool in Python. \Ben[inline]{Why does BGPSim not return ordered
+paths, this seems trivial, but I don't know the implementation details, was it
+just an oversight or did you do something cool?}\toolname{} returns ordered paths from its routing trees, saves
+routing trees to disk after calculation, and is parallelized with MPI via the
+mpi4py library. These improvements have the added benefit of yielding a cross
+platform routing tree algorithm that is ready to work on most hardware.
 
 % Chokepoint calculation
-\par
-Once BGPSimPy generates the routing trees, they can be processed to determine chokepoint potentials. This is done by iterating over every path between each AS-pair. Because
-we use the same random tie-break as BGPSim, there is only 1 path between each AS-pair considered, even if multiple exist in the routing tree. Once a path is determined, it is
-traversed. For each AS visited, the number of paths intercepted by that AS is incremented. This is done for both in-to-out paths and out-to-in paths, so only one traversal is
-necessary per path. Additionally, the number of paths of each type that belong to each nation is tallied, as this makes up the denominator in equation \ref{eqn:chokePointPotential}.
-\toolname{} takes the resulting chokepoint potentials and generates national breakthrough potentials for each nation.
+\par Once \toolname{} generates the routing trees, they can be processed to
+determine chokepoint potentials. This is done by iterating over every path
+between each AS-pair. Because we use the same random tie-break as BGPSim, there
+is only 1 path between each AS-pair considered, even if multiple exist in the
+routing tree.\Ben[inline]{Put stuff in about tiebreaking being rare and
+unlikely to affect results} Once a path is determined, it is traversed. For each AS visited,
+the number of paths intercepted by that AS is incremented. This is done for
+both outgoing paths and incoming paths, so only one traversal is necessary
+per path. Additionally, the number of paths of each type that belong to each
+nation is tallied, as this makes up the denominator in equation
+\ref{eqn:chokePointPotential}.  \toolname{} takes the resulting chokepoint
+potentials and generates national breakthrough potentials for each nation.

### 2018-07-30 Paper/experiment.tex
UPdates to experiments section

@@
 % Chokepoint Evolution
-There is ongoing research interest into ASes that intercept large portions of Internet paths \cite{throats}, \cite{decoy}. Some important questions remain unanswered however. For instance, to what
-extent does the current state of the Internet support national governments' attempts to censor Internet traffic? How does this vary from nation to nation? Is the Internet developing more
-powerful chokepoints, or becoming more evenly accessible?
-\par
-To probe these questions we first used our chokepoint evaluation technique to investigate the chokepoint potential of all nations for the current Internet as well as the
-change in chokepoint potentials over time. We looked at multiple snapshots of AS relationships from the CAIDA dataset (2012-2018).
-For each timestamp, we generated routing trees based on these relationships. Then we calculated the chokepoint potential for every border AS per snapshot. As a result we can investigate
-how countries have changed overtime in their capability to enact censorship and surveillance. This is an attempt to understand what topological trends have developed historically.
-With this test we can compare nations, and see which ones have overtime increased their capability to control the flow of information across their borders.
-\par
+%There is ongoing research interest into ASes that intercept large portions of
+%Internet paths \cite{throats}, \cite{decoy}. Some important questions remain
+%unanswered however. For instance, to what extent does the current state of the
+%Internet support national governments' attempts to censor Internet traffic? How
+%does this vary from nation to nation? Is the Internet developing more powerful
+%chokepoints, or becoming more evenly accessible?
+
+In this section we use chokepoint potential to examine three questions: 1) to
+what extent does the current state of the AS-level Internet topology support
+national governments' ability to censor and surveil Internet traffic? 2) How
+different is this ability vary across nations? and 3)  How has this ability
+evolved over time?
+
+To probe these questions we first used our chokepoint evaluation technique to
+investigate the chokepoint potential of all nations for the current Internet as
+well as the change in chokepoint potentials over time. We looked at multiple
+snapshots of AS relationships from the CAIDA dataset (2012-2018).  For each
+timestamp, we generated routing trees based on these relationships. Then we
+calculated the chokepoint potential for every border AS per snapshot using
+toolname{}. As a result we can investigate how countries have changed overtime
+in their capability to enact censorship and surveillance. This is an attempt to
+understand what topological trends have developed historically.  With this test
+we can compare nations, and see which ones have overtime increased their
+capability to control the flow of information across their borders.
+
 % Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily implement censorship, it stands to reason that their might already be a negative relationship between
-the chokepoint potential of a nation and its Internet freedom. If a significant relationship were to be found it would strengthen chokepoint potential as a measure of a nation's censorship capability
-and it would increase the value in monitoring chokepoint potentials across the globe.
-\par
-We tested whether a significant relationship exists between our measure of national chokepoints and a qualitative evaluation of Internet Freedom. For Internet Freedom we used
-the Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries. Each country receives a numerical score from 0 (the most free)
-to 100 (the least free). We compare our results to the FOTN scores by ranking nations according to national breakthrough potential, particulary the number of border ASes required to intercept
-90\% of paths for each nation and viewing these rankings against the FOTN scores.
+If chokepoint potential can be leveraged to determine if a nation can easily
+implement censorship, it stands to reason that their might already be a
+negative relationship between the chokepoint potential of a nation and its
+Internet freedom. If a significant relationship were to be found it would
+strengthen chokepoint potential as a measure of a nation's censorship
+capability and it would increase the value in monitoring chokepoint potentials
+across the globe.
+
+We tested whether a significant relationship exists between our measure of
+national chokepoints and a qualitative evaluation of Internet Freedom. For
+Internet Freedom we used the Freedom House's Freedom On The Net report
+\cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries.
+Each country receives a numerical score from 0 (the most free) to 100 (the
+least free). We compare our results to the FOTN scores by ranking nations
+according to national breakthrough potential, particulary the number of border
+ASes required to intercept 90\% of paths for each nation and viewing these
+rankings against the FOTN scores.

### 2018-07-30 Paper/results.tex
Just formatting

@@
 
-We have chosen to detail the results of our experiment for 8 nations due to their histories of censorship practice. First we have chosen 4 nations known to enact
-extensive Internet censorship policies. These nations are China, Turkey, Egypt, and Russia. China has historically been the number one nation in regards to Internet
-censorship \cite{censorshipSurvey}. Turkey has recently dramatically increased its online censorship efforts, by blocking social media accesses during turmoil and pressuring ISPs
-to block access to material chosen by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the Russian government access to user statistics \cite{censorshipGeography}. Next we have chosen the nations
-of Germany, the United States, France, and the United Kingdom because, while they have all conducted limited forms of Internet censorship \cite{censorshipSurvey}, Internet access for these 
-nations is considered highly open.
+We have chosen to detail the results of our experiment for 8 nations due to
+their histories of censorship practice. First we have chosen 4 nations known to
+enact extensive Internet censorship policies. These nations are China, Turkey,
+Egypt, and Russia. China has historically been the number one nation in regards
+to Internet censorship \cite{censorshipSurvey}. Turkey has recently
+dramatically increased its online censorship efforts, by blocking social media
+accesses during turmoil and pressuring ISPs to block access to material chosen
+by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered
+severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
+The Russian government has passed surveillance laws requiring ISPs to allow the
+Russian government access to user statistics \cite{censorshipGeography}. Next
+we have chosen the nations of Germany, the United States, France, and the
+United Kingdom because, while they have all conducted limited forms of Internet
+censorship \cite{censorshipSurvey}, Internet access for these nations is
+considered highly open.
 
-\par
-To inspect the changes over time for each nation we arrange all the border ASes belonging to a nation into a list. The list of border ASes is reverse sorted
-so that the first AS has the highest chokepoint potential. The sum of all these chokepoint potentials is 1.0. Then we step through the list, and record
-the cumulative chokepoint potential of all the ASes seen so far. If we plot the number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control different percentages of paths. We repeat this process for each snapshot to highlight
-changes over time.
+To inspect the changes over time for each nation we arrange all the border ASes
+belonging to a nation into a list. The list of border ASes is reverse sorted so
+that the first AS has the highest chokepoint potential. The sum of all these
+chokepoint potentials is 1.0. Then we step through the list, and record the
+cumulative chokepoint potential of all the ASes seen so far. If we plot the
+number of ASes controled vs the cumulative chokepoint potential for that number
+of ASes, we can see how many ASes are required for a certain nation to control
+different percentages of paths. We repeat this process for each snapshot to
+highlight changes over time.
 
-\par
-For instance, consider \figurename \ref{fig:multiTime}. 
-The x-axis (log-scale, so that countries with large differences in AS counts can be compared more easily) for each plot
-is the number of border ASes controlled, and the y-axis is the ratio of in-to-out paths intercepted. Each line represents a different snapshot, with the more red lines being
-farther in the past and more blue lines being more recent.
+For instance, consider \figurename \ref{fig:multiTime}.  The x-axis (log-scale,
+so that countries with large differences in AS counts can be compared more
+easily) for each plot is the number of border ASes controlled, and the y-axis
+is the ratio of in-to-out paths intercepted. Each line represents a different
+snapshot, with the more red lines being farther in the past and more blue lines
+being more recent.
 
-\par
-The United States and China are shown here and exhibit some dramatic differences. First, it is worth pointing out that the United States has many more ASes than China, hence
-it's line extends further to the right in these plots. We also see that China, in all cases, can control a much larger portion of its paths with much fewer ASes than the
-United States. This result is entirely expected. Of more interest is the trends that can be seen over time. The AS-level topology of China's Internet has evolved such that very few
-ASes can intercept nearly all BGP paths. The US has evolved in a different way. While it has become somewhat easier for the US to intercept most of its paths, it has become
-more difficult to intercept around 70\% of its paths and higher. This suggest an expansion of ASes on the interior and more connections, as well as a strengthening of
-the very top ASes.
+The United States and China are shown here and exhibit some dramatic
+differences. First, it is worth pointing out that the United States has many
+more ASes than China, hence it's line extends further to the right in these
+plots. We also see that China, in all cases, can control a much larger portion
+of its paths with much fewer ASes than the United States. This result is
+entirely expected. Of more interest is the trends that can be seen over time.
+The AS-level topology of China's Internet has evolved such that very few ASes
+can intercept nearly all BGP paths. The US has evolved in a different way.
+While it has become somewhat easier for the US to intercept most of its paths,
+it has become more difficult to intercept around 70\% of its paths and higher.
+This suggest an expansion of ASes on the interior and more connections, as well
+as a strengthening of the very top ASes.
 
-\par
-Not all nations have evolved to a state where it easier to control paths. One example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a fairly constant
-trend shows that it has become more difficult to intercept BGP paths. Unlike the other examples, any amount of German border ASes intercept less paths in more recent tests.
+Not all nations have evolved to a state where it easier to control paths. One
+example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a
+fairly constant trend shows that it has become more difficult to intercept BGP
+paths. Unlike the other examples, any amount of German border ASes intercept
+less paths in more recent tests.
 
-\par
-In order to compare multiple nations, we have plotted the number of ASes needed to intercept 90\% of in-to-out paths
-for each timestamp recorded. These results are detailed in \figurename \ref{fig:nodesFree} and \figurename \ref{fig:nodesCensor}.
-The distinct behaviors of each nation are rather striking. We can see a steep decline for some
-nations known to limit Internet freedom (China and Turkey) as well as the increase for other nations
-(Germany and France). Other nations have stayed more stable. It is worth pointing out that if a nation
-has required the same number of border ASes to control many paths over several years, this suggests
-that the chokepoints have grown to intercept more paths. The reason for this is that the number of ASes,
-and thus the number of paths, has increased globally over time. This fact makes the decline in ASes
-needed for some nations even more glaring.
+In order to compare multiple nations, we have plotted the number of ASes needed
+to intercept 90\% of in-to-out paths for each timestamp recorded. These results
+are detailed in \figurename \ref{fig:nodesFree} and \figurename
+\ref{fig:nodesCensor}.  The distinct behaviors of each nation are rather
+striking. We can see a steep decline for some nations known to limit Internet
+freedom (China and Turkey) as well as the increase for other nations (Germany
+and France). Other nations have stayed more stable. It is worth pointing out
+that if a nation has required the same number of border ASes to control many
+paths over several years, this suggests that the chokepoints have grown to
+intercept more paths. The reason for this is that the number of ASes, and thus
+the number of paths, has increased globally over time. This fact makes the
+decline in ASes needed for some nations even more glaring.
 
@@
 
-In our test of the relationship between Internet Freedom and national breakthrough potential, we plotted the Freedom on the Net score
-of each nation vs the number of border ASes that that nation needed to in order to intercept 90\% of in-to-out paths. Additionally,
-we evaluated the relationship with an Ordinarly Least Squares (OLS) fit, and found that the relationship was statistically significant,
-with a p-value $\leq$ 0.002 for 2017. The relationship held for other timestamps as well. The results for 2017 are shown in figure \ref{fig:fotn}
+In our test of the relationship between Internet Freedom and national
+breakthrough potential, we plotted the Freedom on the Net score of each nation
+vs the number of border ASes that that nation needed to in order to intercept
+90\% of in-to-out paths. Additionally, we evaluated the relationship with an
+Ordinarly Least Squares (OLS) fit, and found that the relationship was
+statistically significant, with a p-value $\leq$ 0.002 for 2017. The
+relationship held for other timestamps as well. The results for 2017 are shown
+in figure \ref{fig:fotn}
 
@@
 
-\par
-For each timestamp we see what appears to be two general modes of behavior. Nations that are not free or partly free tend to requrie few ASes
-to intercept 90\% of their paths. Free nations on the otherhand require varying degrees of large numbers of ASes to control the same portion
-of their paths. There are interesting outliers for both situations, however. Countries like Estonia (EE) or Iceland (IS) are very free but require
-few border ASes to control most of their paths. The reason for these outliers is likely that their overall AS counts are very low. On the other hand, Russia (RU)
-is a very interesting outlier in that it is found to be not free by FOTN, but requires a large number of ASes to control most of its paths. This suggests
-that the censorship efforts in Russia might be of types that do not require AS level chokepoints.
+For each timestamp we see what appears to be two general modes of behavior.
+Nations that are not free or partly free tend to requrie few ASes to intercept
+90\% of their paths. Free nations on the otherhand require varying degrees of
+large numbers of ASes to control the same portion of their paths. There are
+interesting outliers for both situations, however. Countries like Estonia (EE)
+or Iceland (IS) are very free but require few border ASes to control most of
+their paths. The reason for these outliers is likely that their overall AS
+counts are very low. On the other hand, Russia (RU) is a very interesting
+outlier in that it is found to be not free by FOTN, but requires a large number
+of ASes to control most of its paths. This suggests that the censorship efforts
+in Russia might be of types that do not require AS level chokepoints.

### 2018-07-30 Paper/conclusion.tex
Jed made some minor edits

@@
 The DNS infrastructure is another potential graph where chokepoints can be
-exploited for censorship and surveillance\~cite{Greschbach2016a} and is at a
+exploited for censorship and surveillance~\cite{Greschbach2016a} and is at a
 higher layer in the network stack than IP routes.  However, we leave analysis of

### 2018-07-30 Paper/experiment.tex
Start of results
Jed made some minor edits

@@
 \section{Experimental Setup}
+\label{sec:exp}
 
@@
 calculated the chokepoint potential for every border AS per snapshot using
-toolname{}. As a result we can investigate how countries have changed overtime
+toolname{}. As a result we can investigate how countries have changed over time
 in their capability to enact censorship and surveillance. This is an attempt to
 understand what topological trends have developed historically.  With this test
-we can compare nations, and see which ones have overtime increased their
+we can compare nations, and see which ones have, over time, increased their
 capability to control the flow of information across their borders.

### 2018-07-30 Paper/results.tex
Start of results
Jed made some minor edits

@@
 \section{Results}
+
+In this section we present the results of several experiments aimed at
+answering the questions presented in Section~\ref{sec:exp}.
+
+\Ben[inline]{We should answer the questions in the order we ask them. 1)
+National Chokepoint potential and implications. 2) Comparison among nations. 3)
+Evoluation over time.}
+
+\subsection{National Chokepoint Potential}
+
+\Ben[inline]{Include results here about general chokepoint potential,
+currently}.
+
 \subsection{Nations Over Time}
@@
 differences. First, it is worth pointing out that the United States has many
-more ASes than China, hence it's line extends further to the right in these
+more ASes than China, hence its line extends further to the right in these
 plots. We also see that China, in all cases, can control a much larger portion

### 2018-07-30 Paper/related.tex
Jed made some minor edits

@@
 chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} didn't consider that
+taken in this paper. First, the authors in \cite{throats} did not consider that
 many of the paths found from the Alexa top websites would have destinations in
@@
 to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study didn't include the United States and Russia in their experiments because they were
+to include more nations. Additionally, this study did not include the United States and Russia in their experiments because they were
 outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for

### 2018-07-30 Paper/intro.tex
writing tuneups of intro

@@
 
-\par Widespread Internet censorship and surveillance pose a
-significant threat to individuals throughout the world. As routine
+\par The Internet is rapidly changing from a collection of networks that span geo-political boundaries and promotes the free flow of information into a network that reflects political and economic constraints.
+%Internet censorship and surveillance pose a
+%significant threat to individuals throughout the world.
+As routine
 tasks, communications, entertainment, and information move online and
@@
 neutrality rules in the U.S.\ were recently overturned, and large
-content services are under enormous pressure to `do
-something' about fake news and bad actors. Also, other forms of Internet
+content services are under enormous pressure to control fake news and bad actors. Other forms of Internet
 tampering such as code injection~\cite{badtraffic,191996} have been observed within the
@@
 trends will likely restructure the Internet in unforeseen ways,
-as organizations respond to new challenges and realities.
+as organizations respond to new challenges and realities, particularly those imposed by legal and regulatory regimes.
 
-Today, most such control is exercised at the country level,
-as governments have recognized both the threat and the opportunity
-that is posed by ubiquitous online communication. For example, much
-of the organization, revolutionary momentum, and broadcasting of
-events during the Arab Spring have been attributed to social media
-communications through the Internet \cite{arabspring}. Several countries
-sought to control these movements by disconnecting in-country networks from the rest of the world \cite{BGPdisrupt}.
-These two factors, increasing levels of content monitoring and
-Internet dependence, point to the need for improved tools and methods
-for studying worldwide Internet censorship and surveillance over time.
-\todo[inline]{A SENTENC OR TWO HERE SUMMARIZING THE STATE OF THE FIELD AND WHAT'S
-  MISSING.}
+Today, most such control is exercised at the country level, as
+governments have recognized both the threat and the opportunity that
+is posed by ubiquitous online communication. For example, much of the
+organization, revolutionary momentum, and broadcasting of events
+during the Arab Spring have been attributed to social media
+communications through the Internet \cite{arabspring}, and several
+countries sought to control these movements by disconnecting
+in-country networks from the rest of the world \cite{BGPdisrupt}.
+These two factors, increasing dependence on the Internet and increased
+content monitoring, point to the need for improved tools and methods
+for studying worldwide Internet architecture over time.  Current
+methods have primarily focused on the node level (individual ASes)
+rather than the country level\Steph[inline]{Josh Karlin and Jen Rexford papers are two exceptions},
+they have not described how AS topology
+has changed over time with respect to national boundaries, and
+existing BGP simulation methods make it difficult to address 
+questions such as these efficienty and at scale.
 
@@
 Autonomous Systems are groups of routers under common management, such as the network of
-a university or an Internet Service Provider (ISP). The change in national AS topologies 
-is of interest to censorship research. The total size of the AS graph (The global network 
+a university or an Internet Service Provider (ISP).
+%The change in national AS topologies 
+%is of interest to censorship research.
+The total size of the AS graph (The global network 
 of all ASes) has grown from about 10,000 ASes in the early 2000s to over 60,000 today. 
@@
 
-\par Every nation has a different layout and number of ASes, and these national networks
-connect to foreign ASes in unique ways. The
-censorship and surveillance strategies of nations also differ. For
-instance, China both conducts keyword filtering in border ASes and in
-internal provincial ASes \cite{chinafiltering}, while Iran routes its
-Internet traffic through a centralized facility \cite{irancensor}.
-A picture of a nation's censorship or surveillance capabilities then
-is not complete without examining its underlying AS topology.
-Having the capability to investigate the AS topology of a nation, then, will both help
-researchers identify nations that could easily conduct censorship and
-also provide possible insight into what kind of censorship is likely
-being conducted.
+\par Every nation has a different layout and number of ASes, and these
+national networks connect to foreign ASes in unique ways. Considering
+just censorship and surveillance, different nations use different
+strategies for this form of content monitoring. For instance, China
+both conducts keyword filtering in border ASes and in internal
+provincial ASes \cite{chinafiltering}, while Iran routes its Internet
+traffic through a centralized facility \cite{irancensor}.  A picture
+of a nation's content monitoring capabilities then is not complete
+without examining its underlying AS topology.  Having the ability to
+investigate the AS topology of a nation will both help researchers
+identify nations that could, for example, easily conduct censorship
+and also provide possible insight into what kind of censorship is
+likely being conducted.  More generally, we are interested in the extent to which the Internet is changing from a borderless structure into one that reflects national boundaries.
 
-We examine how censorship and surveillance might be conducted 
-at the AS-level by evaluating border ASes, or ASes that can connect directly to at least
-one AS from another nation. We introduce a new measure called \emph{chokepoint potential} which
-captures the ability of a node to control communication to the Interent outside a nation's borders.
-We can then aggregate this measure and define \emph{national breakthrough potential} to
-measures the concentration of routing paths into and out of a country. 
-\Ben[inline]{These definitions are important. They need to be a concise and clear definitions. I don't think it's right yet, but we are close}
+%We examine how censorship and surveillance might be conducted 
+%at the AS-level by
+We focus on border ASes, or ASes that can connect directly to at least
+one AS from another nation and introduce a measure called \emph{chokepoint potential}, which quantifies the percentage of BGP paths into or out of a country which pass through that AS.
+At the country level, we are interested in aggregating chokepoint potential in different ways to quantify the tendency of all BGP paths crossing the border to pass through a small (or large) number of chokepoints (border ASes).
+We sometimes refer to this as \emph{national chokepoint potential}.
+%We can then aggregate this measure and define \emph{national breakthrough potential} to
+%measures the concentration of routing paths into and out of a country. 
+%\Ben[inline]{These definitions are important. They need to be a concise and clear definitions. I don't think it's right yet, but we are close}
 With these measures 
@@
 study how national chokepoint potential correlates with two independent
-measures of civil liberty, finding that significant relationship exists between NBP and both measures.
-The paper extends earlier research on AS topology, first by introducing the NBP measure, 
-then by presenting open-source cross-platform tools for simulating Border Gateway Protocol (BGP) paths,
-determining breakthrough potential at the country level, and analyzing how this measure has changed over time for certain countries of interest.
-Our measure of border ASes, chokepoint potential, defined in detail
-in section 3, is a succinct way to estimate the important properties of
-border ASes related to the ability of a nation to censor and surveil, and our aggregate extension of this measure to
-the national level, national breakthrough potential, is an effective way to evaluate and
-compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.\Ben[inline]{This paragraph seems a little redundant now.
+measures of civil liberty, finding that a significant relationship exists between national chokepoint potential and each measure.
+The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasetes for simulating Border Gateway Protocol (BGP) paths,
+determining national chokepoint potential, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
+%Our measure of border ASes, chokepoint potential, defined in detail
+%in section 3, is a succinct way to estimate the important properties of
+%border ASes related to the ability of a nation to censor and surveil, and our aggregate extension of this measure to
+%the national level, national breakthrough potential, is an effective way to evaluate and
+%compare nations and analyze the evolution of the AS-level censorship and surveillance landscape.\Ben[inline]{This paragraph seems a little redundant now.
 It is a combo of other previous writing, but I am not sure what to cut.}
@@
 \par
-The primary contributions of our work are as follows: 
+Specifically, the main contributions of the paper are: 
 \begin{enumerate}
- \item We introduce the measure of chokepoint potential for single border ASes and the 
- aggregate measure national breakthrough potential, 
-motivating their value as a way to interpret the capability for a 
-nation to enact censorship or surveillance of its Internet traffic.
- \item We provide an overview of the evolution of national AS-level chokepoints over time, 
- showing how the evolution of the Internet has either facilitated stronger
- AS-level chokepoints or lead to more open flow of information across borders.
- \item We develop a new tool, \toolname{}, that provides the capability for
-efficiently evaluating chokepoints for a given state of the Internet. 
- \item We show that national breakthrough potential
-has a significant relationship to Internet freedom as measured by two qualitative sources.
+ \item The chokepoint potential measure for single border ASes and the 
+ aggregate measure of national chokepoint potential, motivated as a way to measure the extent to which AS topology reflects national boundaries.
+
+ \item A study of how national AS-level chokepoints have changed over the past nine years, demonstrating that the Internet has evolved to facilitate stronger
+ AS-level chokepoints for some countries and lead to more open flow of information across borders for others.
+ \item An open-source tool, \toolname{}, and datasets for simulating BGP paths and evaluating chokepoints efficiently at different time points for the entire Internet.  
+ \item A study that reveals a significant relationship between chokepointpotential and Internet freedom, as measured by two qualitative sources.
+This result suggests that chokepoint potential can be interpreted as an indicator of a country's capability for conducting censorship or surveillance of its international Internet traffic.
 \end{enumerate}

### 2018-07-30 Paper/experiment.tex
Kirtus: Reworked Experimental Setup Section

@@
 
-In this section we use chokepoint potential to examine three questions: 1) to
+In this section we use chokepoint potential to examine three questions: 1) To
 what extent does the current state of the AS-level Internet topology support
-national governments' ability to censor and surveil Internet traffic? 2) How
-different is this ability vary across nations? and 3)  How has this ability
-evolved over time?
+national governments' ability to censor and surveil Internet traffic and how does
+this ability vary across nations? 2) How has this ability
+evolved over time? and 3) How does our topological chokepoint measure correlate to
+qualitative measures of Internet and press freedom?
 
-To probe these questions we first used our chokepoint evaluation technique to
-investigate the chokepoint potential of all nations for the current Internet as
-well as the change in chokepoint potentials over time. We looked at multiple
-snapshots of AS relationships from the CAIDA dataset (2012-2018).  For each
-timestamp, we generated routing trees based on these relationships. Then we
-calculated the chokepoint potential for every border AS per snapshot using
-toolname{}. As a result we can investigate how countries have changed over time
-in their capability to enact censorship and surveillance. This is an attempt to
-understand what topological trends have developed historically.  With this test
-we can compare nations, and see which ones have, over time, increased their
-capability to control the flow of information across their borders.
+Our experiments begin with the CAIDA AS relationship dataset for a particular timestamp.
+These detail inferred relationships between ASes, with which we can use our simulation framework
+to generate routing trees and then calculates the chokepoint potentials for each AS. We use these 
+relationships to identify border ASes, and we use the Team Cymru ASN lookup service to map each AS to the 
+country with which it is registered. \toolname{} returns both the set of routing trees built via simulation
+and the chokepoint potential that was determined for each AS. The resulting data allows us to conduct
+multiple levels of analysis. The resulting chokepoint potentials allow us to compare nations in regards
+to what kind of border topology they have. We can answer whether many or few ASes are needed for a 
+particular nation to intercept a majority of BGP paths, for instance. We can also compare the results
+of experiments at different timestamps to monitor the evolution of a nation's national chokepoint potential,
+indicating trends toward stronger chokepoints or, alteratively, more open borders.
 
 % Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can easily
-implement censorship, it stands to reason that their might already be a
-negative relationship between the chokepoint potential of a nation and its
-Internet freedom. If a significant relationship were to be found it would
-strengthen chokepoint potential as a measure of a nation's censorship
-capability and it would increase the value in monitoring chokepoint potentials
-across the globe.
+If chokepoint potential can be leveraged to determine if a nation can
+intercept many BGP paths with few ASes, it is possible that there is some correlation between
+this ability and the Internet freedom of that nation. If a higher national chokepoint potential
+for a nation is related to a less free Internet or press, this indicates that topology is linked
+to censorship in some way, perhaps indicating that the flow of information through border ASes
+is an important factor in interpreting a nations ability to censor their traffic. While a causal
+direction cannot be inferred by such a relationship, it would strengthen the importance of chokepoint
+potential as a measure of the evolutionary direction of the Internet.
 
 We tested whether a significant relationship exists between our measure of
-national chokepoints and a qualitative evaluation of Internet Freedom. For
+national chokepoints and two qualitative evaluations of communications Freedom. For
 Internet Freedom we used the Freedom House's Freedom On The Net report
@@
 ASes required to intercept 90\% of paths for each nation and viewing these
-rankings against the FOTN scores.
+rankings against the FOTN scores. FOTN only includes 65 nations in the most recent
+report (2017), so we also compare our measure with Freedom of the Press, which assessed
+201 nations and territories.
+
+% To probe these questions we first used our chokepoint evaluation technique to
+% investigate the national chokepoint potential of all nations for the current Internet as
+% well as the change in evolution of AS chokepoint potentials over time. We looked at multiple
+% snapshots of AS relationships from the CAIDA dataset (2009-2018).  For each
+% timestamp, we generated routing trees based on these relationships. Then we
+% calculated the chokepoint potential for every border AS per snapshot using
+% \toolname{}. As a result we can investigate how countries have changed over time
+% in their capability to enact censorship and surveillance. This is an attempt to
+% understand what topological trends have developed historically.  With this test
+% we can compare nations, and see which ones have, over time, increased their
+% capability to control the flow of information across their borders.
+
+%%%%%%

### 2018-07-30 Paper/background.tex
Added comment about needing censorship info in background
Clarified why BGP paths are different from AS graph paths

@@
 \par Finding BGP paths is more of
-a challenge.\Ben[inline]{Be more clear about the difference between `Finding BGP paths' and what you have discussed above.}
+a challenge. BGP paths are not simply the shortest paths available on the
+AS graph. Instead, they are determined by a combination of the dynamics of
+BGP updates between ASes and the local preferences of each AS to route traffic
+to destinations according to the policy of the AS administrator. 
 Packet based simulations, wherein BGP is simulated
@@
 paper) as part of \toolname{}.
+
+\Ben[inline]{We need to discuss not only the background of BGP routing here, but 
+also background about surveillance and censorship. We could probably shorten/tighten
+the BGP stuff to make room.}

### 2018-07-31 Paper/background.tex
tuned up writing and compressed the background.  SF doesn't think we need background on censorship since that is not our main focus and not required to understand the rest of the paper

--- 2018-07-26 Paper/background.tex
@@
 
-The AS-layer is the highest level of organization of the Internet, with each individual AS represents a separately administered network, for example, a large ISP, a university network, or a government entity.
-The Border Gateway Protocol (BGP) is used to route traffic among ASes \cite{bgp}.
-AS-level routes are known as \emph{paths} and are specified by routing tables stored locally in each AS.
-These tables express known connections to other ASes and local preferences for how data are routed, e.g., if there are multiple paths to a single destination, an AS may prefer the path that costs the least.
+% We already said this in the introductino
+%The AS-layer is the highest level of organization of the Internet, with each
+%individual AS represents a separately administered network, for example, a
+%large ISP, a university network, or a government entity.
+The Border Gateway
+Protocol (BGP) is used to route traffic among ASes \cite{bgp}.  AS-level routes
+are known as \emph{paths} and are specified by routing tables stored locally in
+each AS. These tables express known connections to other ASes and local
+preferences for how data are routed, e.g., if there are multiple paths to a
+single destination, an AS may prefer the path that costs the least.
 
@@
 information stored in the local routing tables, it is challenging to
-infer the exact AS topology at any point in time.  Several strategies have been developed to infer AS topology, but none of them is perfect.  Methods that rely
-on routing table information and BGP
-updates suffer from the fact that BGP paths stored in routing tables
-reflect only the current knowledge of a single AS about the routes
-available~\cite{BGPStream}.
+infer the exact AS topology at any point in time.  Several strategies
+have been developed, each with advantages and drawbacks. For example,
+methods based on empirical traceroute data are known to be incomplete
+or inaccurate~\cite{tracerouteProblems}. Similarly, methods that rely
+on routing table information reflect only the current knowledge of a
+single AS, or small set of ASes, about available
+routes~\cite{BGPStream}, and they rely on ASes that are willing to
+share their routing table information.  Many such ASes are required
 % FIXME: Not sure what BGPStream citation is for.  We need a different citation for RIPE unless that is what BGPStream is.
-Thus one would need to use many sources to infer an accurate
-picture of the BGP topology, which is not easily scalable for
-global studies, particularly those looking at multiple time periods.
-Similarly, methods based on empirical traceroute data are known to be incomplete or inaccurate~\cite{tracerouteProblems}.
+to infer a reliable picture of the BGP topology, an approach that does not scale well for global studies over time.
 
 \par AS-level studies are further complicated by the lack of ground
-truth data for AS relationships and BGP paths. For AS relationships,
-inference is often used based on economic considerations to classify
-the relationships of ASes. This was first done by Gao \cite{gao} by
-maximizing the occurances of certain economic rules on the AS graph by
-choosing a particular set or relationships. This technique was only
-evaluated against a single ISPs set of true relationships, however. As
-part of the CAIDA project, this inference technique was extended,
+truth data for AS relationships and BGP paths. Two ASes might enter a peering relationship,
+exchanging traffic but not announcing their connection. For AS relationships such as these, economic models are often used to classify
+the relationships of ASes. This approach was pioneered by Gao \cite{gao}, who
+assigned AS relationships that optimized certain economic rules on the AS graph.
+%This technique was only
+%evaluated against a single ISPs set of true relationships, however. As
+This inference technique was extended by the CAIDA project,
 leading to the CAIDA AS-relationship dataset \cite{CAIDApaper},
-\cite{CAIDA}. We choose these relationships for our purposes, and they
-are the current research standard.
+\cite{CAIDA}, which are the current research standard.
+%We choose these relationships for our purposes, and they
+%are the current research standard.
 
-\par Finding BGP paths is more of
-a challenge. Packet based simulations, wherein BGP is simulated
-directly, are computationally infeasible for the scales relevant to a
-global study. An accurate, but realistic, simulation technique must be
-chosen to provide useful research potential in this regard.  The
-BGPSim algorithm \cite{quicksand} is one such simulation technique
-that is suitable for this study's purposes. BGPSim takes a set of AS
-relationships as input, such as those provided by the CAIDA dataset
-\cite{CAIDA}, and returns routing trees based off of these
-relationships. These paths are found via a modified breadth-first
-seach (BFS) algorithm. The BFS adds edges to routing trees first
+\par Knowing the AS topology and AS relationships is critical for finding BGP paths, but it is not sufficient because BGP does not always adopt
+the shortest paths available on the
+AS graph. Instead, they are determined by a combination of the dynamics of
+BGP updates between ASes and the local preferences of each AS to route traffic
+to destinations according to local AS policies.
+%Packet based simulations, wherein BGP is simulated
+%directly, are computationally infeasible for the scales relevant to a
+%global study.
+BGPSim is a realistic, simulation technique developed to infer likely BGP paths.~\cite{quicksand}. BGPSim takes as input a set of AS
+relationships, such as those provided by the CAIDA dataset
+\cite{CAIDA}, and returns routing trees based on these
+relationships. A routing tree contains all
+equally reasonable paths (according to economic and connectivity constraints) that between source ASes (within the routing tree) to destination ASes (the
+root of the routing tree). These paths are found using a modified breadth-first
+search (BFS) algorithm. The BFS adds edges to routing trees first
 according to local preference (LP), then shortest path (SP), and
-finally tiebreak (TB).  A resulting routing tree contains all the
-equally reasonable paths (according to economic concerns) that exist
-between source ASes (within the routing tree) to destination ASes (the
-root of the routing tree). We find this technique suitable for our
-purposes, so we extend BGPSim (in ways explained further in this
-paper) as part of \toolname{}.
+finally tiebreak (TB).
+%We find this technique suitable \Ben[inline]{Why is it suitable?} for our
+%purposes.
+We adopt the basic algorithm of BGPSim
+%(in ways explained further in this
+%paper)
+as part of \toolname{}.
+
+\Ben[inline]{We need to discuss not only the background of BGP routing here, but 
+also background about surveillance and censorship. We could probably shorten/tighten
+the BGP stuff to make room.}

### 2018-07-31 Paper/intro.tex
tuned up writing and compressed the background.  SF doesn't think we need background on censorship since that is not our main focus and not required to understand the rest of the paper

@@
 content monitoring, point to the need for improved tools and methods
-for studying worldwide Internet architecture over time.  Current
+for studying worldwide Internet structure over time.  Current
 methods have primarily focused on the node level (individual ASes)
-rather than the country level\Steph[inline]{Josh Karlin and Jen Rexford papers are two exceptions},
+rather than the country level\Steph[inline]{Josh Karlin and Jen Rexford papers are two exceptions and maybe there are more}, 
 they have not described how AS topology

### 2018-07-31 Paper/chokepoint.tex
Some additions, edits

@@
 
-\Ben[inline]{Include a little paragraph here about what this section describes}.
+In this section we motivate the choice to focus our analysis on border ASes and then
+introduce and define out measure of chokepoint potential. We additionally introduce
+the aggregate measure of chokepoint potential, national chokepoint potential (NCP)
+for ranking and comparing nations according to the strength of their border AS chokepoints.
 
@@
 
-\subsection{Border Nodes}
+\subsection{Border ASes}
 

### 2018-07-31 Paper/intro.tex
Some additions, edits

@@
 content monitoring, point to the need for improved tools and methods
-for studying worldwide Internet structure over time.  Current
-methods have primarily focused on the node level (individual ASes)
-rather than the country level\Steph[inline]{Josh Karlin and Jen Rexford papers are two exceptions and maybe there are more}, 
-they have not described how AS topology
-has changed over time with respect to national boundaries, and
-existing BGP simulation methods make it difficult to address 
-questions such as these efficienty and at scale.
+for studying worldwide Internet structure over time. Previous research has
+investigated country level AS topology \cite{DBLP:journals/corr/abs-0903-3218},
+focusing on how the global AS graph layout indicates that interdomain routing
+for many Internet users depends on the infrastructure of other nations. We take
+a different approach by developing a technique to monitor the amount of paths intercepted
+by border ASes, allowing us to compare nations according to how BGP paths enter and exit
+each nation.
+% Current
+% methods have primarily focused on the node level (individual ASes)
+% rather than the country level, 
+% they have not described how AS topology
+% has changed over time with respect to national boundaries, and
+% existing BGP simulation methods make it difficult to address 
+% questions such as these efficienty and at scale.
 
@@
 from nation to nation whether from economic decision making, infrastructural necessities, 
-or efforts to build a powerful censorship and surveillance network.\Ben[inline]{Need citations for economic, infrastructure, censorship}.
+or efforts to build a powerful censorship and surveillance network \cite{economicPeering, irancensor}.
 %The Autonomous Systems (AS) layer of the Internet has grown and changed dramatically over its history.

### 2018-07-31 Paper/results.tex
Some additions, edits

@@
-\section{Results}
+\section{Experimental Results}
 
@@
 
+	To evaluate national chokepoint potentials of each of these nations we
+plot from our \toolname{} generated results the
+
 \subsection{Nations Over Time}
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{nationsCompare}
+	\caption{The national chokepoint potentials of various nations in 2018. k is the
+	percentage of paths used to calculate the national chokepoint potential.}\label{fig:nationsCompare}
+\end{figure}
 \begin{figure}
@@
 
-We have chosen to detail the results of our experiment for 8 nations due to
+We have chosen to detail the results of our national chokepoint potential
+experiment for 8 nations due to
 their histories of censorship practice. First we have chosen 4 nations known to

### 2018-07-31 Paper/chokepoint.tex
writing pass

@@
 need to define routing trees if we only really work on the path}  The
-chokepoint potential of a border AS is the ratio of paths intercepted by that
-border AS to the number of paths intercepted by all border ASes belonging to
-the same nation. Formally, for a country $c$, let $P = \{p_{i \rightsquigarrow
-j} : g(a_i)=c \text{ and } g(a_j)\neq c\} $, i.e.\ the set of all path
-originating in $c$ and ending in outside $c$. Then the chokepoint potential,
+chokepoint potential of a border AS is the ratio of the number of paths in (or out of) the country that contain the border AS to
+the total number of paths in (or out of) the same country.
+%intercepted by all border ASes belonging to
+%the same nation.
+Formally, for country $c$, let $P = \{p_{i \rightsquigarrow
+j} : g(a_i)=c \text{ and } g(a_j)\neq c\} $, i.e.\ the set of all paths
+originating in $c$ and ending outside $c$. Then the chokepoint potential,
 $cp(\cdot)$, of a border AS $a_u$ with $g(a_u) = c$ is
-
+\Steph[inline]{How do we count a parth that originates in a country, routes through a different country, and terminates in the original country?  That case is relevant in US law.}
 \begin{equation}
@@
 
-The above definition is specifically for capturing paths originating in country
-$c$ and routing outside the country, which we will refer to as
-\textit{outgoing} chokepoint potential or $cpo$. We can similarily define
-\textit{incoming} chokepoint potential or $cpi$, by simply redefining $P$ as $\{p_{i
+The above definition applies to paths originating in country
+$c$ and routing to a different country, which we will refer to as
+\textit{outgoing} chokepoint potential or $cpo$. We similarily define
+\textit{incoming} chokepoint potential or $cpi$, by redefining $P$ as $\{p_{i
 \rightsquigarrow j}: g(a_i) \neq c \text{ and } g(a_j) = c\}$. That is all
@@
 
-Given a set of routing trees\Ben[inline]{paths?}, chokepoint potential is an intuitive way to
-compare individual ASes. With this measure, the sum of the chokepoint potential
-for all border ASes for a country is 1.0, as in, all of the border ASes
-collectively control the flow of information over the nation's border. We
-define the aggregate \emph{national chokepoint potential} as the number of
+% SF says 'paths' in response to Ben
+Given a set of BGP paths, chokepoint potential is an intuitive way to
+compare individual ASes. Note that the sum of the chokepoint potentials
+for all border ASes for a given country is 1.0.
+%, as in, all of the border ASes
+%collectively control the flow of information over the nation's border.
+We define the aggregate \emph{national chokepoint potential} as the number of
 border ASes required to control a particular fraction $f$ of paths. Formally,
 given a country,$c$, with a set of $k$ border ASes $\{a_1,a_2,\cdots,a_k\}$ with
-$cp(a_i)>cp(a_j) \forall i<j$ (i.e.\ sorted in descending order), the
-we find the smallest $j$ such that
+$cp(a_i)>cp(a_j) \forall i<j$ (i.e.\ sorted in descending order), we find the smallest $j$ such that
 
@@
 
-and define national chokepoint potential as $CP(c,f) = 1/j$. We use the
-multiplicative inverse here to preserve a clearer semantic meaning i.e.\ higher
-chokepoint potential implies more control over paths. The more border ASes
-required, the more difficult it would likely be for that nation to perform
-censorship or surveillance, as the government would have to restrict more ISPs,
-place more filtering hardware, etc. 
+and define national chokepoint potential as $CP(c,f) = 1/j$. The
+multiplicative inverse preserves a clearer semantic meaning i.e.\ higher
+chokepoint potential implies more control over paths. The fewer border ASes
+required to control $f$ paths, the easier it is to perform
+censorship or surveillance, e.g., by restricting ISPs,
+placing filtering hardware, etc. 
 %Thus, we can define another measure, with
@@
 %selected percentage of AS-level paths. 
-Previous work has used $f=0.90$ as a benchmark indicating strong control of information
+Previous work used $f=0.90$ as a threshold indicating strong control of information
 \cite{throats}, so we use this percentage as well. When used here, however,
-90\% path interception indicates a single nation intercepts 90\% of the paths
-that either enter said nation (for the out-to-in case) or exit said nation (for
-the in-to-out case). National breakthrough potential is calculated taking the
-strongest ASes first, such that the measure reflects the smallest number of
-border ASes needed to intercept the paths.\footnote{We note that a dual measure
+90\% path control indicates the percentage of paths into (our out of) a country.
+%that a single nation intercepts 90\% of the paths
+%that either enter said nation (for the out-to-in case) or exit said nation (for
+%the in-to-out case).
+National chokepoint potential is calculated
+%by taking the
+%strongest ASes first, such that the measure
+as the smallest number of
+border ASes needed to control $f$ of the paths.\footnote{We note that a dual measure
 could be defined as the fraction of paths controlled by $j$ ASes. Our results
@@
 We measure the absolute number of ASes required to control a specific
-fraction rather than a percentage within a country because it makes for more
-equitable comparisons between nations
-Take for instance the United States, which has many more ASes than
+fraction rather than a percentage within a country to make
+comparisons between nations more equitable.
+Consider, for instance, the United States, which has many more ASes than
 most nations. If the United States and another nation both required the same
 percentage of border ASes to intercept 90\% of paths, the similarity of these
-nations would be misleading. Controlling the larger absolute number of ASes
-would require significantly more resources.\Ben[inline]{we need to actually
+nations would be misleading because the cost of controlling the larger absolute number of ASes
+would be significantly higher.\Ben[inline]{we need to actually
 pick another nation where this is true to make this example illustrative.} Our technique provides a comparison more likely to

### 2018-07-31 Paper/tool.tex
Jed made some minor edits
writing pass on tool section.  Ben/Kirtus: Can we do a back of the envelope coplexity analysis for the calculations?

@@
 
-In this section we describe \toolname{}, its relationship to previous attempts
-at identifying routing paths, and how it is used to calculate chokepoint
+This section describes \toolname{}, how it relates to earlier work, and how it is used to calculate chokepoint
 potential.
@@
 % Preprocessing and Data pipeline
-\toolname{} is a new tool that simluates plausible BGP paths and calculates AS
-chokepoint potential and national chokepoint potential.  \toolname{} uses
-readily available AS relationship data, e.g. data made available from
-CAIDA~\cite{CAIDA}, and uses a similar approach to BGPSim \cite{quicksand} to
+\toolname{} determines a set of plausible BGP paths through simulation and calculates AS
+chokepoint potential and national chokepoint potential.  \toolname{} takes as input AS relationship data, such as that provided by
+CAIDA~\cite{CAIDA}.  It also takes as input a set of AS
+country codes (that identifies which nation an AS belongs to) for
+determining chokepoint potential.  \toolname{}
+uses an algorithm based on BGPSim \cite{quicksand} to
 generate a set of routing trees. \Ben[inline]{Need a short description of how
-BGPSim works, maybe below.}. These routing trees, as well as a set of AS
-country codes (for identifying which nation an AS belongs to) are used to
-determine the chokepoint potentials border ASes.  Finally, \toolname{}
-calclates the national breakthrough potential for each nation. In our
-experiments, we used the country codes returned from Team Cymru's IP to ASN
-whois service \cite{cymru} to determine which ASes were registered to which
-nation.  This service only maintains the most recent registration of each AS,
-so to control for uncertain country to AS links we keep ASes that have been
-registered more recently than the current timestamp being tested in the path
-simulation but we do not assign these ASes chokepoint
-potentials. We chose this
-route because the datasets of AS registrations for the various ASN registries
-tended to have more missing data at specific timestamps.  Fortunately, ASes
-with registration changes after each test date tend to not be in the AS graph
-for that test. \Ben[inline]{Two things. 1) This is sort of confusing, see if you
-can clarify it. 2) Is this a big deal? How often does this happen? If it
-happens frequently do we expect it to affect our results? }
+BGPSim works, maybe below.}. \toolname{} 
+calclates both AS and national chokepoint potential for any country.
+
+Our experiments used country codes returned from Team Cymru's IP to
+ASN whois service \cite{cymru} to determine which ASes were registered
+to which nation.  This service maintains only the most recent
+registration of each AS.  To control for this possible source of
+error, we retain ASes in thepath simulation, even if they were
+registered more recently than the current timestamp being studied, but
+we do not assign these ASes chokepoint potentials. We took this approach
+because the datasets of AS registrations for the various ASN
+registries tend to have more missing data at specific timestamps.
+Fortunately, ASes with registration changes after each test date tend
+to not be in the AS graph for that test. \Ben[inline]{Two things. 1)
+  This is sort of confusing, see if you can clarify it. 2) Is this a
+  big deal? How often does this happen? If it happens frequently do we
+  expect it to affect our results? }
 
 % Routing Tree Algorithm
-\par In order to calculate the routing trees, we use an extended version of the
-BGPSim algorithm developed by Gil et. al in \cite{quicksand}. In our work we
-addressed the following limitations of BGPSim: (1) BGPSim returns a set of ASes
-for each path it considers but not the order in which they are visited; (2)
-Once routing trees are determined, they cannot be accessed later without
-recalculation; (3) BGPSim relies on the outdated parallelization framework
-DryadLinq for C\#. To address these issues, we implemented an our routing
-simulation tool in Python. \Ben[inline]{Why does BGPSim not return ordered
-paths, this seems trivial, but I don't know the implementation details, was it
-just an oversight or did you do something cool?}\toolname{} returns ordered paths from its routing trees, saves
-routing trees to disk after calculation, and is parallelized with MPI via the
-mpi4py library. These improvements have the added benefit of yielding a cross
-platform routing tree algorithm that is ready to work on most hardware.
+\par To calculate the routing trees, we use an extended version of the
+BGPSim algorithm developed by Gil et. al in \cite{quicksand}. This
+component of \toolname{} addresses several limitations that prevented
+us using BGPSim directly: (1) BGPSim returns a set of ASes for each
+path it considers but not the order in which they are visited; (2)
+Once routing trees are determined, they cannot be accessed later
+without recalculation; (3) BGPSim relies on the outdated
+parallelization framework DryadLinq for C\#. To address these issues,
+we implemented our routing simulation in
+Python. \Ben[inline]{Why does BGPSim not return ordered paths, this
+  seems trivial, but I don't know the implementation details, was it
+  just an oversight or did you do something cool?}\toolname{} returns
+ordered paths from its routing trees, saves routing trees to disk
+after calculation, and is parallelized with MPI via the mpi4py
+library. These improvements have the added benefit of yielding a cross
+platform routing tree algorithm that is portable to most
+hardware.
+
+\Steph[inline]{Where is the code available?  What about the dataset of routing trees?  that is one of our big contributions.}
 
@@
 \par Once \toolname{} generates the routing trees, they can be processed to
-determine chokepoint potentials. This is done by iterating over every path
-between each AS-pair. Because we use the same random tie-break as BGPSim, there
-is only 1 path between each AS-pair considered, even if multiple exist in the
+determine chokepoint potentials or by other researchers interested in other questions. To calculate chokepoint potential, \toolname{} iterates over every path
+between each AS-pair. Because we use the same random tie-break method as BGPSim,this process returns 
+exactly one path between each AS-pair considered, even if multiple options exist in the
 routing tree.\Ben[inline]{Put stuff in about tiebreaking being rare and
-unlikely to affect results} Once a path is determined, it is traversed. For each AS visited,
-the number of paths intercepted by that AS is incremented. This is done for
-both outgoing paths and incoming paths, so only one traversal is necessary
+unlikely to affect results} Once the path has been determined, it is traversed to identify border nodes and increment their counts, both
+for outgoing and incoming paths.  Thus, only one traversal is conducted
 per path. Additionally, the number of paths of each type that belong to each
-nation is tallied, as this makes up the denominator in equation
+nation is tallied, which becomes the denominator in equation
 \ref{eqn:chokePointPotential}.  \toolname{} takes the resulting chokepoint
-potentials and generates national breakthrough potentials for each nation.
+potentials and generates national chokepoint potentials for each nation.
+
+\Steph[inline]{A brief complexity analysis, even just in words, would help, as woulddata on how long the computations take.  Complexity discussion here, timing data in Experiments?}

### 2018-07-31 Paper/tool.tex
added complexity analysis to 'tool.tex'

@@
 platform routing tree algorithm that is portable to most
-hardware.
-
-\Steph[inline]{Where is the code available?  What about the dataset of routing trees?  that is one of our big contributions.}
+hardware. The source code for \toolname{} and all datasets are publicly available at [URL blinded for peer review].
 
@@
 
-\Steph[inline]{A brief complexity analysis, even just in words, would help, as woulddata on how long the computations take.  Complexity discussion here, timing data in Experiments?}
+Calculating routing trees for the entire AS graph takes time proportional to the $|V|^2$ where $|V|$ is the total number of ASes; calculating chokepoint potential takes $|V|^2D$ where $D$ is the average depth of a routing tree.
+For example,
+it took about 12 wall clock hours to compute all of the routing trees for the Jan.
+2017 dataset, running on 4 Intel Xeon E5-2680 V4 CPUs.
+For each dataset, this is a one-time cost because the routing trees are stored externally.
+[URL blineded for peer review]

### 2018-07-31 Paper/related.tex
Jed added a few citations and put a note in the the related works for them
Jed added a paragraph at the end of related.tex

@@
 nations that could easily implement censorship if their governments decided to.
+
+Routing and its interplay with Internet censorship, surveillance, and related
+issues is a general research area with a broad set of research questions.
+Karlin \emph{et al.}~\cite{DBLP:journals/corr/abs-0903-3218} considered the
+centrality of countries with respect to routing of other nations' traffic, a
+related problem that is distinct from chokepoints to monitor/control traffic
+into and out of a government's own country.  Dainotti \emph{et
+al.}~\cite{Dainotti:2011:ACI:2068816.2068818} anlyze two specific large-scale
+Internet disruptions at the routing level.  Khattak \emph{et
+al.}~\cite{Khattak:2014:LCI:2663716.2663750} performed a detailed analysis of
+censorship at the ISP level in Pakistan.

### 2018-07-31 Paper/experiment.tex
writing pass on experiment.tex

@@
 
-In this section we use chokepoint potential to examine three questions: 1) To
-what extent does the current state of the AS-level Internet topology support
-national governments' ability to censor and surveil Internet traffic and how does
-this ability vary across nations? 2) How has this ability
-evolved over time? and 3) How does our topological chokepoint measure correlate to
+This section describes how we used \toolname{} to study three questions related to national borders and the BGP network:
+(1) To
+what extent do BGP networks today reflect national boundaries
+%support
+%national governments' ability to censor and surveil Internet traffic and
+how do they vary across nations? (2) How has this structure evolved
+over time? and (3) How do topological chokepoints correlate with
 qualitative measures of Internet and press freedom?
 
-Our experiments begin with the CAIDA AS relationship dataset for a particular timestamp.
-These detail inferred relationships between ASes, with which we can use our simulation framework
-to generate routing trees and then calculates the chokepoint potentials for each AS. We use these 
-relationships to identify border ASes, and we use the Team Cymru ASN lookup service to map each AS to the 
-country with which it is registered. \toolname{} returns both the set of routing trees built via simulation
-and the chokepoint potential that was determined for each AS. The resulting data allows us to conduct
-multiple levels of analysis. The resulting chokepoint potentials allow us to compare nations in regards
-to what kind of border topology they have. We can answer whether many or few ASes are needed for a 
-particular nation to intercept a majority of BGP paths, for instance. We can also compare the results
-of experiments at different timestamps to monitor the evolution of a nation's national chokepoint potential,
-indicating trends toward stronger chokepoints or, alteratively, more open borders.
+All of our experiments used the relevant CAIDA AS relationship dataset for 
+each timestamp we studied.  These contain inferred relationships
+between ASes, which the simulation framework uses to
+generate routing trees and  calculate chokepoint potentials.
+We use the relationships to identify border ASes, and
+we use the Team Cymru ASN lookup service to map each AS to the country
+with which it is registered. \toolname{} returns both the set of
+routing trees built via simulation and the chokepoint potential that
+was determined for each AS. These data form the basis for
+multiple levels of analysis: comparing chokepoint potential between countries,
+assessing how many ASes are required to control most international BGP paths, 
+and comparing results across multiple time points.
+%timestamps to monitor the evolution of a nation's national chokepoint
+%potential, indicating trends toward stronger chokepoints or,
+%alteratively, more open borders.
 
 % Chokepoints and Internet Freedom
-If chokepoint potential can be leveraged to determine if a nation can
-intercept many BGP paths with few ASes, it is possible that there is some correlation between
-this ability and the Internet freedom of that nation. If a higher national chokepoint potential
-for a nation is related to a less free Internet or press, this indicates that topology is linked
-to censorship in some way, perhaps indicating that the flow of information through border ASes
-is an important factor in interpreting a nations ability to censor their traffic. While a causal
-direction cannot be inferred by such a relationship, it would strengthen the importance of chokepoint
-potential as a measure of the evolutionary direction of the Internet.
-
-We tested whether a significant relationship exists between our measure of
-national chokepoints and two qualitative evaluations of communications Freedom. For
-Internet Freedom we used the Freedom House's Freedom On The Net report
-\cite{FOTN}. FOTN scores quantify the level of Internet freedom in countries.
-Each country receives a numerical score from 0 (the most free) to 100 (the
-least free). We compare our results to the FOTN scores by ranking nations
-according to national breakthrough potential, particulary the number of border
-ASes required to intercept 90\% of paths for each nation and viewing these
-rankings against the FOTN scores. FOTN only includes 65 nations in the most recent
-report (2017), so we also compare our measure with Freedom of the Press, which assessed
-201 nations and territories.
+Chokepoints provide one leverage point for nations interested in censorship or surveillance.
+%potential can be leveraged to determine if a nation can
+%intercept many BGP paths with few ASes, it is possible that there is some correlation between
+Thus, we are interested in the extent to which the Internet freedom of a country correlates with its chokepoint potential.
+%If a higher national chokepoint potential
+%for a nation is related to a less free Internet or press, this indicates that topology is linked
+%to censorship in some way, perhaps indicating that the flow of information through border ASes
+%is an important factor in interpreting a nations ability to censor their traffic.
+Although correlation certainly does not imply a causal relation, correlations can highlight overall trends and even provide clues about how censorship might be implemented in countries that are known to control content.
+%direction cannot be inferred by such a relationship, it would strengthen the importance of chokepoint
+%potential as a measure of the evolutionary direction of the Internet.
+We tested the statistical relationship between national chokepoint potential
+and two relevant qualitative evaluations of freedom. First, we used
+Freedom House's Freedom On The Net report \cite{FOTN}. FOTN scores
+quantify the level of Internet freedom in countries.  Each country
+receives a numerical score from 0 (the most free) to 100 (the least
+free). We compare our results to FOTN scores (for years 2014-2017) by ranking nations
+according to the number of border ASes required to intercept 90\% of
+paths for each nation.  FOTN includes only 65 nations in the most
+recent report (2017), so we also compare to the Freedom of
+the Press evaluation (2014-2017), which assesses 201 nations and territories.
 

### 2018-07-31 Paper/results.tex
National Chokepoint Results

@@
 
-\Ben[inline]{We should answer the questions in the order we ask them. 1)
-National Chokepoint potential and implications. 2) Comparison among nations. 3)
-Evoluation over time.}
-
 \subsection{National Chokepoint Potential}
 
-\Ben[inline]{Include results here about general chokepoint potential,
-currently}.
+	We have chosen to detail the results of our national chokepoint potential
+experiment for 8 nations due to
+their histories of censorship practice. First we have chosen 4 nations known to
+enact extensive Internet censorship policies. These nations are China, Turkey,
+Egypt, and Russia. China has historically been the number one nation in regards
+to Internet censorship \cite{censorshipSurvey}. Turkey has recently
+dramatically increased its online censorship efforts, by blocking social media
+accesses during turmoil and pressuring ISPs to block access to material chosen
+by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered
+severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
+The Russian government has passed surveillance laws requiring ISPs to allow the
+Russian government access to user statistics \cite{censorshipGeography}. Next
+we have chosen the nations of Germany, the United States, France, and the
+United Kingdom because, while they have all conducted limited forms of Internet
+censorship \cite{censorshipSurvey}, Internet access for these nations is
+considered somewhat open.
 
-	To evaluate national chokepoint potentials of each of these nations we
-plot from our \toolname{} generated results the
+We present the national chokepoint potentials for these nations in 2018 with increasing
+$f$ in \figurename \ref{fig:nationsCompare}. For each value of $f$, (on the x-axis),
+the naitonal chokepoint potential, the inverse of the number of ASes needed to intercept
+the fraction $f$ of paths is plotted. For all results we detail only the in-to-out results
+because we find little difference between the results for in-to-out paths and out-to-in paths
+for all our experiments.
 
-\subsection{Nations Over Time}
+\figurename \ref{fig:nationsCompare} shows us the diverse nature of national topologies. China is one
+extreme example, where upto more than 80 percent of paths can be intercepted by a single AS, as indicated
+by a NCP of 1.0. Other nations exhibiting strong chokepoints are Egypt, Turkey, and the United Kingdom. The
+nations of the US, France, Germany, and Russia show considerably lower NCP for 2018.
 
-\begin{figure}
+\begin{figure}[ht]
 	\centering
@@
 \end{figure}
+
+\subsection{Nations Over Time}
+
 \begin{figure}
@@
 \end{figure*}
-
-We have chosen to detail the results of our national chokepoint potential
-experiment for 8 nations due to
-their histories of censorship practice. First we have chosen 4 nations known to
-enact extensive Internet censorship policies. These nations are China, Turkey,
-Egypt, and Russia. China has historically been the number one nation in regards
-to Internet censorship \cite{censorshipSurvey}. Turkey has recently
-dramatically increased its online censorship efforts, by blocking social media
-accesses during turmoil and pressuring ISPs to block access to material chosen
-by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered
-severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the
-Russian government access to user statistics \cite{censorshipGeography}. Next
-we have chosen the nations of Germany, the United States, France, and the
-United Kingdom because, while they have all conducted limited forms of Internet
-censorship \cite{censorshipSurvey}, Internet access for these nations is
-considered highly open.
 

### 2018-07-31 Paper/background.tex
Jed added a reference and a paragraph at the end of background.tex

@@
 the BGP stuff to make room.}
+
+The interplay between the Internet and policy has a history as long as the
+Internet itself, but in terms of national-scale interference with packets the
+first prominent case of a state-sponsored system to achieve this is the
+so-called ``Great Firewall of China,'' which was first documented in the
+technical research literature in 2006 by Clayton \emph{et
+al.}~\cite{Clayton:2006:IGF:2166520.2166523}.  As of 2017, 49 countries are
+reported by the Freedom on the Net project~\cite{FOTN} to be ``partly free'' or
+``not free''.  Large-scale Internet surveillance is also prominent and has a
+long history, but surveillance is not as easy to measure as censorship so only
+information dumps such as the 2013 Snowden revelations can give us a sense of
+the prevalence and time periods related to surveillance.  National-scale
+manipulation of packets for reasons other than censorship, for example to inject
+malware or carry out distributed denial-of-service attacks, was first seen being
+carried out by China's ``Great Cannon'' in 2015~\cite{badtraffic}.  More
+recently Egypt was documented doing similar injections using a commercial
+product~\cite{191996}.  This phenomena is more targeted and therefore more
+stealthy than national-scale censorship, so may be more prevalent than just
+these two countries.

### 2018-07-31 Paper/conclusion.tex
Seperated conclusion

@@
-\section{Conclusions and Discussion}
-
-% Chokepoint Potential
-In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
-and national breakthrough potential, on entire nations, that together allow us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating AS-level topology using
-\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
-We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
-% Ecolution over time
-\par
-Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
-interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
-more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
-we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
-% Internet Freedom
-\par
-We have validated national breakthrough potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
-by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
-at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.
+\section{Discussion}
 
 \subsection{Routing Trees Dataset}
-In the hope to further AS topology research, we have open sourced the routing tree datasets generated in this study. While we generated routing trees
-with an efficient algorithm, it still requires considerable time to calculate them, particulary for multiple timestamps. Additionally, each set of routing trees
-takes up on the order of 50GB of disk space. By releasing these datasets we hope that researchers looking for a particular set of routing trees will find
-working with these simpler than recalculating them. This also provides an alternative to research projects that might otherwise use measurement tools to estimate
-BGP paths.
+Despite the use of efficient simulation techniques, collecting BGP path data is still not a simple task.
+For a global, multiple timestamp study such as this one, compute resources can be a limiting factor to
+researchers without access to a compute cluster.
+In the hope to further AS topology research, we have open sourced the routing tree datasets produced for
+this study, available at [removed for peer review] generated in this study. By releasing these datasets we 
+hope that researchers looking for a particular set of routing trees will find working with these simpler 
+than recalculating them, or creating new datasets by extensive measurement or
+inference.
 
@@
 censorship events and chokepoint potential changes, and as such lies beyond the scope of this study.
+
+\section{Conclusion}
+% Chokepoint Potential
+In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
+and national breakthrough potential, on entire nations, that together allow us to
+evaluate the level of difficulty that a government faces when trying to censor the Internet
+based entirely on AS-level topology. Our technique for evaluating AS-level topology using
+\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
+We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
+design principles so that this tool is readily deployable for future research.
+% Ecolution over time
+\par
+Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
+interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
+more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
+we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
+% Internet Freedom
+\par
+We have validated national breakthrough potential as having a relationship to actual censorship by showing
+the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
+by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
+at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.

### 2018-07-31 Paper/related.tex
Jed added a paragraph to the end of related.tex

@@
 censorship at the ISP level in Pakistan.
+
+While there have been attempts to systematize different technical approaches to
+censoring Internet content~\cite{Tschantz2016a,Khattak2016a} there is no
+comprehensive list of all the different ways a state actor can manipulate
+traffic.  For some methods the AS graph is relevant, such as IP address
+blacklisting, traffic throttling~\cite{DBLP:journals/corr/Anderson13}, URL or
+packet filtering, packet injection, physically shutting down infrastructure, and
+BGP attacks.  For other methods the AS graph is less relevant, such as creating
+internal national networks~\cite{DBLP:journals/corr/abs-1209-6398}, portal
+censorship such as search engine filtering and social media post deletion,
+propaganda campaigns, and manipulation of the DNS system.  Our work is
+complementary to other studies that focus more on Internet routing's effect on
+high-level content than on a country's desire to enforce policies on all traffic
+that crosses their borders, such as Edmundson \emph{et
+al.}~\cite{Edmundson:2018:NHI:3209811.3211887}.

### 2018-07-31 Paper/results.tex
fixed a caption

--- 2018-07-30 Paper/results.tex
@@
-\section{Results}
+\section{Experimental Results}
 
-In this section we present the results of several experiments aimed at
-answering the questions presented in Section~\ref{sec:exp}.
+This section reports results for experiments aimed at
+answering the three research questions described in Section~\ref{sec:exp}.
 
-\Ben[inline]{We should answer the questions in the order we ask them. 1)
-National Chokepoint potential and implications. 2) Comparison among nations. 3)
-Evoluation over time.}
+\subsection{National Chokepoint Potential (NCP)}
 
-\subsection{National Chokepoint Potential}
+We report results for eight countries, some that are known to practice censorship and some that are considered more open.
+We selected four well-known countries with
+extensive Internet censorship policies: China, Turkey,
+Egypt, and Russia. China has historically been the most aggressive in terms of
+Internet censorship \cite{censorshipSurvey}. Turkey recently
+began blocking social media
+accesses during turmoil and pressuring ISPs to block access to certain material~\cite{turkeyCensor}. Egypt's Internet suffered
+severe shutdowns during the Arab Spring \cite{arabspring}.
+Russian has passed surveillance laws requiring ISPs to provide access to user statistics \cite{censorshipGeography}. 
+We also selected Germany, the United States, France, and the
+United Kingdom because Internet access for these nations is
+considered to be open even though they each have conducted limited Internet
+censorship \cite{censorshipSurvey},
 
-\Ben[inline]{Include results here about general chokepoint potential,
-currently}.
+\figurename~\ref{fig:nationsCompare} shows national chokepoint
+potential for 2018 for different values of $f$ (x-axis). The
+y-axis reports the inverse of the number of ASes required to control fraction $f$ of
+paths. We report only in-to-out results
+because the results for out-to-in are very similar
+for all experiments.  \figurename~\ref{fig:nationsCompare} shows the diverse nature of national topologies. China is one
+extreme, where over 80\% of paths can be controlled by a single AS,
+%as indicated
+%by a NCP of 1.0.
+Other nations exhibiting strong chokepoints are Egypt, Turkey, and the United Kingdom. The
+nations of the US, France, Germany, and Russia show considerably lower NCP for 2018.
+These results are as expected, except for Russia.
+As an early adopter of the Internet, we speculate that Russia's AS topology evolved before stringent regulations were adopted and this early history dominates our measurements.
 
-\subsection{Nations Over Time}
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{nationsCompare}
+	\caption{The national chokepoint potentials of various nations in 2018. f is the
+	fraction of paths used to calculate the national chokepoint potential.}\label{fig:nationsCompare}
+\end{figure}
+
+\subsection{Topology Over Time}
 
@@
 	\includegraphics[width=\textwidth]{multi}
-	\caption{The capability for various nations to intercept in-to-out paths over multiple years \timerange. Each plot shows for a number of border ASes controlled (x-axis),
-	what fraction of in-to-out paths are intercepted (y-axis). Years in blue are more recent, and years in red are further in the past.}\label{fig:multiTime}
+	\caption{Chokepoint potential (in-to-out) for multiple years
+          \timerange. Each plot shows the number of border ASes
+          controlled (x-axis) vs. $f$ (y-axis).}\label{fig:multiTime}
 \end{figure*}
 
-We have chosen to detail the results of our experiment for 8 nations due to
-their histories of censorship practice. First we have chosen 4 nations known to
-enact extensive Internet censorship policies. These nations are China, Turkey,
-Egypt, and Russia. China has historically been the number one nation in regards
-to Internet censorship \cite{censorshipSurvey}. Turkey has recently
-dramatically increased its online censorship efforts, by blocking social media
-accesses during turmoil and pressuring ISPs to block access to material chosen
-by the Turkish government \cite{turkeyCensor}. The Internet of Egypt suffered
-severe shutdowns during the Arab Spring of the early 2010s \cite{arabspring}.
-The Russian government has passed surveillance laws requiring ISPs to allow the
-Russian government access to user statistics \cite{censorshipGeography}. Next
-we have chosen the nations of Germany, the United States, France, and the
-United Kingdom because, while they have all conducted limited forms of Internet
-censorship \cite{censorshipSurvey}, Internet access for these nations is
-considered highly open.
+To study changes over time, we first plotted the
+number of border ASes required to intercept 90\% of in-to-out paths
+for each year (\figurename \ref{fig:nodesFree} and \figurename
+\ref{fig:nodesCensor}).  The results are striking with a steep decline
+over time for some nations (China and Turkey) and increases for others
+(Germany and France), while some countries remained relatively
+stable. Given that the overall number of ASes and paths has increased
+globally over time, it is likely that even nations that look stable
+have in fact become more 'chokepointy' over time as some border nodes
+must be intercepting more paths.  This fact makes the decline observed
+for some countries even more striking.
 
-To inspect the changes over time for each nation we arrange all the border ASes
-belonging to a nation into a list. The list of border ASes is reverse sorted so
-that the first AS has the highest chokepoint potential. The sum of all these
-chokepoint potentials is 1.0. Then we step through the list, and record the
-cumulative chokepoint potential of all the ASes seen so far. If we plot the
-number of ASes controled vs the cumulative chokepoint potential for that number
-of ASes, we can see how many ASes are required for a certain nation to control
-different percentages of paths. We repeat this process for each snapshot to
-highlight changes over time.
+Next, for each country studied we sorted its border ASes so
+that the first AS has the highest chokepoint potential.
+%The sum of all these
+%chokepoint potentials is 1.0.
+Then, we calculated the cumulative chokepoint potential
+%we step through the list, and record the
+%cumulative chokepoint potential of all the ASes seen so far. If we
+and plotted the
+number of border ASes controlled vs cumulative chokepoint potential for each year in the study,
+%This shows how many ASes are required to control a given
+%percentage of paths.
+as shown in
+\figurename \ref{fig:multiTime}.  The x-axis (log-scale,
+so countries with large differences can be compared more
+easily) for each plot is the number of border ASes controlled, and the y-axis
+is the fraction of in-to-out paths intercepted. Each line represents a different
+year.
+%, with the more red lines being farther in the past and more blue lines
+%being more recent.
 
-For instance, consider \figurename \ref{fig:multiTime}.  The x-axis (log-scale,
-so that countries with large differences in AS counts can be compared more
-easily) for each plot is the number of border ASes controlled, and the y-axis
-is the ratio of in-to-out paths intercepted. Each line represents a different
-snapshot, with the more red lines being farther in the past and more blue lines
-being more recent.
-
-The United States and China are shown here and exhibit some dramatic
-differences. First, it is worth pointing out that the United States has many
-more ASes than China, hence its line extends further to the right in these
-plots. We also see that China, in all cases, can control a much larger portion
-of its paths with much fewer ASes than the United States. This result is
-entirely expected. Of more interest is the trends that can be seen over time.
-The AS-level topology of China's Internet has evolved such that very few ASes
-can intercept nearly all BGP paths. The US has evolved in a different way.
-While it has become somewhat easier for the US to intercept most of its paths,
-it has become more difficult to intercept around 70\% of its paths and higher.
-This suggest an expansion of ASes on the interior and more connections, as well
-as a strengthening of the very top ASes.
-
-Not all nations have evolved to a state where it easier to control paths. One
-example is Germany, as shown in \figurename \ref{fig:multiTime}. For Germany, a
-fairly constant trend shows that it has become more difficult to intercept BGP
-paths. Unlike the other examples, any amount of German border ASes intercept
-less paths in more recent tests.
-
-In order to compare multiple nations, we have plotted the number of ASes needed
-to intercept 90\% of in-to-out paths for each timestamp recorded. These results
-are detailed in \figurename \ref{fig:nodesFree} and \figurename
-\ref{fig:nodesCensor}.  The distinct behaviors of each nation are rather
-striking. We can see a steep decline for some nations known to limit Internet
-freedom (China and Turkey) as well as the increase for other nations (Germany
-and France). Other nations have stayed more stable. It is worth pointing out
-that if a nation has required the same number of border ASes to control many
-paths over several years, this suggests that the chokepoints have grown to
-intercept more paths. The reason for this is that the number of ASes, and thus
-the number of paths, has increased globally over time. This fact makes the
-decline in ASes needed for some nations even more glaring.
+The U.S. and China exhibit dramatic differences. The U.S. has many
+more ASes than China, so its line extends further to the right in the
+plots. In all cases, China can control a much larger portion of its
+paths with fewer ASes than the U.S., as expected. More interesting are
+the trends over time.  China's topology has evolved such that very few
+ASes can intercept nearly all BGP paths. For The U.S. it has become
+somewhat easier to control most of its paths but more difficult to
+intercept around 70\% of its paths and higher (yellow lines crossing
+green lines).  This suggest an expansion of interior ASes with more
+connections, as well as a strengthening of the very top ASes.
+Not all nations have developed higher chokepoint potential, e.g., Germany shows a trend
+in the opposite direction, which persists for any level of border ASes.
 
@@
 
-In our test of the relationship between Internet Freedom and national
-breakthrough potential, we plotted the Freedom on the Net score of each nation
-vs the number of border ASes that that nation needed to in order to intercept
-90\% of in-to-out paths. Additionally, we evaluated the relationship with an
-Ordinarly Least Squares (OLS) fit, and found that the relationship was
-statistically significant, with a p-value $\leq$ 0.002 for 2017. The
-relationship held for other timestamps as well. The results for 2017 are shown
-in figure \ref{fig:fotn}
+To test the correlation between Internet Freedom and NCP, we plotted
+FOTN scores vs. the number of border ASes required to intercept 90\% of in-to-out
+paths and computed the Ordinary
+Least Squares (OLS) fits, finding that the relationship is
+statistically significant (e.g., p-value $\leq$ 0.002 for 2017). The
+relationship held for other years, and 
+Fig. \ref{fig:fotn} shows a representative result (2017).  The relationship also holds for Freedom of the Press scores compared to NCP (Fig. \ref{fig:fotp}), p-value $\leq XXX$.
 
@@
 	\includegraphics[width=\linewidth]{fotn}
-	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FOTN score
-	of those nations in 2017. A higher FOTN score indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotn}
+	\caption{National Chokepoint Potential (in-to-out) vs. FOTN score
+	for 2017. Higher FOTN scores indicate less freedom. The OLS fit is also plotted.}\label{fig:fotn}
 \end{figure}
@@
 	\includegraphics[width=\linewidth]{fotp}
-	\caption{The National Breakthrough Potential (in-to-out) for nations plotted against the FPI
-	of those nations in 2017. A higher FPI indicates a less free nation. Our OLS fit is also plotted.}\label{fig:fotp}
+	\caption{The National Chokepoint Potential (in-to-out) vs. FPI
+	2017. Higher FPI indicates less freedom. The OLS fit is also plotted.}\label{fig:fotp}
 \end{figure}
 
-For each timestamp we see what appears to be two general modes of behavior.
-Nations that are not free or partly free tend to requrie few ASes to intercept
-90\% of their paths. Free nations on the otherhand require varying degrees of
-large numbers of ASes to control the same portion of their paths. There are
-interesting outliers for both situations, however. Countries like Estonia (EE)
-or Iceland (IS) are very free but require few border ASes to control most of
-their paths. The reason for these outliers is likely that their overall AS
-counts are very low. On the other hand, Russia (RU) is a very interesting
-outlier in that it is found to be not free by FOTN, but requires a large number
-of ASes to control most of its paths. This suggests that the censorship efforts
-in Russia might be of types that do not require AS level chokepoints.
+For each year there appear to be two general modes.  Nations that are
+not free or partly free tend to have high NCP, while free nations have
+variable NCPs. There are interesting outliers for both situations,
+however. Countries like Estonia (EE) or Iceland (IS) are very free but
+require few border ASes to control most of their paths. This is likely
+because their overall AS counts are low. As we saw earlier, Russia
+(RU) is considered not free by FOTN, but requires a large number of
+ASes to control most of its paths. This suggests that censorship
+in Russia is implemented without AS-level
+chokepoints.

### 2018-07-31 Paper/tool.tex
Explained CC missing problem more

@@
 registration of each AS.  To control for this possible source of
-error, we retain ASes in thepath simulation, even if they were
-registered more recently than the current timestamp being studied, but
-we do not assign these ASes chokepoint potentials. We took this approach
+error, we retain ASes in the path simulation, even if they were
+registered more recently than the current timestamp being studied. This is because
+the relationships of these ASes and paths through these ASes are still valid.
+We do not assign these ASes chokepoint potentials, as it is not clear which nation
+these ASes are registered for. We took this approach
 because the datasets of AS registrations for the various ASN
 registries tend to have more missing data at specific timestamps.
-Fortunately, ASes with registration changes after each test date tend
-to not be in the AS graph for that test. \Ben[inline]{Two things. 1)
-  This is sort of confusing, see if you can clarify it. 2) Is this a
-  big deal? How often does this happen? If it happens frequently do we
-  expect it to affect our results? }
+We do not expect this to have a considerable effect on our results
+because few ASes do not have a definite nation (in the best case in the most recent
+test this is less than 1\% of ASes, in the worst case in the earliest test this is around 12\% of ASes.)
+Additionally, not all of the ASes that are unlinked to a nation are border ASes, and those will not
+affect our results at all. Furthermore, it is unlikely that this small portion of ASes will be a majority
+of border ASes for any single country, so the country level experiments are unlikely to suffer from this
+hole in the data.
 

### 2018-07-31 Paper/background.tex
removed a few repretitive sentences

@@
 
-\Ben[inline]{We need to discuss not only the background of BGP routing here, but 
-also background about surveillance and censorship. We could probably shorten/tighten
-the BGP stuff to make room.}
-
 The interplay between the Internet and policy has a history as long as the
@@
 technical research literature in 2006 by Clayton \emph{et
-al.}~\cite{Clayton:2006:IGF:2166520.2166523}.  As of 2017, 49 countries are
-reported by the Freedom on the Net project~\cite{FOTN} to be ``partly free'' or
-``not free''.  Large-scale Internet surveillance is also prominent and has a
+al.}~\cite{Clayton:2006:IGF:2166520.2166523}. Large-scale Internet surveillance is also prominent and has a
 long history, but surveillance is not as easy to measure as censorship so only

### 2018-07-31 Paper/chokepoint.tex
Explained in-to-out and out-to-in

@@
 composed of every acceptable BGP path from the various source ASes that can
-reach the destination AS.\Ben[inline]{Is this always a tree? Seems like it
-could be a DAG sometimes (especially if random choices are made). Also do we
-need to define routing trees if we only really work on the path}  The
+reach the destination AS. Note that if multiple paths exist from a source
+to a destination, the routing tree is technicailly a DAG, but we use the traditional convention and call
+all such data structures routing trees. The
 chokepoint potential of a border AS is the ratio of the number of paths in (or out of) the country that contain the border AS to
@@
 potential for a nations is illustrated in \figurename
-\ref{fig:chokepoint}.
-
-\Ben[inline]{A couple of questions here. Why mess with incoming (out to in) and
-outgoing (in to out) measures? Can we just define it as $\frac{|p_{in}| +
-|p_{out}|}{|P_{in}| + |P_{out}|}$, that is the incoming and outgoing paths over
-the total paths? Would this be that much different? Is there are an argument to
-be made that out-to-in or in-to-out is a better? My guess is they are all
-pretty correlated, I can check on that}.
-% Formal Definition
-%\begin{equation}\label{eqn:chokePointPotential}
-%CP(b) = \frac{|\{p \in P_c \textrm{ s.t. } \{b\} \subseteq B_p\}|}{\sum_{p \in P_c}|B_p|}
-%\end{equation}
+\ref{fig:chokepoint}. We define both $cpo$ and $cpi$ because past work
+has revealed that country-level paths are often asymmetric, meaning that the
+forward path from AS $a$ to AS $b$ does not necessarily match the reverse path
+\cite{characterizingAndAvoiding}.
 

### 2018-07-31 Paper/related.tex
writing

--- 2018-07-31 Paper/related.tex
@@
 \par
-Previous studies have used BGP path models to find ASes that
-intercept a high fraction of paths. One such project in \cite{throats}
-identified that 90\% of paths on the Internet could be intercepted with only 30
-or so ASes. The researchers in \cite{throats} generated paths by first using
-only paths to top websites as defined by the Alexa top websites project,
-and then appending additional edges to those paths from the full
-AS graph according to the principles defined by Gao \cite{gao}. Many of the
-ASes that were found to intercept a large number of paths were found to be
-within nations that conduct censorship. As an extension of these results,
-another paper \cite{decoy} revealed that ASes that intercept many paths could
-be utilized for decoy routing. This approach for identifying AS level
-chokepoints has several potential pitfalls that are remedied by the approach
-taken in this paper. First, the authors in \cite{throats} did not consider that
-many of the paths found from the Alexa top websites would have destinations in
-nations that censor Internet traffic, particularly China due to its large
-Internet population. This artificially inflates the chokepoint nature of
-Chinese ASes. As an alternative, in this paper we consider paths from every
-source-destination AS pair, and we make a distinction between in-to-out paths
-and out-to-in paths. Secondly, chokepoints have previously only been
-identified at a single snapshot of the Internet. This makes it difficult to
-discuss the evolution of Internet topology, and whether or not chokepoints are
-anomalous or common. Finally, the aforementioned approach for identifying
-chokepoints does not allow a clear comparison between different nations. It
-can be said that one nation controls a large portion of Internet paths, but
-not how easily traffic directed through that country could be intercepted on a
-national level. For this, some aggregate measure across all of the border ASes
-within a country must be considered, as it is in this paper.     
+Previous studies also used BGP path models to find ASes that
+intercept a high fraction of paths, e.g.,~\cite{throats} which
+reported that 90\% of paths on the Internet could be intercepted by only 30
+or so ASes. This work generated paths starting with
+paths to top websites
+as defined by the Alexa top websites project,
+and then appending additional edges from the full
+AS graph according to the Gao algorithm \cite{gao}.
+%The
+%ASes that intercept many paths are often
+%within nations that conduct censorship.
+This work was extended
+in \cite{decoy}, which showed how ASes that intercept many paths could
+be used for decoy routing. Our paper extends this work in several ways.
+%This approach for identifying AS level
+%chokepoints has several potential pitfalls that are remedied by the approach
+%taken in this paper.
+First, 
+%the authors in \cite{throats} did not consider that
+many of the websites in the Alexa dataset are in China, due to its large
+Internet population, potentially biasing results.
+%which artificially inflates the effect of
+%Chinese ASes.
+while we consider paths between each
+source-destination AS pair.  More importantly, our work studies how chokepoints have changed over time, rather than considering 
+a single snapshot.
+%This makes it difficult to
+%discuss the evolution of Internet topology, and whether or not chokepoints are
+%anomalous or common.
+Finally, we compare chokepoints quantitatively between different
+countries by defining a measure, rather than simply identifying those
+countries with ASes that intercept the most paths.
+
+%the aforementioned approach for identifying
+%% chokepoints does not allow a clear comparison between different nations. It
+%% can be said that one nation controls a large portion of Internet paths, but
+%% not how easily traffic directed through that country could be intercepted on a
+%% national level. For this, some aggregate measure across all of
+%the border ASes
+%within a country must be considered, as it is in this paper.     
 \par
@@
 to identify where keyword filtering occurred. They found that the most
-effective ASes with which to deploy keyword filtering devices are those in the
-backbone of the Chinese AS topology. A relevant contribution of
-\cite{chinafiltering} is that, while most filtering occurs in border ASes,
-some filtering at the time of the study was controlled by non-central provincial
-ASes. China had a
-diverse strategy for Internet censorship, both targeting chokepoints and the
-Chinese provincial network, but this may have changed since. The potential for various forms of censorship in
-regards to various AS level topologies motivates the question: Is centralized
-censorship or decentralized censorship more common? Instead of directly
-identifying censorship devices on the AS graph, we instead quantify the
-chokepoint potential of ASes on the national level, and then compare that with
-qualitative Internet freedom measures and empirical censorship events.
-\par
-We are not the first to investigate the relationship between Internet freedom and
-AS-level topology. Similar techniques have been used to classify nations according
-to the connectivity of their ASes \cite{politicsrouting}. This has only been done for a
-single moment in time, however, making the results limited in terms of stability and predictability.
-Additionaly, previous work has not used country level chokepoints as the link between Internet topology
-and censorship or surveillance practices. The work in \cite{politicsrouting} chose to relate Internet topology
-to the Freedom of the Press measure from freedom house instead of the Freedom On The Net score. They chose to do this
-to include more nations. Additionally, this study did not include the United States and Russia in their experiments because they were
-outliers in regards to their topologies. Through our approach we hope to extend this previous work by finding an interesting measure for
-understanding the dynamics of all nations, as well as targeting our results more specifically to Internet freedom by using the Freedom On
-The Net score as our measure for Internet freedom. Through our techniques, we provide a simple measure that not only 
-sheds light on the relationship between topology and Internet freedom, but reveals currently free
-nations that could easily implement censorship if their governments decided to.
+effective ASes
+%with which to deploy keyword filtering devices
+are those in the backbone of the Chinese AS topology. A relevant
+contribution of \cite{chinafiltering} is that, while most filtering
+occurs in border ASes, some filtering occurs in provincial ASes. China
+had a diverse strategy for Internet censorship at the time, targeting
+both chokepoints and the Chinese provincial network, but this may have
+changed since. The potential for various forms of censorship in
+regards to various AS level topologies motivates the question: Is
+centralized censorship or decentralized censorship more common?
+Instead of directly identifying censorship devices on the AS graph, we
+instead have quantified the chokepoint potential of ASes on borders,
+and then compared that to  Internet freedom measures.
+
+\par We are not the first to investigate
+the relationship between Internet freedom and AS-level
+topology. Similar techniques have been used to classify nations
+according to the connectivity of their ASes
+\cite{politicsrouting} for a single snapshot in time,
+limiting the results. Our work focuses on country-level
+chokepoints as the link between Internet topology and censorship or
+surveillance. Finally, our study includes
+the U.S. and Russia which the earlier work did not.
 
@@
 into and out of a government's own country.  Dainotti \emph{et
-al.}~\cite{Dainotti:2011:ACI:2068816.2068818} anlyze two specific large-scale
+al.}~\cite{Dainotti:2011:ACI:2068816.2068818} analyze two specific large-scale
 Internet disruptions at the routing level.  Khattak \emph{et
@@
 censorship at the ISP level in Pakistan.
+
+While there have been attempts to characterize methods for
+censoring Internet content~\cite{Tschantz2016a,Khattak2016a} there is no
+comprehensive list of all the different ways a state actor can manipulate
+traffic.  For some methods, the AS graph is relevant, such as IP address
+blacklisting, traffic throttling~\cite{DBLP:journals/corr/Anderson13}, URL or
+packet filtering, packet injection, physically shutting down infrastructure, and
+BGP attacks.  For other methods, the AS graph is less relevant, such as creating
+internal national networks~\cite{DBLP:journals/corr/abs-1209-6398}, portal
+censorship such as search engine filtering and social media post deletion,
+propaganda campaigns, and manipulation of the DNS system.  Our work is
+complementary to other studies that focus more on Internet routing's effect on
+high-level content than on a country's desire to enforce policies on all traffic
+that crosses their borders, such as Edmundson \emph{et
+al.}~\cite{Edmundson:2018:NHI:3209811.3211887}.

### 2018-07-31 Paper/tool.tex
removed missing equation reference
clarified some of Ben's questions

@@
 uses an algorithm based on BGPSim \cite{quicksand} to
-generate a set of routing trees. \Ben[inline]{Need a short description of how
-BGPSim works, maybe below.}. \toolname{} 
+generate a set of routing trees. \toolname{} 
 calclates both AS and national chokepoint potential for any country.
@@
 component of \toolname{} addresses several limitations that prevented
-us using BGPSim directly: (1) BGPSim returns a set of ASes for each
-path it considers but not the order in which they are visited; (2)
+us using BGPSim directly: (1) BGPSim was not used to test statistics on paths, so it
+doesn't return ordered paths for calculating AS-level statistics; (2)
 Once routing trees are determined, they cannot be accessed later
@@
 we implemented our routing simulation in
-Python. \Ben[inline]{Why does BGPSim not return ordered paths, this
-  seems trivial, but I don't know the implementation details, was it
-  just an oversight or did you do something cool?}\toolname{} returns
+Python. \toolname{} returns
 ordered paths from its routing trees, saves routing trees to disk
@@
 exactly one path between each AS-pair considered, even if multiple options exist in the
-routing tree.\Ben[inline]{Put stuff in about tiebreaking being rare and
-unlikely to affect results} Once the path has been determined, it is traversed to identify border nodes and increment their counts, both
+routing tree. Tiebreaks do not have a noticable effect on our results. For instance, the maximum
+standard deviation for the chokepoint potential of any AS in 5 runs of one of our timestamps was 
+$\leq$ 0.005, meaning the chokepoint potentials for each AS barely changed depending on the tiebreak
+decisions.
+Once the path has been determined, it is traversed to identify border nodes and increment their counts, both
 for outgoing and incoming paths.  Thus, only one traversal is conducted
 per path. Additionally, the number of paths of each type that belong to each
-nation is tallied, which becomes the denominator in equation
-\ref{eqn:chokePointPotential}.  \toolname{} takes the resulting chokepoint
+nation is tallied. \toolname{} takes the resulting chokepoint
 potentials and generates national chokepoint potentials for each nation.

### 2018-07-31 Paper/chokepoint.tex
Addressed more comments

@@
 the aggregate measure of chokepoint potential, national chokepoint potential (NCP)
-for ranking and comparing nations according to the strength of their border AS chokepoints.
+for ranking and comparing countries according to the strength of their border AS chokepoints.
 
@@
 the number of border ASes of a country to all ASes in that country.
-%hints toward the abillity of that nation to intercept information flows across
+%hints toward the abillity of that country to intercept information flows across
 %their borders.
@@
 %decreased. 
-It would seem that this fact reflects an ever increasing capability for nations
+It would seem that this fact reflects an ever increasing capability for countries
 to intercept AS-level paths.  However, an important consideration is not only
@@
 generally occurs at a national level so we further define an aggregate measure
-for nations, which we call \emph{national chokepoint potential}.
+for countries, which we call \emph{national chokepoint potential}.
 
 % % Why Border ASes?
-% In order to identify AS chokepoints and compare nations, we need a measure that can be calculated from the many paths between
-% ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another nation. 
+% In order to identify AS chokepoints and compare countries, we need a measure that can be calculated from the many paths between
+% ASes. First, we decided to use a measure that is evaluated only on border ASes, or ASes that lie one hop from an AS belonging to another country. 
 % The evolution of AS relationships and national boundaries on the AS graph hints that border ASes are an important feature in regards to the flow
@@
 %intercepted by all border ASes belonging to
-%the same nation.
+%the same country.
 Formally, for country $c$, let $P = \{p_{i \rightsquigarrow
@@
 $cp(\cdot)$, of a border AS $a_u$ with $g(a_u) = c$ is
-\Steph[inline]{How do we count a parth that originates in a country, routes through a different country, and terminates in the original country?  That case is relevant in US law.}
 \begin{equation}
@@
 The incoming chokepoint
-potential for a nations is illustrated in \figurename
+potential for a countries is illustrated in \figurename
 \ref{fig:chokepoint}. We define both $cpo$ and $cpi$ because past work
@@
 %, as in, all of the border ASes
-%collectively control the flow of information over the nation's border.
+%collectively control the flow of information over the country's border.
 We define the aggregate \emph{national chokepoint potential} as the number of
@@
 %this one being a national measure. We define the \emph{national breakthrough
-%potential} of a nation to be the minimum number of ASes required to intercept a
+%potential} of a country to be the minimum number of ASes required to intercept a
 %selected percentage of AS-level paths. 
@@
 90\% path control indicates the percentage of paths into (our out of) a country.
-%that a single nation intercepts 90\% of the paths
-%that either enter said nation (for the out-to-in case) or exit said nation (for
+%that a single country intercepts 90\% of the paths
+%that either enter said country (for the out-to-in case) or exit said country (for
 %the in-to-out case).
@@
 could be defined as the fraction of paths controlled by $j$ ASes. Our results
-are substantively the same for either definition.\Ben[inline]{Check to make
-sure}}.
+are substantively the same for either definition.}
 
@@
 fraction rather than a percentage within a country to make
-comparisons between nations more equitable.
-Consider, for instance, the United States, which has many more ASes than
-most nations. If the United States and another nation both required the same
+comparisons between countries more equitable.
+Consider, for instance, the U.S., which has many more ASes than
+most countries, and China, which has a much smaller number of ASes. If the United States and China 
+were to both require the same
 percentage of border ASes to intercept 90\% of paths, the similarity of these
 nations would be misleading because the cost of controlling the larger absolute number of ASes
-would be significantly higher.\Ben[inline]{we need to actually
-pick another nation where this is true to make this example illustrative.} Our technique provides a comparison more likely to
+would be significantly higher. Our technique provides a comparison more likely to
 reflect reality. 

### 2018-07-31 Paper/results.tex
Addressed more comments

@@
 Least Squares (OLS) fits, finding that the relationship is
-statistically significant (e.g., p-value $\leq$ 0.002 for 2017). The
+statistically significant (e.g., p-value $\leq$ 0.0001 for 2017). The
 relationship held for other years, and 
-Fig. \ref{fig:fotn} shows a representative result (2017).  The relationship also holds for Freedom of the Press scores compared to NCP (Fig. \ref{fig:fotp}), p-value $\leq XXX$.
+Fig. \ref{fig:fotn} shows a representative result (2017).  The relationship also holds for Freedom of the Press scores compared to NCP (Fig. \ref{fig:fotp}), p-value $\leq$ 0.001.
 

### 2018-07-31 Paper/intro.tex
BGP Path Guide -> BGP-SAS
Changed breakthrough to chokepoint in intro

@@
 We developed a
-suite of tools, called \toolname{}, for studying national chokepoints on
+suite of tools, called BGP Simulation Analysis and Storage (\toolname{}), for studying national 
+chokepoints on
 the AS graph efficiently. To illustrate these ideas and tools, we
@@
 to the problems we are investigating; Section 3 introduces and defines the measures of chokepoint
-potential and national breakthrough potential; Section 4 details our tool for simulating and evaluating 
+potential and national chokepoint potential; Section 4 details our tool for simulating and evaluating 
 AS topologies \toolname{}; Section 5

### 2018-07-31 Paper/results.tex
Fixed decline and increase with flipped plots

@@
 To study changes over time, we first plotted the
-number of border ASes required to intercept 90\% of in-to-out paths
+national chokepoint potential of each of the 8 selected countries
 for each year (\figurename \ref{fig:nodesFree} and \figurename
-\ref{fig:nodesCensor}).  The results are striking with a steep decline
-over time for some nations (China and Turkey) and increases for others
+\ref{fig:nodesCensor}).  The results are striking with a steep increase
+over time for some nations (China and Turkey) and decreases for others
 (Germany and France), while some countries remained relatively
@@
 globally over time, it is likely that even nations that look stable
-have in fact become more 'chokepointy' over time as some border nodes
-must be intercepting more paths.  This fact makes the decline observed
+have in fact become more 'chokepointy' over time as some border ASes
+must be intercepting more paths.  This fact makes the increased NCP observed
 for some countries even more striking.

### 2018-07-31 Paper/conclusion.tex
ran out of gas for a soaring finish

@@
 % Chokepoint Potential
-In this paper we have introduced two novel measures, chokepoint potential, on single border ASes,
-and national breakthrough potential, on entire nations, that together allow us to
-evaluate the level of difficulty that a government faces when trying to censor the Internet
-based entirely on AS-level topology. Our technique for evaluating AS-level topology using
-\toolname{} provides a monitoring capability of BGP dynamics that was previously not possible.
+This paper addresses the question of how Internet structure relates to 
+international boundaries and how it is changing over time.  By defining chokepoint
+potential, both for individual border ASes and aggregated by country,
+we provide a quantitative way to measure these effects in terms of the ease
+with which a government can control routing paths across its borders. 
+Our technique for generating BGP paths and evaluating chokepoints
+using \toolname{} provides public domain software and routing trees for the entire AS graph.
+%a monitoring capability of BGP dynamics that was previously not possible.
 We have taken advantage of efficient simulation, standard AS relationship datasets, and cross-platform
-design principles so that this tool is readily deployable for future research.
+design principles so that this tool will be readily deployable for future research.
 % Ecolution over time
-\par
-Our application of \toolname{} to investigate evolution of the global AS graph over time shows us
-interesting trends in AS topologies. We have identified the apparent lack of major nations to develop
-more open Internet borders and gateways, despite the increasing connectivity and size of the AS graph. Additionally,
-we find that the chokepoints in some nations known to conduct censorship and surveillance have increased in strength.
+\par We used \toolname{} to investigate how the global AS graph has
+evolved over time, finding interesting trends and outliers. These
+studies reveal a tendency in large nations to concentrate paths across
+their borders into a small number of ASes, despite the increasing
+connectivity and size of the AS graph. Additionally, we find that border
+chokepoints in some nations known to conduct censorship and
+surveillance have steadily increased in strength.
 % Internet Freedom
 \par
-We have validated national breakthrough potential as having a relationship to actual censorship by showing
-the significance of the relationship between the chokepoint potential of nations and the evaluations of freedom provided
-by both freedom on the net and freedom of the press. This relationship implies that when censorship and surveillance are being evaluated
-at the national level, it is worthwhile to consider the border AS chokepoints that are resident to the nation in question.
+We studied the relationship between NCP and two evaluations of openness, Freedom of the Net  and Freedom of the press, finding stastistically significant relationships in both cases.
+%These results suggest that it is worthwhile to consider border AS chokepoints when studying
+This relationship implies that when censorship and surveillance are
+being evaluated at the national level, it is worthwhile to consider
+the border AS chokepoints that are resident to the nation in question.

### 2018-08-01 Paper/intro.tex
more removing for review
merged some stuff

--- 2018-07-31 Paper/intro.tex
@@
 are mediated by the Internet, most of us have little choice about
-whether or not to rely on the Internet.  According to the
+whether or not to rely on it.  According to the
 International Telecommunication Union (ITU), the number of individual
@@
 focusing on how the global AS graph layout indicates that interdomain routing
-for many Internet users depends on the infrastructure of other nations. We take
-a different approach by developing a technique to monitor the amount of paths intercepted
+for many Internet users depends on the infrastructure of other nations. Here we take
+a different approach, presenting a technique to monitor the of paths intercepted
 by border ASes, allowing us to compare nations according to how BGP paths enter and exit
@@
 \par
-In this paper, we focus on the Autonomous Systems (AS) level of the Internet.
-Autonomous Systems are groups of routers under common management, such as the network of
+In this paper, we focus on the AS level of the Internet.
+ASes are groups of routers under common management, such as the network of
 a university or an Internet Service Provider (ISP).
@@
 The structural properties of national subgraphs have evolved differently
-from nation to nation whether from economic decision making, infrastructural necessities, 
+from nation to nation, whether from economic decision making, infrastructural necessities, 
 or efforts to build a powerful censorship and surveillance network \cite{economicPeering, irancensor}.
@@
 
-\par Every nation has a different layout and number of ASes, and these
-national networks connect to foreign ASes in unique ways. Considering
-just censorship and surveillance, different nations use different
-strategies for this form of content monitoring. For instance, China
-both conducts keyword filtering in border ASes and in internal
-provincial ASes \cite{chinafiltering}, while Iran routes its Internet
-traffic through a centralized facility \cite{irancensor}.  A picture
-of a nation's content monitoring capabilities then is not complete
-without examining its underlying AS topology.  Having the ability to
-investigate the AS topology of a nation will both help researchers
-identify nations that could, for example, easily conduct censorship
-and also provide possible insight into what kind of censorship is
-likely being conducted.  More generally, we are interested in the extent to which the Internet is changing from a borderless structure into one that reflects national boundaries.
+\par Every nation has a different number and organization of ASes, and
+these national networks connect to foreign ASes in unique
+ways. Considering just censorship and surveillance, different nations
+use different strategies for this form of content monitoring. For
+instance, China conducts keyword filtering in both border ASes and
+internal provincial ASes \cite{chinafiltering}, while Iran routes its
+Internet traffic through a centralized facility \cite{irancensor}.  A
+picture of a nation's content monitoring capabilities then is not
+complete without examining its underlying AS topology.  Having the
+ability to investigate the AS topology of a nation will both help
+researchers identify nations that could, for example, easily conduct
+censorship and also provide possible insight into what kind of
+censorship is likely being conducted.  More generally, we are
+interested in the extent to which the Internet is changing from a
+borderless structure into one that reflects national boundaries.
 
@@
 %at the AS-level by
-We focus on border ASes, or ASes that can connect directly to at least
-one AS from another nation and introduce a measure called \emph{chokepoint potential}, which quantifies the percentage of BGP paths into or out of a country which pass through that AS.
+We focus on border ASes, i.e. ASes that can connect directly to at least
+one AS from another nation and introduce a measure called \emph{chokepoint potential}, which quantifies the percentage of BGP paths into or out of a country that pass through the AS.
 At the country level, we are interested in aggregating chokepoint potential in different ways to quantify the tendency of all BGP paths crossing the border to pass through a small (or large) number of chokepoints (border ASes).
-We sometimes refer to this as \emph{national chokepoint potential}.
+We  refer to this as \emph{national chokepoint potential (NCP)}.
 %We can then aggregate this measure and define \emph{national breakthrough potential} to
@@
 We developed a
-suite of tools, called \toolname{}, for studying national chokepoints on
+suite of tools, called BGP Simulation, Analysis, and Storage (\toolname{}), for studying national 
+chokepoints on
 the AS graph efficiently. To illustrate these ideas and tools, we
-study how national chokepoint potential correlates with two independent
-measures of civil liberty, finding that a significant relationship exists between national chokepoint potential and each measure.
+study how NCP correlates with two independent
+measures of civil liberty, finding that a significant relationship exists between NCP and each measure.
 The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasetes for simulating Border Gateway Protocol (BGP) paths,
-determining national chokepoint potential, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
+determining NCP, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
 %Our measure of border ASes, chokepoint potential, defined in detail
@@
  \item A study of how national AS-level chokepoints have changed over the past nine years, demonstrating that the Internet has evolved to facilitate stronger
- AS-level chokepoints for some countries and lead to more open flow of information across borders for others.
+ AS-level chokepoints for some countries and led to the more open flow of information across borders for others.
  \item An open-source tool, \toolname{}, and datasets for simulating BGP paths and evaluating chokepoints efficiently at different time points for the entire Internet.  
- \item A study that reveals a significant relationship between chokepointpotential and Internet freedom, as measured by two qualitative sources.
+ \item A study showing a significant relationship between chokepoint potential and Internet freedom, as measured by two qualitative sources.
 This result suggests that chokepoint potential can be interpreted as an indicator of a country's capability for conducting censorship or surveillance of its international Internet traffic.
@@
 \par
-The layout of the rest of this paper is as follows: Section 2 provides relavent background information
-to the problems we are investigating; Section 3 introduces and defines the measures of chokepoint
-potential and national breakthrough potential; Section 4 details our tool for simulating and evaluating 
-AS topologies \toolname{}; Section 5
-explains our experimental setups; Sections 6, 7, and 8 provide results, discussion, and related work
-respectively.
+The remainder of the paper is organized as follows: Section II provides relevent background for
+the problems we are investigating; Section III introduces and defines the measures of chokepoint
+potential and national chokepoint potential; Section IV describes \toolname{} for simulating and evaluating 
+BGP networks; Section V
+explains the experimental setup and data sources; Sections VI, VII, and VIII contain experimental results, related work, and discussion,
+respectively.  Section IX concludes the paper.

### 2018-08-01 Paper/chokepoint.tex
Spellcheck run through

@@
 reach the destination AS. Note that if multiple paths exist from a source
-to a destination, the routing tree is technicailly a DAG, but we use the traditional convention and call
+to a destination, the routing tree is technically a DAG, but we use the traditional convention and call
 all such data structures routing trees. The
@@
 $c$ and routing to a different country, which we will refer to as
-\textit{outgoing} chokepoint potential or $cpo$. We similarily define
+\textit{outgoing} chokepoint potential or $cpo$. We similarly define
 \textit{incoming} chokepoint potential or $cpi$, by redefining $P$ as $\{p_{i

### 2018-08-01 Paper/conclusion.tex
Spellcheck run through

@@
 of overall chokepoint potential, and information such as geography and the physical locations of routers does not add meaningfully to the analysis without considering every individual piece of Internet
-infrastructue on a case-by-case basis.
+infrastructure on a case-by-case basis.
 Additionally, the dramatic shifts in chokepoint potential over time depicted in our analysis are unrelated to constant geography, so our AS-level study is important
@@
 actual censorship events and the shifts in the AS graph. The Open Observatory of Network Interference, or OONI, \cite{OONI} provides Internet users around
-the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anamolies of various types, and the results are
+the world with a tool called the ooniprobe. The ooniprobe lets users run a suite of tests to identify censorship anomalies of various types, and the results are
 recorded in the large OONI database. Matching up changes in OONI measurements, such as increased censorship campaigning in an authoritarian nation, with shifts
@@
 \par
-We studied the relationship between NCP and two evaluations of openness, Freedom of the Net  and Freedom of the press, finding stastistically significant relationships in both cases.
+We studied the relationship between NCP and two evaluations of openness, Freedom of the Net  and Freedom of the press, finding statistically significant relationships in both cases.
 %These results suggest that it is worthwhile to consider border AS chokepoints when studying

### 2018-08-01 Paper/intro.tex
Spellcheck run through

@@
 measures of civil liberty, finding that a significant relationship exists between NCP and each measure.
-The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasetes for simulating Border Gateway Protocol (BGP) paths,
+The paper extends earlier research on AS topology in several ways: it introduces chokepoint potential, it presents open-source cross-platform tools and datasets for simulating Border Gateway Protocol (BGP) paths,
 determining NCP, and analyzing changes over time.  In addition, we report and analyze data for several countries of interest.
@@
 \par
-The remainder of the paper is organized as follows: Section II provides relevent background for
+The remainder of the paper is organized as follows: Section II provides relevant background for
 the problems we are investigating; Section III introduces and defines the measures of chokepoint

### 2018-08-01 Paper/results.tex
Spellcheck run through

@@
 
-For each year there appear to be two general modes.  Nations that are
+The trend holds for all years investigated that countries that are
 not free or partly free tend to have high NCP, while free nations have
 variable NCPs. There are interesting outliers for both situations,
-however. Countries like Estonia (EE) or Iceland (IS) are very free but
+however. Countries like Estonia and Iceland are very free but
 require few border ASes to control most of their paths. This is likely
 because their overall AS counts are low. As we saw earlier, Russia
-(RU) is considered not free by FOTN, but requires a large number of
+is considered not free by FOTN, but requires a large number of
 ASes to control most of its paths. This suggests that censorship

### 2018-08-01 Paper/tool.tex
Minor trim in tool section
Spellcheck run through

@@
 generate a set of routing trees. \toolname{} 
-calclates both AS and national chokepoint potential for any country.
+calculates both AS and national chokepoint potential for any country.
 
@@
 Additionally, not all of the ASes that are unlinked to a nation are border ASes, and those will not
-affect our results at all. Furthermore, it is unlikely that this small portion of ASes will be a majority
-of border ASes for any single country, so the country level experiments are unlikely to suffer from this
-hole in the data.
+affect our results at all.
 
@@
 exactly one path between each AS-pair considered, even if multiple options exist in the
-routing tree. Tiebreaks do not have a noticable effect on our results. For instance, the maximum
+routing tree. Tiebreaks do not have a noticeable effect on our results. For instance, the maximum
 standard deviation for the chokepoint potential of any AS in 5 runs of one of our timestamps was 
@@
 For each dataset, this is a one-time cost because the routing trees are stored externally.
-[URL blineded for peer review]
+[URL blinded for peer review]

### 2018-08-01 Paper/tool.tex
small typo changes

@@
 determine chokepoint potentials or by other researchers interested in other questions. To calculate chokepoint potential, \toolname{} iterates over every path
-between each AS-pair. Because we use the same random tie-break method as BGPSim,this process returns 
+between each AS-pair. Because we use the same random tie-break method as BGPSim, this process returns 
 exactly one path between each AS-pair considered, even if multiple options exist in the

### 2018-08-01 Paper/intro.tex
typos

@@
 for many Internet users depends on the infrastructure of other nations. Here we take
-a different approach, presenting a technique to monitor the of paths intercepted
+a different approach, presenting a technique to monitor the number of paths intercepted
 by border ASes, allowing us to compare nations according to how BGP paths enter and exit
